### 2022-11-15 Scripts/data_cleaning.Rmd
Adding event counts and protest sizes from Mass Mobilization dataset (note: quality checks still needed)

---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
```

Loading data

Note: check on how massmob data merges with nelda. We want cases where this is no protest after the election to score as 0.

```{r}
nelda <- read_xls(here::here("Data", "NELDA.xls"))
ecav <- read_xls(here::here("Data", "ECAV datatset_Version 1.2.xls"))
vdem.data <- vdem #At the time of coding, this was v. 11.1
lji <- read.csv(here::here("Data", "LJI-estimates-20140422.csv"))
massmob <- read.csv(here::here("Data", "mmALL_073119_csv.csv"))
```

Cleaning nelda variables--this will have to be done for all used variables
```{r}
nelda <- nelda %>% rename(first.mp.election = nelda2)
nelda <- nelda %>% rename(protests = nelda29)
nelda <- nelda %>% rename(inc.fav.cancel = nelda34)
nelda <- nelda %>% rename(inc.lose = nelda24)
nelda <- nelda %>% rename(inc.termlimit = nelda8)
nelda <- nelda %>% rename(econ.crisis.nelda = nelda18)
nelda <- nelda %>% rename(opp.votegain = nelda27)
nelda <- nelda %>% rename(nelda.protests.fraud = nelda30)
nelda <- nelda %>% rename(new.elections.held = nelda37)
nelda <- nelda %>% rename(inc.replaced.prot = nelda41)

nelda <- nelda %>% mutate(first.mp.election = ifelse(first.mp.election == "yes", 1, ifelse(first.mp.election == "no", 0, NA)))
nelda <- nelda %>% mutate(protests = ifelse(protests == "yes", 1, ifelse(protests == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel = ifelse(inc.fav.cancel == "yes", 1, ifelse(inc.fav.cancel == "no", 0, NA)))
nelda <- nelda %>% mutate(inc.fav.cancel.protests = ifelse(nelda35 == "yes", 1, 0))
nelda <- nelda %>% mutate(opp.fav.cancel = ifelse(nelda32 == "yes", 1, 0))
nelda <- nelda %>% mutate(inc.lose = ifelse(inc.lose == "yes", 1, ifelse(inc.lose == "no", 0, NA)))

nelda <- nelda %>% mutate(inc.termlimit.num = ifelse(types == "Legislative/Parliamentary" & inc.termlimit == "N/A", 0, ifelse(types == "Executive" & inc.termlimit == "yes", 1,                                                                                                                       ifelse(types == "Executive" & inc.termlimit == "no", 0, NA))))

nelda <- nelda %>% mutate(econ.crisis.nelda = ifelse(econ.crisis.nelda == "yes", 1, ifelse(econ.crisis.nelda == "no", 0, NA)))
nelda <- nelda %>% mutate(opp.votegain = ifelse(opp.votegain == "yes", 1, ifelse(opp.votegain == "no", 0, NA)))
nelda <- nelda %>% mutate(nelda.protests.fraud = ifelse(protests == 0, 0, ifelse(nelda.protests.fraud == "yes", 1, ifelse(nelda.protests.fraud == "no", 0, NA))))
nelda <- nelda %>% mutate(inc.replaced.prot = ifelse(inc.replaced.prot == "yes", 1, 0))
nelda <- nelda %>% mutate(new.elections.held = ifelse(new.elections.held == "yes", 1, 0))

```

Coding anti-gov violence from ecav

```{r}
## Coding anti-government initiation of violence

ecav <- ecav %>% mutate(antigov.violence = 
                          ifelse(EventViolence == 0, 0,
                                 ifelse(ViolenceInitiator == "-99", NA,
                                                  ifelse(ViolenceInitiator == ".1" & Actor1Side == 1, 1, 
                                                  ifelse(ViolenceInitiator == ".2" & Actor2Side == 1, 1,
                                                         ifelse(ViolenceInitiator == "1.1" & Target1Side == 1, 1,
                                                                ifelse(ViolenceInitiator == "2.1" & Target2Side == 1, 1, 0)))))))
```

Adding a vdem ID to the other packages

Note: the missing values from NELDA are cases not included in VDEM (e.g. Bahamas, Dominica.)
```{r}
nelda <- nelda %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "gwn", destination = "vdem"))

lji <- lji %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "cown", 
                                               destination = "vdem"))
massmob <- massmob %>% mutate(country_id = countrycode(sourcevar = ccode, origin = "p4n", 
                                               destination = "vdem"))

nelda <- nelda %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
lji <- lji %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
vdem.data <- vdem.data %>% mutate(country_id_year = paste(country_id, year, sep = "_"))
massmob <- massmob %>% mutate(country_id_year = paste(country_id, year, sep = "_"))

```

VDEM variables to lag
v2x_polyarchy, v2x_libdem, v2xpartipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat

The approach below works for lagging variables by group. Next step is to lag all of the above variables, lag LJI variables, and then merge with nelda.

```{r}
vdem.recent <- vdem.data %>% filter(year >= 1944)

vdem.recent <- vdem.recent %>% mutate(elcompindex.m = ifelse(v2xel_frefair == 0, 0, v2x_mpi / v2xel_frefair))
vdem.recent$v2elirreg.inv <- vdem.recent$v2elirreg * -1

vdem.recent$reform_positive <- NA
vdem.recent$reform_positive[vdem.recent$v2jureform_ord <= 1] <- 0
vdem.recent$reform_positive[vdem.recent$v2jureform_ord == 2 ] <- 1

vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elintmon.lag = lag(v2elintmon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(reform_positive.lag = lag(reform_positive))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(v2elirreg.inv.lag = lag(v2elirreg.inv))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elcompindex.m.1lag = lag(elcompindex.m))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(polyarchy.1lag = lag(v2x_polyarchy))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(libdem.1lag = lag(v2x_libdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partipdem.1lag = lag(v2x_partipdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(delibdem.1lag = lag(v2x_delibdem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(egaldem.1lag = lag(v2x_egaldem))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(altinf.1lag = lag(v2x_freexp_altinf))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(api.1lag = lag(v2x_api))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mpi.1lag = lag(v2x_mpi))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frassoc.1lag = lag(v2x_frassoc_thick))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(suffr.1lag = lag(v2x_suffr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(frefair.1lag = lag(v2xel_frefair))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(elecoff.1lag = lag(v2x_elecoff))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(liberal.1lag = lag(v2x_liberal))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jucon.1lag = lag(v2x_jucon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partip.1lag = lag(v2x_partip))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(cspart.1lag = lag(v2x_cspart))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(eqdr.1lag = lag(v2xeg_eqdr))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(oppaut.1lag = lag(v2psoppaut))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jureform.1lag = lag(v2jureform))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupurge.1lag = lag(v2jupurge))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupoatck.1lag = lag(v2jupoatck))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jupack.1lag = lag(v2jupack))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juaccnt.1lag = lag(v2juaccnt))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juhcind.1lag = lag(v2juhcind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(juncind.1lag = lag(v2juncind))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(mecenefm.1lag = lag(v2mecenefm))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdppcln.1lag = lag(e_migdppcln))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(gdpgro.1lag = lag(e_migdpgro))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(inflat.1lag = lag(e_miinflat))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(physinteg.1lag = lag(v2x_clphy))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcorrupt.lag = lag(v3lgcrrpt))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(exbribe.lag = lag(v2exbribe))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(exembez.lag = lag(v2exembez))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(pubseccorrup.lag = lag(v2excrptps))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partybarriers.lag = lag(v2psbars))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(natlpartyorg.lag = lag(v2psorgs))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(partybranch.lag = lag(v2psprbrch))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(csrepress.1lag = lag(v2csreprss))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(osorg.1lag = lag(v2cseeorgs))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(civlib.1lag = lag(v2x_clpol))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(jucon.1lag = lag(v2x_jucon))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(legcon.1lag = lag(v2xlg_legcon))

vdem.recent <- vdem.recent %>% group_by(COWcode) %>% mutate(opposition.oversight.lag = lag(v2lgoppart))

##Leading variables for protest after election years
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(prodem.protest.1lead = lead(v2cademmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(proauth.protest.1lead = lead(v2caautmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(protest.scale.1lead = lead(v2cagenmob))
vdem.recent <- vdem.recent %>% group_by(country_id) %>% mutate(concen.protest.1lead = lead(v2caconmob))

```
Subtracting election integrity variable from additive polyarchy index and judicial constraints from liberal component index
```{r}
vdem.recent <- vdem.recent %>% mutate(polyarchy_no_ei = v2x_api - (.25*v2xel_frefair))

vdem.recent <- vdem.recent %>% mutate(liberal_no_jucon = (v2xcl_rol + v2xlg_legcon)/2)
```

Imporant variables to include but not lag: v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm

```{r}
lji <- lji %>% group_by(ccode) %>% mutate(lji.lag = lag(LJI))
```

Cleaning nelda date variable and massmob date
```{r}
nelda <- nelda %>% mutate(mmdd = ifelse(mmdd < 1000, paste0(0, mmdd), mmdd))
nelda <- nelda %>% mutate(election_mdy = paste0(mmdd, year))
nelda <- nelda %>% mutate(election_mdy = mdy(election_mdy))

massmob <- massmob %>% mutate(startday = ifelse(startday < 10, paste0(0, startday), startday))
massmob <- massmob %>% mutate(startmonth = ifelse(startmonth < 10, paste0(0, startmonth), startmonth))
massmob <- massmob %>% mutate(protest_mdy = ifelse(is.na(startday) == T | is.na(startmonth) == T, NA,  paste0(startmonth, startday, year)))

massmob <- massmob %>% mutate(protest_mdy = mdy(protest_mdy))
massmob <- massmob %>% filter(is.na(protest_mdy) == F)
```

Merging

```{r}
lji <- lji %>% filter(is.na(country_id) == FALSE)
nelda <- nelda %>% filter(is.na(country_id) == FALSE)
massmob <- massmob %>% filter(is.na(country_id) == FALSE)

#Lagging econ crisis

nelda <- nelda %>% group_by(country_id) %>% mutate(econ.crisis.nelda.1lag = lag(econ.crisis.nelda))

#Merging
nelda2 <- nelda %>% merge(lji %>% dplyr::select(LJI, lji.lag, country_id_year), by = "country_id_year")

nelda2 <- nelda2 %>% merge(vdem.recent %>% dplyr::select(v2elintmon.lag, v2elintmon, v2eltvrexo, v2elvotlrg, v2elvotsml, v2ellostlg, v2ellostsl, v2ellostsm, v2ellostss, v2ellovtlg, v2ellovtsm, v2x_polyarchy, v2x_libdem, v2x_partipdem, v2x_delibdem, v2x_egaldem, v2x_freexp_altinf, v2x_api, v2x_mpi, v2x_frassoc_thick, v2x_suffr, v2xel_frefair, v2x_elecoff, v2x_liberal, v2x_jucon, v2x_partip, v2x_cspart, v2xeg_eqdr, v2psoppaut, v2jureform, v2jupurge, v2jupoatck, v2jupack, v2juaccnt, v2juhcind, v2juncind, v2mecenefm, e_migdppcln, e_migdpgro, e_miinflat, v2xlg_legcon, legcon.1lag, polyarchy.1lag, libdem.1lag, partipdem.1lag, 
                                                  delibdem.1lag,
                                                  elcompindex.m.1lag,
                                                  opposition.oversight.lag,
                                                  v2elirreg.inv,
                                                  reform_positive.lag,
                                                  e_mipopula,
                                                  e_miurbani,
                                                  e_miurbpop,
                                                  e_wb_pop, #Population
                                                  v2xps_party, #Party institutionalization
                                                  #v2regoppgroupssize, #Opp group size (categorical)
                                                  v2pscohesv, #Legislative party cohesion
                                                  v2exl_legitratio, 
                                                  v2exl_legitperf,
                                                  v2exl_legitlead,
                                                  v2exl_legitideol,
                                                  v2xnp_regcorr,
                                                  v2xeg_eqprotec,
                                                  v2xeg_eqdr,
                                                  v2xeg_eqaccess,
                                                  v3lgcrrpt,
                                                  legcorrupt.lag,
                                                  exbribe.lag,
                                                  v2exbribe,
                                                  exembez.lag,
                                                  v2exembez,
                                                  pubseccorrup.lag,
                                                  v2excrptps,
                                                  partybarriers.lag,
                                                  v2psbars,
                                                  natlpartyorg.lag,
                                                  v2psorgs,
                                                  partybranch.lag,
                                                  v2psprbrch,
                                                  v2stfisccap_ord,
                                                  v2stfisccap,
                                                  v2pepwrses, #Power distribution by wealth
                                                  v2pepwrsoc, #Power distribution by social group
                                                  v2csreprss,
                                                  csrepress.1lag,
                                                  osorg.1lag,
                                                  v2cseeorgs,
                                                  v2x_clpol,
                                                  civlib.1lag,
                                                  jucon.1lag,
                                                  legcon.1lag,
                                                  v2ex_elechog,
egaldem.1lag ,
altinf.1lag,
api.1lag,
mpi.1lag,
frassoc.1lag,
suffr.1lag,
frefair.1lag,
elecoff.1lag,
liberal.1lag,
jucon.1lag,
partip.1lag,
cspart.1lag,
eqdr.1lag,
oppaut.1lag,
jureform.1lag,
jupurge.1lag,
jupoatck.1lag,
jupack.1lag,
juaccnt.1lag,
juhcind.1lag,
juncind.1lag,
mecenefm.1lag,
gdppcln.1lag,
gdpgro.1lag,
inflat.1lag,
v2elvotbuy,
v2elirreg,
v2elintim, v2elfrfair,
physinteg.1lag, v2x_clphy,
polyarchy_no_ei,
liberal_no_jucon, v2elaccept,
v2cademmob, #pro-democracy mass mob
v2caautmob, #pro-autoc mass mob
v2cagenmob, #scale of mobilization
v2caconmob, #concentration of mobilization in the capital
v2x_regime,
prodem.protest.1lead,
proauth.protest.1lead,
protest.scale.1lead,
concen.protest.1lead,
country_id_year), by = "country_id_year")

nelda2 <- nelda2 %>% mutate(winner.margin = ifelse(types == "Executive", v2elvotlrg/100, v2ellovtlg/100))
nelda2 <- nelda2 %>% mutate(presidential = ifelse(types == "Executive", 1, 0))

#

nelda2 <- nelda2 %>% mutate(weakstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 0 | v2stfisccap_ord == 1, 1, 0)))
nelda2 <- nelda2 %>% mutate(rentierstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 2, 1, 0)))
nelda2 <- nelda2 %>% mutate(taxstate = ifelse(is.na(v2stfisccap_ord) == T, NA, ifelse(v2stfisccap_ord == 3 | v2stfisccap_ord == 4, 1, 0)))

#nelda3 <- nelda2 %>% merge(massmob, by = "country_id_year")

nelda3 <- nelda %>% dplyr::select(ccode, country, year, electionid, election_mdy)
nelda3 <- nelda3 %>% mutate(election_year = year)
nelda3 <- massmob %>% left_join(nelda3, by = c("ccode", "year")) #Nelda3 is in progress, need to figure out a) if needed, b) how to include pre-election and post-election mobilization
```

Cutting out protests before elections for nelda 3, 
```{r}
#Add electionid to the year before the election and year after (so that grouping variable can capture those protests if needed)
nelda3 <- nelda3 %>% group_by(ccode) %>% mutate(election_date_upcoming = as_date(ifelse(is.na(election_year)==F, election_mdy, 
                                                  ifelse(is.na(election_year)==T & is.na(lead(election_mdy))==F,
                                                                          lead(election_mdy), NA))))

nelda3 <- nelda3 %>% group_by(ccode) %>% mutate(election_date_lagged = as_date(ifelse(is.na(election_year)==F, election_mdy,                                                         ifelse(is.na(election_year)==T & is.na(lag(election_mdy))==F, lag(election_mdy), NA))))

## Now filter by date of protest

# Steps:
  #1) Remove rows where election_upcoming and election_lagged are both NA. 
  #2) This results in a dataframe with only election years +-1 year; create interval variables (lagged and leading days between protest and election)
  #3 Filter to cases that are six months before to 3 months after election (i.e 180 days and 90 days)

nelda3.sub <- nelda3 %>% filter(is.na(election_date_upcoming)==F | is.na(election_date_lagged) == F)

nelda3.sub <- nelda3.sub %>% mutate(protest_preelex_interval = ifelse((election_mdy - protest_mdy) >0, (election_mdy-protest_mdy), NA))

nelda3.sub <- nelda3.sub %>% mutate(protest_postelex_interval = ifelse((election_mdy - protest_mdy) <= 0, (protest_mdy - election_mdy), NA))

nelda3.sub <- nelda3.sub %>% group_by(ccode) %>% filter(protest_preelex_interval <= 180 | protest_postelex_interval <= 90)

# Next step is to aggregate needed variables (participants) by electionid
```

Aggregating participant variables
```{r}

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(participants_category == "", NA, 
                                            ifelse(participants_category == "50-99", 1, 
                                            ifelse(participants_category == "100-999", 2,
                                            ifelse(participants_category == "1000-1999", 3,
                                            ifelse(participants_category == "2000-4999", 4,
                                            ifelse(participants_category == "5000-10000", 5,
                                                   6)))))))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100s", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000s", 4, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50", 1, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50", 1, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000", 3, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "50+", 1, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000", 5, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "2000", 4, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "300", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "3000", 4, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100+", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000", 5, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000s", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds", 2, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "15000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1500", 3, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 50", 1, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "4000", 4, participants_category_ordered))  ## Stopping here; all other categories have fewer than 20 observations. Beginning second stopping rule:

## Second rule: collect all observations greater than 10,000

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "hundreds of thousands", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "<20000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 40000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 42000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">35000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">40000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">50000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "110000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14059", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "16000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "18000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "24000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "250000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "30000+", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "35000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "36000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "5000-50000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "55000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "650000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "> 10000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "1000000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "13000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "500000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">15000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">20000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "14000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "400000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "Tens of thousands", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">100000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "150000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "20000+", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "60000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "90000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "10000+", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "100000s", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "200000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== ">10000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "70000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "80000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "40000", 6, participants_category_ordered))

nelda3.sub <- nelda3.sub %>% mutate(participants_category_ordered = ifelse(is.na(participants_category_ordered) == T & participants== "25000", 6, participants_category_ordered))

#nelda3.sub <- nelda3.sub %>% group_by(electionid) %>% 
```

Remove post-election protests unrelated to election

```{r}
nelda3.sub <- nelda3.sub %>% filter(protest_mdy < election_mdy | (protest_mdy >= election_mdy & str_detect(notes, regex(pattern = "election*|vot*|ballot*", ignore_case = T))))
```

Getting grouped summarize for number and size
```{r}
nelda3.sub.pre <- nelda3.sub %>% filter(protest_mdy < election_mdy)
nelda3.sub.post <- nelda3.sub %>% filter(protest_mdy >= election_mdy)

count_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(n.events.massmob.pre = sum(protestnumber))
count_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(n.events.massmob.post = sum(protestnumber))

size_data_mm_pre <- nelda3.sub.pre %>% group_by(electionid) %>% summarize(median.participants.pre = median(participants_category_ordered, na.rm = T))
size_data_mm_post <- nelda3.sub.post %>% group_by(electionid) %>% summarize(median.participants.post = median(participants_category_ordered, na.rm = T))

```

##Nelda2 looks at NELDA variables as DV

```{r}
rm(vdem.data)
rm(vdem.recent)
rm(lji)
rm(nelda)
rm(massmob)
```

```{r}
nelda2.sub <- nelda2 %>% filter(v2x_regime == 1 | v2x_regime == 2) #Note that coding scheme for electoral democracies allows for "some degree of fraud and irregularities but these did not in the end affect the outcome" (v2elfrfair, which is a component of v2x_regime) 
```

```{r}
ecav.pre <- ecav %>% filter(Date < Electiondate)
ecav.pre <- ecav.pre %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests

ecav <- ecav %>% filter(Date >= Electiondate)
ecav <- ecav %>% filter(str_detect(EventName, "rotest") == T) #Only captures protests
```

Grouped summaries
```{r}
count_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(n.events.pre = n())
count_data_post <- ecav %>% group_by(NeldaID) %>% summarize(n.events.post = n())

##Duration
max_date <- ecav %>% group_by(NeldaID) %>% summarize(latest.event = max(Date))
min_date <- ecav %>% group_by(NeldaID) %>% summarize(earliest.event = min(Date))
duration_data <- left_join(min_date, max_date, by = "NeldaID")
duration_data <- duration_data %>% mutate(d.events = as.duration(interval(earliest.event, latest.event)))
duration_data <- duration_data %>% mutate(d.events.num = as.numeric(d.events, "days"))
duration_data <- duration_data %>% mutate(d.events.num = ifelse(d.events.num == 0, 1, d.events.num)) #1-day events

##Size
ecav <- ecav %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
size_data <- ecav %>% group_by(NeldaID) %>% summarize(median.participants = median(ParticipantNumber, na.rm = T))

ecav.pre <- ecav.pre %>% mutate(ParticipantNumber = ifelse(ParticipantNumber == -99, NA, ParticipantNumber))
size_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(median.participants.pre = median(ParticipantNumber, na.rm = T))

## Violence
violence_data <- ecav %>% group_by(NeldaID) %>% summarize(antigov.violence = sum(antigov.violence, na.rm = T))
violence_data_pre <- ecav.pre %>% group_by(NeldaID) %>% summarize(antigov.violence.pre = sum(antigov.violence, na.rm = T))

##Combined
ecav_data <- left_join(count_data_post, duration_data, by = "NeldaID")
ecav_data <- left_join(ecav_data, size_data, by = "NeldaID")
ecav_data <- left_join(ecav_data, violence_data, by = "NeldaID")

ecav_data_pre <- left_join(count_data_pre, size_data_pre, by = "NeldaID")
ecav_data_pre <- left_join(ecav_data_pre, violence_data_pre, by = "NeldaID")

ecav_data <- left_join(ecav_data, ecav_data_pre, by = "NeldaID")

## Merge back with NELDA 2
```

Merge ECAV with NELDA

```{r}
nelda2.sub <- nelda2.sub %>% rename(NeldaID = electionid)
nelda2.sub <- left_join(nelda2.sub, ecav_data, by = "NeldaID")

##Cleaning

nelda2.sub <- nelda2.sub %>% mutate(antigov.violence = ifelse(is.na(antigov.violence)==T, 0, antigov.violence))
nelda2.sub <- nelda2.sub %>% mutate(n.events.post = ifelse(is.na(n.events.post)==T, 0, n.events.post))
nelda2.sub <- nelda2.sub %>% mutate(d.events.num = ifelse(is.na(d.events.num)==T, 0, d.events.num))
#nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(is.na(median.participants)==T, 0, median.participants))

nelda2.sub <- nelda2.sub %>% mutate(any.ecav.protest = ifelse(n.events.post >= 1, 1, 0))

##Making the assumption that NAs for pre-election protest are zero
nelda2.sub <- nelda2.sub %>% mutate(n.events.pre = ifelse(is.na(n.events.pre) == T, 0, n.events.pre))
nelda2.sub <- nelda2.sub %>% mutate(antigov.violence.pre = ifelse(is.na(antigov.violence.pre) == T, 0, antigov.violence.pre))

```

```{r}
#nelda3 <- nelda3 %>% mutate(protests.postelex = ifelse(protests == 1 & protesterdemand1 == "political behavior, process", 1, 0))

#nelda3 <- nelda3 %>% mutate(protests.postelex.quant = ifelse(protests.postelex == 1, protestnumber, 0))
```

## World Bank urbanization data

```{r}
urban <- read.csv(here::here("Data", "urbanization_worldbank.csv"))
test <- urban %>%
  pivot_longer(cols = starts_with("X"),
               names_to = "year")
test <- test %>% mutate(year = gsub("X", "", year))

#Next is line up country codes in test and nelda 2, then merge

library(countrycode)

test <- test %>% mutate(stateid = countrycode(sourcevar = Country.Code, origin = 'iso3c', destination = 'gwc'))
test <- test %>% rename(urban.pct = value)
test <- test %>% dplyr::select(year, urban.pct, stateid)

nelda2.sub <- nelda2.sub %>% mutate(year = as.character(year))

nelda2.sub <- nelda2.sub %>% left_join(test, by = c("stateid", "year"))

```

## Merge with winner share updated

```{r}
winner.shares <- read.csv(here::here("Data", "amended_vote_shares.csv"))
nelda2.sub <- nelda2.sub %>% left_join(winner.shares, by = "NeldaID")
```

## Merge data from nelda3.sub with nelda2.sub

```{r}
##Combined post data together, then pre data, then both
mm_data_post_comb <- left_join(count_data_mm_post, size_data_mm_post, by = "electionid")
mm_data_pre_comb <- left_join(count_data_mm_pre, size_data_mm_pre, by = "electionid")

mm_data_comb <- left_join(mm_data_pre_comb, mm_data_post_comb, by = "electionid")

mm_data_comb <- mm_data_comb %>% rename(NeldaID = electionid)

nelda2.sub <- nelda2.sub %>% left_join(mm_data_comb, by = "NeldaID")  ## Double check to make sure this joining works

```

## Writing

```{r}
write.csv(nelda2.sub, here::here("Data", "nelda2_sub.csv"))
```

```{r}
write.csv(nelda3.sub, here::here("Data", "nelda3_sub.csv"))

```