### 2022-11-17 Output/appendix2.Rmd
Working to build a v3 draft that highlights MM data

---
title: "R Notebook"
output: html_notebook
---

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(cowplot)
library(pscl)
library(modelsummary)
library(ggeffects)
library(MASS)

nelda3.sub <- read.csv(here::here("Data", "nelda3_sub.csv"))

nelda3.sub <- nelda3.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda3.sub <- nelda3.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda3.sub <- nelda3.sub %>% rename(boycott = nelda14)
nelda3.sub <- nelda3.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda3.sub <- nelda3.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda3.sub <- nelda3.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda3.sub <- nelda3.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda3.sub <- nelda3.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

##Removing cases before 1990, when ECAV starts
nelda3.sub <- nelda3.sub %>% filter(year >= 1990)

## Removing cases where inc. loses
nelda3.sub <- nelda3.sub %>% filter(inc.lose == 0)

##Intmon as a factor
nelda3.sub <- nelda3.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda3.sub <- nelda3.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda3.sub <- nelda3.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

#Crisis as factor

nelda3.sub <- nelda3.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Margins with updates

nelda3.sub <- nelda3.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

#Successful protest = new elections or new incumbent

nelda3.sub <- nelda3.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

#Protest cases only

protest.cases <- nelda3.sub %>% filter(any.ecav.protest == 1)

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda3.sub <- nelda3.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Changing participants from numeric category to text
nelda3.sub <- nelda3.sub %>% mutate(median.participants = round(median.participants))
nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = ifelse(is.na(median.participants)==T, NA, ifelse(median.participants == 1, "Less than 10",
                     ifelse(2, "10 to 99", ifelse(3, "100 to 999",
                                                  ifelse(4, "1000 to 9,999", ifelse(5, "Greater than 10,000", NA)))))))
nelda3.sub <- nelda3.sub %>% mutate(particpants_categorical = as.factor(particpants_categorical))

## Median particiants (MM) as factor

nelda3.sub <- nelda3.sub %>% mutate(median.participants.mm.post = as.factor(median.participants.mm.post))

## Inverting the vote buying measure so it is not reverse coded

nelda3.sub <- nelda3.sub %>% mutate(v2elvotbuy_inv = -1*v2elvotbuy)

```

```{r}
incl.vars <- nelda3.sub %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp, any.ecav.protest, median.participants.mm.pre, median.participants.mm.post)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))

datasummary_skim(incl.vars)
```

## Protest size (categories) using mass mobilization data

```{r}
model.size.base <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.size.base)

model.size.base2 <- polr(median.participants.mm.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.size.base2)

model.size.griev <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + harass_opp +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.size.griev)

model.size.griev2 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + + harass_opp + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub)
summary(model.size.griev2)

model.size.griev3 <- polr(as.factor(median.participants.mm.post) ~ jucon.1lag + legcon.1lag + altinf.1lag + harass_opp +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + median.participants.mm.pre + v2elvotbuy_inv, data = nelda3.sub)
summary(model.size.griev3)
```

Plotting above models

```{r}
##Winning margin
dat <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-.47, 1.77]"))
dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "1" | response.level == "2" | response.level == "3")
plot(dat.sub)

dat.sub <- dat %>% dplyr::filter(response.level == "4" | response.level == "5" | response.level == "6")
plot(dat.sub)

dat_max <- ggpredict(model.size.base2, terms = c("winner.share.upd", "v2elirreg.inv [-2.3, 3.2]"))
dat.sub <- dat_max %>% dplyr::filter(response.level == "2" | response.level == "4" | response.level == "6")
plot(dat.sub) + coord_cartesian(ylim = c(0, .25))
```

```{r}
## Econ grievance
dat <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv"))
dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "2" | response.level == "4" | response.level == "6")
plot(dat.sub)

dat_max <- ggpredict(model.size.griev2, terms = c("econ.crisis.nelda.1lag", "v2elirreg.inv [-2.3, 3.2]"))
dat.sub <- dat %>% dplyr::filter(response.level == "0" | response.level == "3" | response.level == "6")
plot(dat.sub)
```

## Hurdle models with MM quantity

```{r models-hurdle-mm, warning = FALSE, message = FALSE, include = F}
model.n.base.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag + gdpgro.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base.hurd.mm)

model.n.base2.hurd.mm <- hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre  + v2elvotbuy_inv|  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + pubseccorrup.lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub %>% filter(gdpgro.1lag < 1), dist = "negbin")
summary(model.n.base2.hurd.mm)

model.n.griev.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv| jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, dist = "negbin", data = nelda3.sub %>% filter(gdpgro.1lag < 1))
summary(model.n.griev.hurd.mm)

model.n.griev2.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag  + altinf.1lag + osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   econ.crisis.nelda.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub, dist = "negbin")
summary(model.n.griev2.hurd.mm)

model.n.griev3.hurd.mm <- hurdle(n.events.massmob.post ~ jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv | jucon.1lag + legcon.1lag + altinf.1lag +  pubseccorrup.lag + osorg.1lag +  natlpartyorg.lag + v2elirreg.inv + pubseccorrup.lag +   pubseccorrup.lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + v2elvotbuy_inv, data = nelda3.sub, dist = "negbin")
summary(model.n.griev3.hurd.mm)
```

## Plots for hurdles

```{r}
dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.78]")) 
p.mm.margin <- plot(dat)
p.mm.margin + coord_cartesian(ylim = c(0, 100))
```

```{r custom-tidiers, include=FALSE}
tidy.hurdle.count <- function(x, conf.int = FALSE, conf.level = 0.95, ...) {

  result <- summary(x)$coefficients$count %>%
    tibble::as_tibble(rownames = "term") #%>%
    #dplyr::rename(estimate = Estimate,
    #              std.error = `Std. Error`,
    #              statistic = `t value`,
     #             p.value = `Pr(>|t|)`)

  if (conf.int) {
    ci <- confint(x, level = conf.level)
    result <- dplyr::left_join(result, ci, by = "term")
  }

  result
}

#test <- summary(model.n.base.hurd)$coefficients$count

test <- tidy.hurdle.count(model.n.base.hurd)

#tbl_regression(test)
```

```{r zero-model-table, echo=FALSE}

models <- list()
models[['Model 1']] <- model.n.base.hurd
models[['Model 2']] <- model.n.base2.hurd
models[['Model 3']] <- model.n.griev.hurd
models[['Model 4']] <- model.n.griev2.hurd
models[['Model 5']] <- model.n.griev3.hurd

modelsummary::modelsummary(models, coef_omit = "count", gof_omit = "R2", stars = T)

```