### 2023-07-11 Output/appendix_v4.Rmd
Continuing efforts to address R comments

---
title: "Appendix: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
---

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
library(pscl)
library(survival)
library(modelsummary)
library(glmmTMB)
library(huxtable)
library(kableExtra)
library(simPH)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2023-05-30.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

##Removing cases before 1990, when ECAV starts
nelda2.sub <- nelda2.sub %>% filter(year >= 1990)

## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)

##Updated winner share
nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

#Crisis as factor

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Margins with updates

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(types == "Executive", v2elvotlrg_amend/100, v2ellovtlg_amend/100))

#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))

## Coding max_size as zero for elections with no protest
nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Loading massmob

massmob <- read.csv(here::here("Data", "nelda_massmob_all.csv"))
massmob <- massmob %>% rename("NeldaID" = "electionid")
massmob <- massmob %>% dplyr::select(NeldaID, n.events.massmob.pre, n.events.massmob.post)

nelda2.sub <- nelda2.sub %>% left_join(massmob, by = "NeldaID")

```

# Alternative protest data: Mass Mobilization dataset

```{r models-hurdle-mm, message = FALSE, echo=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(n.events.massmob.pre.scale = scale(n.events.massmob.pre))

model.n.base.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
                                         zero.dist = "binomial",
                                          dist= "negbin",
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.base.hurd.mm)

model.n.base2.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
                                         zero.dist = "binomial",
                                          dist= "negbin",
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.base2.hurd.mm)

model.n.griev.hurd.mm <- pscl::hurdle(n.events.massmob.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +     gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre +  unscheduled_election + regional_success_lag,
                                         zero.dist = "binomial",
                                          dist= "negbin",
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.griev.hurd.mm)

```

The Mass Mobilization dataset [@massmob] aims to collect data on every protest event with greater than 50 participants in over 160 countries, from 1990 to 2014. Since the 50-participant threshold undercounts small protests, I do not use it in the main text and I do not use it here to test hypotheses on protest size. To match the ECAV format, I limit the window for protests to 180 days prior to the election and 90 days after. Unlike with ECAV, the counted protests need not be explicitly tied to the election. Using Mass Mobilization data generates results similar to those in the main text, as shown in Tables \@ref(tab:table-mm-binary) and \@ref(tab:table-mm-count). Fraud severity has no standalone effect on protest initiation or protest count. Election fraud does have a significant relationship to protest initiation in Model 2, at low incumbent vote-shares, but as Figure \@ref(fig:plot-margin-mm) indicates, this relationship is not apparent at typical levels of election fraud. As in the main text, the main drivers of protest count appear to be institutional and structural\textemdash repressive capacity and opposition organization.

```{r table-mm-binary, echo=FALSE, warning=FALSE, message=FALSE}
models_a1 <- list()
models_a1[['Model 1']] <- model.n.base.hurd.mm
models_a1[['Model 2']] <- model.n.base2.hurd.mm
models_a1[['Model 3']] <- model.n.griev.hurd.mm

modelsummary::modelsummary(models_a1, stars = T, longtable=T, coef_omit = "count", title = 'Binary portion of hurdle models of post-election protest', coef_rename = 
               c("zero_(Intercept)" = "Intercept", "zero_jucon.1lag" = "Judical independence (lag)", "zero_legcon.1lag" = "Leg. constraints (lag)", 
                 "zero_gdpgro.1lag" = "GDP growth rate (lag)", "zero_altinf.1lag" = "Alternative info. (lag)", 
                 "zero_osorg.1lag" = "Civil soc. openness (lag)", "zero_econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "zero_natlpartyorg.lag" = "National party organization (lag)", "zero_presidential" = "Presidential election", "zero_libdem.1lag" = "Liberal democracy index",
                 "zero_inc.termlimit.num" = "Incumbent term-limited",  "zero_regional_success_lag" = "Regional diffusion",
                 "zero_urban.pct" = "Urbanization", "zero_rentierstate" = "Rentier state", "zero_taxstate" = "Tax state",
                 "zero_n.events.massmob.pre" = "Number of pre-election protests", "zero_v2elirreg.inv" = "Election fraud",
                 "zero_physinteg.1lag" = "Physical integrity (lag)", "zero_physintlag.sq" = "Physical integrity (lag), squared", "zero_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "zero_unscheduled_election" = "Off-schedule election"))
```

\newpage

```{r table-mm-count, echo=FALSE, warning=FALSE, message=FALSE}
modelsummary::modelsummary(models_a1, stars = T, longtable=T, coef_omit = "zero", title = 'Count portion of hurdle models of post-election protest', coef_rename = 
               c("count_(Intercept)" = "Intercept", "count_jucon.1lag" = "Judical independence (lag)", "count_legcon.1lag" = "Leg. constraints (lag)", 
                 "count_gdpgro.1lag" = "GDP growth rate (lag)", "count_altinf.1lag" = "Alternative info. (lag)", 
                 "count_osorg.1lag" = "Civil soc. openness (lag)", "count_econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "count_natlpartyorg.lag" = "National party organization (lag)", "count_presidential" = "Presidential election", "count_libdem.1lag" = "Liberal democracy index",
                 "count_inc.termlimit.num" = "Incumbent term-limited",  "count_regional_success_lag" = "Regional diffusion",
                 "count_urban.pct" = "Urbanization", "count_rentierstate" = "Rentier state", "count_taxstate" = "Tax state",
                 "count_n.events.massmob.pre" = "Number of pre-election protests", "count_v2elirreg.inv" = "Election fraud",
                 "count_physinteg.1lag" = "Physical integrity (lag)", "count_physintlag.sq" = "Physical integrity (lag), squared", "count_winner.share.upd" = "Incumbent vote-share", "v2elirreg.inv" = "Election fraud", "count_unscheduled_election" = "Off-schedule election"))
```

```{r plot-margin-mm, echo=FALSE, message=FALSE, warning=F, fig.cap="Fraud, winning vote-share, and probability of protest initiation (Mass Mobilization data)", out.width = '85%', fig.pos="h!"}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(is.na(n.events.massmob.post)==T, NA, ifelse(n.events.massmob.post > 0, 1, 0)))

model.margin.binary_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.massmob.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.margin.binary_n)

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary_n, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.64, 1.55]"))
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

## Count portion

dat <- ggpredict(model.n.base2.hurd.mm, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.64, 1.55]"), type = "count") 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw()

plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_typ_n, p.n.margin, ncol=1)
plot.combined.prob

```

```{r, include=FALSE}
model.binary.griev1_n <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2elirreg.inv +  +   gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.binary.griev1_n)

plot_binary_gdp_typ_n <- sjPlot::plot_model(model.binary.griev1_n, type = "pred", terms = c("gdpgro.1lag [all]", "v2elirreg.inv [-.45, 1.78]"))
plot_binary_gdp_typ_n <- plot_binary_gdp_typ_n +  labs(x = "GDP growth (lagged)", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3))

## Count portion

dat <- ggpredict(model.n.griev.hurd.mm, terms = c("gdpgro.1lag", "v2elirreg.inv [-.64, 1.55]"), type = "count") 
p.n.gdp <- plot(dat)
p.n.gdp <- p.n.gdp +  labs(x = "GDP growth rate (lag)", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(xlim = c(-.4, 3), ylim = c(0, 200))
```

# Repression and pre-election protest

The main analysis already controls for possible confounding by the state's repressive capacity; i.e. the possibility that in highly repressive states, protest is deterred and fraud can proceed unopposed. However, it may still be useful to check an alternative explanation\textemdash perhaps protest is deterred in highly repressive states with considerable pre-election unrest (that is, when protest appears to be exceeding the ability of the state to contain). To check this possibility, a variation of Model 1 from the main text is run, but with an interaction between the number of pre-election protests and the lagged *physical integrity* variable. If protest risk deters manipulation in this way, there should be a negative and significant effect for the number of pre-election protests in highly repressive states.

Figure \@ref(fig:prot-rep-fraud) shows the absence of such an effect. In fact, there is no significant difference in the predicted level of fraud at lower levels of physical integrity, even when pre-election protest is widespread. A significant effect is seen at the high end of protection of physical integrity rights, but it is present only in extreme cases\textemdash the value of pre-election protests chosen here, 25, is above the 98th percentile for the variable\textemdash and is in the opposite direction predicted by this counterargument. That is, more numerous pre-election protest is associated with higher predicted levels of fraud in less repressive states than when the pre-election period is quiet. This robustness check shows that protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to attempt to claim a large margin of victory or to prevent elite or bureaucratic splits in a challenging period.

```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity", fig.pos="h!"}
m1_rep <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre*physinteg.1lag  +  unscheduled_election  + (1 | stateid), data = nelda2.sub %>% filter(v2x_regime <= 2)) 
#summary(m1_rep)
p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("physinteg.1lag", "n.events.pre [0, 25]"))
p.m1rep + labs(title = "Predicted values of election fraud", x = "Physical integrity (lag)", y = "Election fraud", color = "N. of pre-election protests") + theme_bw()
```

# Prior manipulation and expecations

Another explanation for post-election protest could hinge on changes over time; elections that are much more fraudulent than prior elections may exacerbate citizens' grievances or convince opposition groups that the incumbent is unexpectedly weak. This possibility is tested in two ways. First, the *election fraud* variable is replaced with *fraud change*, the difference between the current election and the one immediately prior. Since the prior election's value is subtracted from the 'current' one, positive values mean that the current election was more fraudulent than its predecessor, while negative values indicate election fraud declined. Using this conceptualization of election fraud does result in some support for the deterrence thesis: an election that is markedly more fraudulent than the prior election does appear associated with more numerous protest events at higher incumbent vote-shares. However, as in other examples, this effect only becomes statistically significant at the extremes. The values for change in fraud used in the marginal effects plots in Figures \@ref(fig:n-protests-ch) and \@ref(fig:size-protests-ch) represent very unusual changes, roughly between the 5th and 95th percentiles. In other words, the expected number and duration of protest events are only statistically significantly different when comparing elections that are far cleaner than the previous election and those that are far worse\textemdash a set of options that is unlikely to confront most incumbents.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(fraud.change = v2elirreg.inv - v2elirreg.inv.lag) #Positive values mean worsening fraud

model.n.base2.lag <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
#summary(model.n.base2.lag)

#model.d.base2.lag <- glmmTMB::glmmTMB(d.events.num ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + fraud.change*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag ,
#                                          zi = ~.,
#                                          family= nbinom2,
#                                          data = nelda2.sub %>% filter(v2x_regime <= 2))
#summary(model.d.base2.lag)

model.size.base2.lag <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  fraud.change*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
                                family=truncated_nbinom2)
#summary(model.size.base2.lag)

```

```{r table-lagged-manip, results='asis', echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 4']] <- model.n.base2.lag
#models3[['Model 5']] <- model.d.base2.lag
models3[['Model 5']] <- model.size.base2.lag

t4a <- huxreg(models3, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Fraud change" = "fraud.change", "Incumbent vote-share" = "winner.share.upd", "Fraud change:Incumbent vote-share" = "fraud.change:winner.share.upd"),
              statistics = character(0)) %>% set_caption("Models of number, duration, and size of protests") %>% insert_row("", "Number", "Size", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 10)
theme_compact(t4a, header_rows = TRUE, header_cols = FALSE)  %>% set_bottom_border(10, everywhere)

```

```{r n-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and number of protest events", echo=FALSE, message=FALSE, fig.pos="h!"}
dat <- ggpredict(model.n.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.12, 1.12]"), type = "count") 
p.n.margin.ch <- plot(dat)
p.n.margin.ch <- p.n.margin.ch +  labs(x = "Incumbent vote-share", y = "Number of events", title = "", color = "Change in fraud") + theme_bw()
p.n.margin.ch
```

```{r size-protests-ch, fig.cap="Change in fraud, incumbent vote-share, and size of protest events", echo=FALSE, message=FALSE, fig.pos="h!"}
dat <- ggpredict(model.size.base2.lag, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "fraud.change [-1.22, 1.22]"), type = "count") 
p.s.margin.ch <- plot(dat)
p.s.margin.ch <- p.s.margin.ch +  labs(x = "Incumbent vote-share", y = "Size of protest wave", title = "", color = "Change in fraud") + theme_bw() + coord_cartesian(ylim = c(0, 70000))
p.s.margin.ch 
```

Second, the NELDA dataset includes a binary variable noting whether "there was evidence of domestic or international concern that the election process was not going to be free or fair" in advance of the election (nelda11). While this is likely to be rough indicator, it still provides an opportunity to test the prediction that election fraud may be more likely trigger protest when it exceeds a baseline expectation. In particular, when fraud is high and there are no pre-election expectations of severe manipulation, this view might predict a higher likelihood and severity of protest. This is tested by interacting *intentional voting irregularities* with the binary variable from NELDA, called *pre-election expectations* here.

```{r expectations-table, echo=FALSE, message=FALSE, warning=FALSE}
model.n.concerns.hurd <- glmmTMB::glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    preelection_concerns*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.concerns.hurd)

model.size.concerns.hurd <- glmmTMB::glmmTMB(max_size ~  jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    preelection_concerns*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.size.concerns.hurd)

models8 <- list()
models8[['Model 6']] <- model.n.concerns.hurd
models8[['Model 7']] <- model.size.concerns.hurd

t8a <- huxreg(models8, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Pre-election concerns" = "preelection_concerns", "Election fraud : Pre-election concerns" = "v2elirreg.inv:preelection_concerns"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0)
theme_compact(t8a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
```

As Table \@ref(tab:expectations-table) and Figures \@ref(fig:n-protests-concerns) and \@ref(fig:protests-size-concerns) indicate, there is no apparent effect for high levels of fraud when there are no significant pre-election concerns about the freeness and fairness of the election (as coded by NELDA). Taken together, these results suggest that the alternative, 'exceeding expectations' framework for the protest model of election manipulation is not supported.

```{r, echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.concerns.hurd, terms = c("v2elirreg.inv", "preelection_concerns"), type = "count") 
p.n.concerns <- plot(dat)
p.n.concerns <- p.n.concerns + labs(x = "Intentional irregularities", y = "Predicted number of protests" , title = "Predicted count of protest events", color = "Pre-election concerns")  + theme_bw()

dat <- ggpredict(model.size.concerns.hurd, terms = c("v2elirreg.inv", "preelection_concerns")) 
p.size.concerns <- plot(dat)
p.size.concerns <- p.size.concerns +  labs(x = "Intentional irregularities", y = "Predicted size of protests" , title = "Predicted max. size of protest events", color = "Pre-election concerns")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))

```

```{r n-protests-concerns, fig.cap="Election fraud, pre-election expectations, and number of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
p.n.concerns

```

```{r protests-size-concerns, fig.cap="Election fraud, pre-election expectations, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
p.size.concerns

```

# Chosen successors

The models in the main text control for the presence of term-limited executive incumbents, based on the logic in Hale [@hale_patronal_2014]. However, it could be the case that the nature of a term-limited election affects the behavior of protesters and election manipulators alike; that is, perhaps when there is a designated successor, the heir apparent is better able to mobilize the resources of the state to manipulate elections and suppress protest. Using a variable from NELDA that codes for designated successors, the following models add a categorical variable for the incumbent's status, which can take on the following values: *executive eligible*, *chosen successor*, *contested executive election*, and *legislative election*. As Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text. Higher levels of fraud does make the occurrence of protest more likely at low incumbent vote-shares, but this does not translate into a higher risk of more numerous or larger protest events.

```{r successor-table, echo=FALSE, message=FALSE, warning=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Executive eligible"))
nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.n.base2.succ <- glmmTMB::glmmTMB(n.events.post ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                         zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.base2.succ)

model.n.base2.succ.bin <- glm(no_protest ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                              data = nelda2.sub  %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.n.base2.succ.bin)

plot_binary_margin_succ_n <- sjPlot::plot_model(model.n.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
plot_binary_margin_succ_n <- plot_binary_margin_succ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()

####

model.d.base2.succ <- glmmTMB::glmmTMB(d.events.num ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.d.base2.succ)

model.size.base2 <- glmmTMB::glmmTMB(max_size ~ incumb.status + jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  v2elirreg.inv*winner.share.upd  +    physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0" & v2x_regime <= 2),
                                family=truncated_nbinom2)
#summary(model.size.base2)

model.size.base2.succ.bin <- glm(no_size ~ incumb.status + jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  +   physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag,
                              data = nelda2.sub  %>% filter(inc.lose == "0" & v2x_regime <= 2))
#summary(model.size.base2.succ.bin)

plot_binary_margin_succ_size <- sjPlot::plot_model(model.size.base2.succ.bin, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-.45, 1.78]"))
plot_binary_margin_succ_size <- plot_binary_margin_succ_size +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of no protest", color = "Election fraud") + theme_bw()

### Table

models4 <- list()
models4[['Model 8']] <- model.n.base2.succ
#models4[['Model 6']] <- model.d.base2.hurd
models4[['Model 9']] <- model.size.base2

t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Number", "Size", after = 0)
#theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)

```

```{r n-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
dat <- ggpredict(model.n.base2.succ, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Number of events", title = "Predicted number of protests", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 40))
```

```{r size-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
dat <- ggpredict(model.size.base2, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), type = "count") 
p.s.margin <- plot(dat)
p.s.margin <- p.s.margin +  labs(x = "Incumbent vote-share", y = "Size of events", title = "Predicted size of protests", color = "Election fraud") + theme_bw() + coord_cartesian(ylim = c(0, 18000))
```

```{r successor-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
plot.combined.prob <- cowplot::plot_grid(plot_binary_margin_succ_n, p.n.margin, p.s.margin, ncol=1)
plot.combined.prob
```

# Electoral authoritarian regimes only

The models in the main text make use of observations from country-years coded as electoral authoritarian and electoral democratic by the 'v2x_regime' variable in V-Dem. This approach is taken because it reflects the widespread availability of election-manipulation techniques across this range of regime types. But it may be that considering the two groups simultaneously obscures some within-group patterns. For example, it may be that electoral protest is more likely to be observed in electoral authoritarian regimes following the logic in Tucker [-@tucker2007], and given several prominent examples of electoral revolution in that category (e.g. Ukraine 2004, Georgia 2003). The models in Table \@ref(tab:auth-table) show a statistically significant coefficient for the interaction of election fraud and winner's vote-share, but Figures \@ref(fig:n-protests-margin-auth) and \@ref(fig:size-protests-margin-auth) show that this effect is not apparent for typical values of election fraud in that subset of the population (the first and third quartiles, respectively). As a result, there is no strong evidence to suggest that election fraud has different effects on protest dynamics in electoral authoritarian vs. electoral democratic regimes.

```{r, echo=FALSE}
model.n.base2.auth <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(v2x_regime == 1))
#summary(model.n.base2.auth)

model.size.base2.auth <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub %>% filter(v2x_regime == 1))
#summary(model.size.base2.auth)

```

```{r auth-table, echo=FALSE, warning=FALSE, message=FALSE}

models5a <- list()
models5a[['Model 10']] <- model.n.base2.auth
models5a[['Model 11']] <- model.size.base2.auth

t6b <- huxreg(models5a, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, only electoral authoritarian regimes") %>% insert_row("", "Number", "Size", after = 0)
theme_compact(t6b, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
```

```{r, echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.base2.auth, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]"), type = "count") 
p.n.margin.auth <- plot(dat)
p.n.margin.auth <- p.n.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw()

dat <- ggpredict(model.size.base2.auth, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [.35, 1.86]")) 
p.size.margin.auth <- plot(dat)
p.size.margin.auth <- p.size.margin.auth + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw()

```

```{r n-protests-margin-auth, fig.cap="Election integrity, incumbent vote-share, and number of protest events in authoritarian regimes", echo=FALSE, out.width = '70%', fig.align='center'}
p.n.margin.auth

```

```{r size-protests-margin-auth, fig.cap="Election integrity, incumbent vote-share, and maximum size of protest events in authoritarian regimes", echo=FALSE, out.width = '70%', fig.align='center'}
p.size.margin.auth

```

# Including liberal democracies

The main text excludes liberal democracies from the analysis, on the grounds that the prior literature on protest and fraud typically confines itself to regimes that hold multiparty elections, and the liberal-democratic cases exhibit little variation in their (low) levels of election fraud. But perhaps it is in liberal democracies where the deterrent logic operates most effectively, settling into an equilibrium of high protest risk and low fraud. Including liberal democracies in the analysis results in little difference from the findings in the main text, however. Election fraud does not appear associated with a higher number of protest events, or larger protest sizes.

```{r lib-table, echo=FALSE, warning=FALSE}

model.n.base2.lib <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag +   winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub)
#summary(model.n.base2.lib)

model.size.base2.lib <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                          zi = ~.,
                                          family= nbinom2,
                                          data = nelda2.sub)
#summary(model.size.base2.lib)

models5 <- list()
models5[['Model 10']] <- model.n.base2.lib
models5[['Model 11']] <- model.size.base2.lib

t6a <- huxreg(models5, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,  omit_coefs = c("sd__(Intercept)", "sd__Observation"), coefs = c("Intercept" = "(Intercept)", "Election fraud" = "v2elirreg.inv", "Winner vote-share" = "winner.share.upd", "Election fraud : Winner vote-share" = "v2elirreg.inv:winner.share.upd"), statistics = character(0)) %>% set_caption("Models of number and size of protests, inclu. liberal democracies") %>% insert_row("", "Number", "Size", after = 0)
theme_compact(t6a, header_rows = TRUE, header_cols = FALSE)  %>% insert_row("Controls", "Yes", "Yes", after = 10) %>% set_bottom_border(10, everywhere)
```

# Alternative measures of election manipulation

To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity from V-Dem. This variable is `v2xel_frefair,' which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election. This variable ranges from 0 to 1, with higher values indicating freer and fairer elections. The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the ECAV data on protest.

```{r manipdv-full, echo=FALSE}
m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag   +  presidential + libdem.1lag + urban.pct + rentierstate + taxstate + n.events.pre +  physinteg.1lag + physintlag.sq    + unscheduled_election   + (1 | stateid), data = nelda2.sub)

#summary(m1)

models <- list()
models[['Model 12']] <- m1

modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity', gof_omit = "AIC|BIC",
             coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", "presidential" = "Presidential", "n.events.pre" = "N. of pre-election protest",
                 "physinteg.1lag" = "Physical integrity (lag)",
                 "physintlag.sq" = "Phys. integ. (lag), squared", "unscheduled_election" = "Off-schedule election",
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests"))
```

The first model shows that the number of pre-election protests is not associated with improved overall election integrity. This suggests that pre-election unrest is not associated with broad-scale improvements in election integrity across the electoral environment.

```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F}
model.n.base2.elfr <- glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +  v2xel_frefair*winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)
summary(model.n.base2.elfr)

model.size.base2.elfr <- lmer(log(max_size) ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + econ.crisis.nelda.1lag + natlpartyorg.lag + v2xel_frefair + pubseccorrup.lag +   winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election +  as.factor(inc.lose) + (1 | stateid), data = nelda2.sub %>% filter(max_size > 0))
summary(model.size.base2.elfr)

model.n.griev.elfr <- glmmTMB::glmmTMB(n.events.post ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair +  v2xel_frefair*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)

model.size.base2.elfr <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag + v2xel_frefair + gdpgro.1lag  +   winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag ,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)

model.size.griev.elfr <- glmmTMB::glmmTMB(max_size ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  osorg.1lag +  natlpartyorg.lag + v2xel_frefair +  v2xel_frefair*gdpgro.1lag  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                                ziformula=~.,
                                data = nelda2.sub %>% filter(inc.lose == "0"),
                                family=truncated_nbinom2)

```

Moving to the hurdle models of post-election protest, we do not observe much of a relationship between overall election integrity and protest. There is no statistically significant relationship between election integrity and protest size or quantity in any of the models, either alone or interaction with winner vote-share and lagged GDP growth. Figure \@ref(fig:n-protests-margin-fulldv) shows the absence of any effect for the number of protest events, while Figure \@ref(fig:protests-size-fulldv) shows the same for protest size. The values for *overall election integrity* chosen for these marginal effects plots represent the first and third quartiles, respectively. As a whole, the results are in accordance with the theory proposed in the main paper. There is no consistent evidence that lower-quality elections are associated with either a higher risk of protest initiation, or with a larger number of post-election protests.

```{r zero-model-table-fulldv, echo=FALSE, results='asis', warning=FALSE}

models <- list()
#models[['Model XXa']] <- model.n.base2.elfr
models[['Model 13']] <- model.n.base2.elfr
models[['Model 14']] <- model.n.griev.elfr
models[['Model 15']] <- model.size.base2.elfr
models[['Model 16']] <- model.size.griev.elfr

t7 <-  huxreg(models, tidy_args = c("component" = "cond"), borders = 0, number_format = 2,
 statistics = character(0), coefs = c("Intercept" = "(Intercept)", "Electoral integrity" = "v2xel_frefair", "Incumbent vote-share" = "winner.share.upd", "GDP growth (lag)" = "gdpgro.1lag", "Electoral integrity:Incumbent vote-share" = "v2xel_frefair:winner.share.upd",  "Election integrity : GDP growth (lag)" = "gdpgro.1lag:v2xel_frefair")) %>% set_caption("Conditional hurlde models of protest number and size") %>% insert_row("", "Number", "", "Size", "", after = 0)
theme_compact(t7, header_rows = TRUE, header_cols = FALSE) %>% insert_row("Controls", "Yes", "Yes", "Yes", "Yes", after = 14) %>% set_bottom_border(14, everywhere)

```

\newpage

```{r, echo=FALSE, message=FALSE}
dat <- ggpredict(model.n.base2.elfr, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2xel_frefair [.197, .61]"), type = "count") 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin + labs(x = "Incumbent vote-share", y = "Predicted number of protests" , title = "Predicted count of protest events")  + theme_bw() + coord_cartesian(ylim=c(0, 75))

dat <- ggpredict(model.size.base2.elfr, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2xel_frefair [.197, .61]")) 
p.size.margin <- plot(dat)
p.size.margin <- p.size.margin + labs(x = "Incumbent vote-share", y = "Predicted size of protests" , title = "Predicted maximum of protest events")  + theme_bw() + coord_cartesian(ylim = c(0, 50000))

```

```{r n-protests-margin-fulldv, fig.cap="Election integrity, incumbent vote-share, and number of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
p.n.margin

```

```{r protests-size-fulldv, fig.cap="Election integrity, incumbent vote-share, and size of protest events", echo=FALSE, out.width = '70%', fig.align='center'}
p.size.margin

```

\newpage

# References