### 2024-01-31 Scripts/authoritarian_only.Rmd
Edits through data and methods

---
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
indent: true
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{caption}
  - \captionsetup[table]{labelformat=empty}
---

\doublespace

On August 9, 2020, Belarus held an election in which President Aleksander Lukashenko ran for a sixth term. The official results credited Lukashenko with over 80\% of the vote, a result which observers regarded as wildly inflated [@GolosPlatformPresents2020]. In response, hundreds of thousands of Belarusians took to the streets over the course of several months. Though the protests put significant strain on the regime, a harsh campaign of repression (and support from neighboring Russia) ultimately quelled the protests; Lukashenko was inaugurated into a new five-year term.

The Belarusian case is typical in some ways, and an outlier in others. Like many previous elections in Belarus, and in authoritarian regimes around the world, the contest was heavily managed and the results inflated by fraud. However, the sustained, massive protest wave that occurred post-election is extremely unusual. However, the regime's ultimate victory over the protest movement\textemdash while costly and never guaranteed\textemdash is also the typical outcome of post-election protest.

These patterns highlight a puzzle generated by existing research on the relationship between election manipulation and popular protest. The topic has significant implications for authoritarian governance and democratization, and as such has been a very active subject of study. However, the various theoretical perspectives brought to bear on the topic have been built on different assumptions, and have different theoretical implications. To briefly summarize: for some models, election fraud is a tool of weak incumbents. As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@Fearon2011]. For others, fraud is committed by weak and strong incumbents alike, making protest a risky gamble the opposition undertakes for exogenous reasons. In these models, fraud may or may not be deterred based on other factors: a dubious, though not completely dismissive, prospect for democratization and self-enforcing democracy. Lastly, a third school of thought holds that election fraud is an indicator of genuine incumbent strength. This approach, which is expanded upon below, is a pessimistic account for democracy. If this model is correct, it is authoritarian politics rather than democracy that is self-enforcing\textemdash the ability to commit election fraud helps win elections, signal strength, deter challengers under most conditions, and put down protest movements if they do begin. This paper argues that the third model best fits the overall pattern of the data, and is supported by specific hypothesis tests.

In this paper, I argue that\textemdash in keeping with the idea of 'informational autocracy' (ADD CITES HERE)\textemdash an election outcome produces a signal of the relative strength of the incumbent. As Magaloni (2008) and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society. In some theories, this only works if the votes for the incumbent are largely 'genuine' (i.e. not fabricated), or produced using the incumbent's organizational resources (i.e. vote-buying and intimidation). (ADD CITES HERE, INCLUDING TO HARVEY AND M). These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski (2013) (ADD CITE): legitimation and cooptation. Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it.

However, these models overlook that engineering widespread fraud requires control over patronage resources and the criminal and judicial apparatus (ADD CITES to HARVEY). As a result, they overlook the fact that high levels of fraud are a sign of incumbent strength in both cooptation and Gerschewski's third pillar: repression. Both of these tools can be useful for defusing or dismantling an emergent protest movement. As a result, effective fraudsters should also be effective dismantlers of protest. This helps explain the broad pattern, where most fraudulent elections go unprotested, and among those that do generate protest, the vast majority fail to dislodge the incumbent. Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.

Testing this proposition in more depth is difficult. If, for example, in equilibrium incumbents manipulate just the right amount to avoid mass protest\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation. High and low levels of fraud would generally be associated with no protest. Beginning to untangle the relationship between election fraud and protest requires looking beyond standard binary measures of protest. Instead, this paper tests outcomes "off the equilibrium path," to use the language of formal modeling. Does greater risk to the incumbent in the pre-election period result in reduced election fraud? When protest movements do arise after an election, is fraud associated with weaker governments? Using data from 614 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that the answer to both questions is no. In fact, more severe election fraud is associated with a *faster* demise for protest waves. This indicates that, far from being a sign of weakness, in most cases the ability to generate significant electoral fraud is an indicator of durable regime strength. This finding holds even when incumbents claim small vote-shares, suggesting that electoral fraud capacity is an independent component of regime stability.

Combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019],... [MOVE]

Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has likely been overemphasized. Even in the unlikely event that protests do break out in response to large-scale fraud, the government has many tools it may deploy to curtail them. Fraud itself is a sign of the robustness of some of those tools\textemdash the control over resources that incentivizes fraud can be used to coopt the opposition or constrain elites, while control over the judicial and prosecutorial systems can be used to intensify repression during a crisis.

[REWRITE LATER] One troubling implication of these results is that ruling parties should follow an 'all-in' strategy on election fraud, a process that entails consolidating control over patronage resources and further eroding the rule of law. More fraud is not associated with longer-lasting protest waves, but a large margin of victory is associated with a lower risk of protest initiation. These results help explain both the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory. They also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation.

# Literature review: election manipulation and mass protest

Prior research has often considered that electoral manipulation has both costs and benefits for ruling parties. The benefits include better odds of winning the official tally [@lehoucq2002], of course, as well as any signals of strength and stability that a larger vote-share [@magaloni2006] or display of power that manipulation might provide [@simpser2013;@gehlbach2015electoral]. On the other side, election manipulation is generally characterized as having both operational costs (which the incumbent may be unwilling or unable to pay [@Greene2010]) and legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011]. The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.

The possibility of post-election protest regularly figures as a cost that incumbents must weigh when determining whether (and how much) to inflate their vote share. In these approaches, the cost of protest be conceptualized as the risk that the incumbent may be unseated and/or the resource cost needed to repress or buy off the incumbent's opponents. Various theories model the decision to engage in post-election protest either as a strategic one by opposition groups [@magaloni2010;@Luo2020;@przeworski2015], or more spontaneous coordination by citizens expressing a shared grievance [@tucker2007;@Brancati2016]. In either case, the risk of protest is a classic component of models of electoral manipulation, as a restraint on incumbents who would prefer to use fraud to increase their vote-shares, all else held equal. This argument can also be found in theories about the conduct of non-democratic elections, the focus here, as well as in the literature on self-enforcing democracy [@Fearon2011;@przeworskiSelfenforcingDemocracy2006]. In the latter literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results. While the focus here is on elections in authoritarian cases, the implications for self-enforcing democracy are discussed at the close of the paper.

<!-- Broadly, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework). The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs *may* manifest as protest, but may also manifest as a reduction in general cooperation with the regime. However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can *induce* cooperative behavior from some agents.], factor which affects the demand-side cost of manipulation in most models. -->

<!-- In the major formulations of this argument, such as Fearon  [-@fearon_self-enforcing_2011] and Przeworski [-@przeworskiSelfenforcingDemocracy2006], 'democracy' is understood in procedural terms. For Przeworski, democracy requires that "authorization to rule is derived from elections" which designate winners and losers according to a set of rules [@przeworskiSelfenforcingDemocracy2006, p. 312]. Fearon adopts a similar definition for electoral democracy [@fearon_self-enforcing_2011, p. 1662]. Such arrangements are not limited to liberal democracies, and in fact constitute the large majority of regimes today. Many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020]. In most such models, protests are successful in overthrowing the government with some probability, but impose a cost on both the protesters and the government regardless of success (e.g. [@Luo2020]); the logic thus extends even to highly repressive and entrenched regimes. Nevertheless, this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.  -->

Much of the work on this subject has been done in the formal modeling tradition. In game-theoretic research where the risk of protest is shown to deter fraud, the election is usually conceived as a signaling game. In such models, incumbents are characterized as 'weak' or 'strong' based on their support in the electorate; weak incumbents may need to resort to fraud in order to secure re-election. As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent. For example, the model developed by Little, Tucker, and LaGatta [-@Little2015] predicts that "Just as stronger election results for the incumbent lead to less protest because they indicate that she is popular, reports of more fraud lead to more protest because they indicate that the incumbent is, for a given election result, less popular" (p. 1153). In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud\textemdash a pro-democracy outcome. Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010]. These are normatively optimistic models: fraud can be kept in check, and ultimately reduced, through the development of election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].

A second class of formal models holds that fraud operates as a distortion of the signal of the incumbent's true strength. In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity. However, greater transparency makes such bluffs more costly. The typical result in these models is that protest risk can *partially* deter fraud. In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less. High levels of fraud can be committed by strong or weak incumbents, and as such these models do not predict a relationship between the risk of protest and the level of fraud per se. As Egorov and Sonin [-@egorov2014incumbency] write when discussing their model of this type, "We see that even dictators who are sure to win may use fraud; the reason is that if they do not, citizens will believe that they are less popular and used fraud anyway [thus inviting protest]." [ADD PAGE] In these models, protest initiation is most likely when fraud is higher than expected and when the incumbent's vote-share is low (see, e.g., Little [-@little_elections_2012].

Similar predictions emerge from another school of thought, which holds that electoral protest is driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016]. In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009]. Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991]. Just as above, these models tend to predict greater protest risk when margins of victory are low and/or grievances are high in the presence of fraud.

Empirical work has found some support for the microfoundations of this theory. For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021]. This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].

The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance. A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b]. Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013]. Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006]. These studies imply only a limited deterrent effect of protest on fraud in non-democracies. If the incumbent may pay the cost of protest whether or not it engages in fraud (due to subjective, and perhaps mistaken, beliefs about fraud and economic conditions), it may as well engage in manipulation in order to gain the benefits.

The strategic and social-psychological explanations are, of course, interconnected. As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something. Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304].

Finally, there is a third school of thought, which holds that election manipulation may indicate regime strength, especially if manipulation is blatant or produces large margins of victory [@simpser2013]. In the next section, I expand on this school of thought in more detail.

There is a general, though not complete, consensus in empirical work that election fraud is associated with increased risk of protest. Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019]. A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012]. However, Hafner-Burton et al [-@Hafner-Burton2016] find that the positive association between election fraud and post-election protest falls short of conventional statistical significance. Likewise, looking at the 2011 election and protest cycle in Russia, Ananyev and Poyker [-@ananyevDictatorsSignalStrength2022] show that post-election protest was not associated with the severity of fraud using several measures^[Lankina and Skovoroda [-@Lankina2017] find that post-election protest was positively associated with a different measure of election fraud in that case, however.] Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e. whether opponents are allowed to run), rather than by the election-day manipulation that usually figures in theoretical work. This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].

While these works have identified a connection between fraud and protest, in general, they cannot distinguish between the competing assumptions of the two theories discussed here, or the third possible approach discussed in below. All of the relevant theories predict that protest should be most likely when evidence of fraud is high and the incumbent's vote-share is low (or, relatedly, grievances are high). However, the implications of the theories are substantively distinct. Protest risk only acts as a deterrent to fraud in theories where fraud is a signal of regime weakness. So, it is necessary but insufficient to ask whether the occurrence of post-election protests is associated with higher fraud. It is also necessary to test the assumptioins underlying the different models. In particular, do post-election protest movements encounter weak incumbents as signaling models predict?  If not, the deterrence model is undermined.

<!--  One critique of this view, although still within the information/strategic framework, comes from Simpser's [-@Simpser2011] pathbreaking book. Simpser argues that when wanipulation efforts are excessive---well beyond what is needed simply to win---they demonstrate the organizational capacity, resources, and manpower of the ruling party [@Simpser2013]. Large-scale manipulation is, thus, a signal of strength rather than weakness, and can deter protest. By contrast, partial manipulation efforts that fail to secure a large majority vote-share signals the government lacks both popular support and mobilizational capacity, making opposition protest more likely. -->

[MOVE TO DISCUSSION AND REWRITE] Generally speaking, protest-based models predict that manipulation should decline as opposition groups become stronger relative to incumbents. This dynamic creates a sort of virtuous cycle, wherein strong opposition groups' credible threat of protest makes elections fairer, which in turn improves the strength of the opposition--leading ultimately to the ideal of self-enforcing democracy. The alternative school of thought, among other predictions, argues that opposition strength drives front-line agent behavior more than that of leaders, and causes a shift to less dispersed, less observable kinds of manipulation [@asunka2017electoral;@Harvey2019]. Election integrity can thus remain low, even in competitive contexts. Putting the protest-deterrent logic to the test, then, helps resolve significant questions in the study of election manipulation and democratization.

  Of course, electoral manipulation is not understood to be the sole cause of protest\textemdash protests occur for all sorts of reasons\textemdash but the literature has drawn a close connection between the risk of specifically electoral protest and the conduct of elections.

# Theory

As discussed above, one view of election fraud and protest views fraud as a tool of weak incumbents, while another views it as a tool that is employed by weak and strong alike. A third view, associated with Simpser [-@simpser2013], holds that election fraud demonstrates incumbent strength. In Simpser's original account, we should expect electoral manipulation to indicate strength when it is blatant\textemdash i.e. readily observable\textemdash or excessive, producing large margins of victory. A strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver]. This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are weakly associated with more election fraud.

[CONSIDER Ananyev and Poyker, Trinh 2022--do they fit in?]

I build on Simpser's logic, and subsequent research, by arguing that that extensive election fraud itself requires the use of resources that are sources of incumbent strength, whether or not these are observed. Simpser's model is concerned with whether or or not other actors perceive the regime's signal of strength, leading to the focus on excessive or blatant manipulation. I argue here that even when large-scale election fraud is marginal (i.e. produces narrow margins of victory), it can indicate regime strength independent of popularity. This argument  draws on the recent strand of research on election manipulation that emphasizes the principal-agent problems inherent in such campaigns. Political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].

control over the bulk of patronage resources in society , and the ability to shield agents from criminal and legal risks [@Harvey2019;]. Each of these resources can be used to minimize protest waves if they do occur after the election.

To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraud. Large-scale election fraud requires that the incumbent have consolidated control over patronage resources in society; this makes cooperation with the incumbent attractive for front-line agents, who are less willing to risk losing access to the incumbent's network [@Harvey2019]. However, such efforts also require that incumbents be able to shield front-line agents from negative consequences if their illegal efforts are publicly exposed [@Harvey2022]. In other words, the incumbent must have sufficient control over the police and the courts, such that ordinary individuals expect little punishment for committing criminal election fraud on their behalf. Each of these capacities is required to commit large-scale fraud, and both are useful for disarming protest movements.

The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014]. Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019]. In either case, large-scale election fraud indicates control over considerable patronage resources, which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).

Ruling parties may also lean on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007]. Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022]. The ability to protect regime agents from any legal consequences for carrying out orders, a requirement for successful election fraud, is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].

This argument\textemdash that successful, large-scale fraud draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head. Instead of signalling weakness, large-scale fraud indicates a well of resilience. Consequently, we should not expect incumbent regimes capable of committing mass fraud to be deterred from doing so by the risk of protest. This argument clearly echoes the seminal argument of Alberto Simpser [@simpser2013]. It is distinct, however; for Simpser, election manipulation only communicates strength to other actors when it is excessive (generating high margins of victory) or blatant. As Simpser notes, "...the marginal versus excessive dimension of electoral manipulation does not fully overlap with the question of scale or extent: while excessive electoral manipulation is generally associated with large-scale manipulation, the marginal kind can result either from a low amount of manipulation (e.g., in a tight race) or from a large amount [due to an unpopular incumbent]" (p. 7).

All together, I argue that a large margin of victory combined with election fraud sends an unambiguous signal of ruling-party strength, with uncertainty around its components. That is, large margins of victory can be composed of a number of values of 'natural' popularity and election fraud. As margins of victory decline, the relative value of the signal becomes clearer. A narrow margin of victory combined with extensive fraud is a clear signal of low 'natural' popularity. While this may indicate an opportune moment for opposition actors, the large extent of fraud is nevertheless itself an independent indicator of regime strength. The combination of patronage resources and control over the criminal and judicial systems required for mass fraud can be turned to coopting or repressing protest movements.

This theory makes predictions for protest initiation, protest duration, and the level of fraud employed. First, if large-scale election fraud indicates strength, we should expect that incumbents will respond to pre-election signs of opposition strength by increasing the level of fraud. This runs directly counter to the prediction of the classic models, which hold that a greater risk of post-election protest should deter fraud. The prediction of second-generation [USE THIS AS SHORTHAND?] models is embedded in the null hypothesis of H1, since these models hold that protest risk is not tethered to the level of fraud itself (e.g. [@egorov2014incumbency]). In those models, a greater risk of protest makes it more important to incumbents that they satisfy the opposition's expected level of fraud, which may be high or low.

> H1: Pre-election environments that are more uncertain for the incumbent will be associated with increased election fraud.

Likewise, if election fraud and incumbent vote-share are both indicator of incumbent strength, we should expect protest initiation to be most likely when both are low. A high vote-share in the absence of large-scale fraud indicates broader regime popularity, which dampens protest risk. As argued here, a high level of fraud can help compensate for a low vote-share by indicating independent sources of regime strength (i.e. resource capacity, agent loyalty, and biased rule of law). Only when both signals are low should we expect a high risk of protest. This hypothesis is consistent across all three schools of thought.

> H2: The likelihood of protest initiation will be highest when incumbent vote-share and election fraud are low.

Finally, if large-scale protest is indicative of regime resiliency, we should expect that election fraud will be associated with reduced protest duration among incumbents who successfully fended off protest movements. That is, among strong incumbents who were capable of defeating post-election protest movements, the resources associated with large-scale fraud should enable them to more swiftly quell the movement. Those who lack such resources are expected to take longer to dispatch the protest movement. The effect should run in the opposite direction for incumbents who were removed through post-election protest. These types, revealed ex post to be weak, should survive longer against the protest movement if they control the resources needed to carry out large-scale fraud. Those who lack such resources should be defeated more quickly. Again, these expectations are counter to the standard protest-deterrent logic and distinct from the null hypothesis expected in second-generation [USE?] models.

> H3a: Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.

>H3b: Among incumbents that are defeated by poste-election protest movements, more extensive election should be associated with longer protest waves.

That election dates are known in advance creates opportunities for opposition mobilization; anticipating this, regimes can more easily engage in pre-emptive repression [@Truex2019;@Sullivan2016;@Bhasin2013].

In addition, governments may learn from the experiences of other governments, just as opposition groups may learn from international opposition movements. Anticipating threats from below may encourage less democratic governments to invest more in the security services, or to reform political and legal institutions in order to stifle potential protests before they take shape [@hallAuthoritarianLearningConceptual2017a;@Morgenbesser2020], further reducing the expected cost of protest.

In particular,

# Data and methods

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(vdemdata)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
library(glmmTMB)
library(modelsummary)
library(huxtable)
library(simPH)
library(broom)
library(broom.helpers)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-01-30.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))

#Coding as factors

nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Replacing NA with 0 for protest size

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))

## Fixing missing values of winner.share.upd for liberal democracies

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))

## Set max_size zero for elections with no protest

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Excluding liberal democracies

nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.

## Remove outlier
nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1)

## Coding median size as 0 when n.events = 0

nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))

```

These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA. This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections. I limit the sample to authoritarian regimes which hold multiparty national elections. This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1.]

While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, democracies are excluded from this analysis for two reasons. First, the theory presented above is derived from prior work on authoritarian elections; the factors that drive the theory are expected to be weaker in democratic regimes. As regimes trend toward liberal democracy, patronage consolidation should decline and rule of law should increase by definition. Consequently, the theory expects that fraud should be rare in such cases, even in the absence of a protest deterrent: incumbents simply lack the tools. At the same time, the more liberal-democratic a state becomes, the more alternatives their are to protest as a means to check the executive. More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect to see a more pronounced role for protest risk in curtailing election fraud. The case selection strategy here is thus one that both fits the theory and makes for a strong test of the classic fraud-protest model.

Of course, treating regime type as a sharply defined binary indicator, rather than as a continuum, will introduce some bias around edge cases. As a robustness check for this case selection process, I also run the main analyses on a larger sample that adds electoral democratic regimes. While the strength of the findings is somewhat attenuated, as is expected, the results are generally upheld in the broader sample as shown in the appendix.

Models where the dependent variable is the level of election fraud include data from all cases. For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the theoretical expectations of protest risk following an election in which the incumbent won, possibly through reliance on election fraud. For models of protest-wave duration, only cases with a positive number of post-election protests are included. The date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012. The resulting dataset includes 614 total election-period observations across 97 countries.

## Dependent variables and modeling strategies

To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021]. This measure of election integrity should thus serve as a strong test of the protest-oriented model. It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest initiation and duration. As an additional check in the appendix, a fuller measure of overall election integrity is used instead.

ECAV records protests and other contentious actions related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election). To focus the study on protest, I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups. ECAV data is the event-day format; to capture the *number of protests* before the election, I take the sum number of events recorded in ECAV before election day. To evaluate hypotheses regarding *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day. While the main analysis assumes that the last protest in the dataset marks the end of the protest movement, an alternative specification in the appendix assumes right-censoring for a portion of the observations due to the 90-day observation window.]

The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nrow(nelda2.sub)` elections, no protests are recorded. Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest number and duration. As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration. This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that most fraudulent elections do not produce signals of weakness that prompt opposition protest. Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression. *Election fraud*, which follows a more normal distribution, is modeled using multilevel models with random intercepts by country. Lastly, protest wave duration is modeled using Cox proportional hazards models.

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()

p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()

p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()

 ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()

#  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
```

```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and protest duration", fig.height=7, echo=FALSE, message=FALSE, warning = FALSE}
plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.duration, ncol=1)
plot.combined.hist
```

```{r, include=F}
incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
                                            'Legislative constraints' = legcon.1lag, 
                                            'Alternative information' = altinf.1lag,
                                          'Liberal democracy index' = libdem.1lag,
                                            'GDP growth rate' = gdpgro.1lag, 
                                            'Civil society openness' = osorg.1lag, 
                                            #'Public sector corruption' = pubseccorrup.lag,
                                            'National party organization' = natlpartyorg.lag, 
                                             'Election fraud' = v2elirreg.inv, 
                                            'Presidential election' = presidential,
                                            'Incumbent term-limited' = inc.termlimit.num,
                                            'Physical integrity index' = physinteg.1lag,
                                            'Urbanization' = urban.pct, 
                                            'Rentier state' = rentierstate,
                                            'Tax state' =  taxstate,
                                            'Winning vote-share' = winner.share.upd, 
                                            'Number of pre-election protests' = n.events.pre,
                                            'Off-schedule election' = unscheduled_election,
                                          'Regional protest diffusion' = regional_success_lag,
                                         "Privatization" =  state.owned.1lag,
                                         "Professional bureaucracy" = patronage.appts.1lag,
                                        "Anti-system" = antisys.1lag,
                                        "EMB autonomy"= v2elembaut)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))

incl.vars <- incl.vars %>% mutate(`Presidential election` = as.numeric(`Presidential election`))
incl.vars <- incl.vars %>% mutate(`Rentier state` = as.numeric(`Rentier state`))
incl.vars <- incl.vars %>% mutate(`Tax state` = as.numeric(`Tax state`))
incl.vars <- incl.vars %>% mutate(`Off-schedule election` = as.numeric(`Off-schedule election`))

```

\singlespace

```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)
datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics and variable distributions ")  %>%
    kable_styling(font_size = 9, full_width = FALSE)
```
\doublespace

```{r, include=FALSE}
library(ggcorrplot)
corr <- round(cor(incl.vars, use = "pair"), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower")
```

To add: Beaulieu book; Daxecker 2019 et al (fraud perceptions), Daxecker 2012 (cost of exposing cheating), Donno 2013 (book) , Beaulieu 2014 book, hyde and marinov 2014, Morrison, Donno, Savun, and Davutoglu (2024)

## Explanatory variables

The first set of models test Hypothesis 1, on the relationship between protest risk and fraud. Protest risk is operationalized in three ways. First, the *number of pre-election protests* in ECAV is taken as *prima facie* evidence of a more turbulent and risky election environment. Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that incumbents will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b]. On the other hand, H1 predicts that incumbents will generate more manipulation in this context.

[MOVE] Moreover, in general, incumbents tend to avoid mass repression in the pre-election phase in favor of targeted harassment of opposition activists [@Bhasin2013].

Second, since a state with greater capacity for repression may be better able to deter protest and thus 'get away with' more manipulation [@Egorov2021,@Ritter2016], I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year. Higher values for this variable indicate greater freedom from torture and political killing. In one approach, higher values represent more permissive environments, and should be associated with greater protest risk (and thus reduced fraud). Alternatively, in keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and thus most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000]. I thus include the variable's square term to account for the 'murder in the middle' framework.

Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby. This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA. A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down. The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'. All told there are 15 successful electoral protests in the dataset: Bulgaria 1990, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]

The models incorporate several potential explanatory variables, in addition to the *election fraud* variable described above. The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance. The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} Finally, whether or not the chief executive is *term limited* is taken from NELDA.

## Control variables

As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), and the available of *alternative sources of information*. Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019]. All models include *legislative constraints* on the executive, a V-Dem variable which captures the extent to which the legislative body in the country acts independently in practice to limit executive power. Additional V-Dem variables capture *privatization* of the economy ('v2clstown') and *bureaucratic professionalization* ('v2stcritrecadm'), both of which might give incumbents influence over election fraud and subsequent protest. To capture the overall level of pre-election protection for rights associated with liberal democracy, I include the lagged *liberal democracy index* from V-Dem. *Physical integrity* is used as a control here, to address the risk of selection bias (if, for example, high-repression states select into a high-fraud condition with the hope that the repressive apparatus will deter protest).

Several binary control variables are also included. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest. Finally, I include dummy variables that provide a categorical evaluation of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. States that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018]. Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.

Table \@ref(tab:summary-table) shows summary statistics for all of these variables. To address concerns about multicollinearity, I present models in the appendix (page 25) that exclude control variables associated with pre-election protest.

\newpage

\singlespace

```{r manipdv, echo=FALSE}

m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub) 
#summary(m1a)

m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag  + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)

#summary(m1b)

m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1c)

m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)

#summary(m1)

m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1a)

m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)

#summary(m1b)

m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + state.owned.1lag + patronage.appts.1lag + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1d)

modelsummary(c(m1, m1b, m1c, m1a), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
                 )) %>% kable_styling(latex_options = "HOLD_position")

##Note: state owned variable is inverse coded, so should be labeled as private ownership in economy
##Note: Patronage appointments ialso inverse coded. Higher values = more professionalism
```

\newpage

\doublespace

# Results

## Election-day fraud and protest risk

The results of the first set of multilevel models are not promising for argument that protest risk deters election manipulation. Hypothesis 1a is not supported in the models shown in Table  \@ref(tab:manipdv). Models 1 through 4 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature. In Model 1, the number of pre-election protests is associated with an increase in election-day fraud. While the effect is small\textemdash ten pre-election protests only tend to increase election fraud by about six percent of a standard deviation\textemdash it runs counter to the expectations of the deterrence logic. A similar effect is found for *diffusion* in Model 4; recent successful protest in a nearby country is also associated with a modest increase in expected fraud. This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results and/or to signal organizational strength.

Model 2 does not demonstrate a significant linear relationship between repression and election fraud. However, Model 3 shows that fraud first increases with physical integrity, but then declines. In other words, fraud is lowest on the extremes of repressive capacity\textemdash where repression is severe and where it is most limited. This is contrary to the expectation that a higher risk of protest is associated with reduced fraud.  Under the expectation that protest risk is highest for these middle values [@pierskallaProtestDeterrenceEscalation2010], H1a is not supported. By contrast, H1b is supported by Models 1, 3, and 4: by each measure, fraud is higher when protest risk is higher.

## Fraud and the likelihood of protest

The second set of models challenge the idea that more severe electoral manipulation is associated with an increased risk of protest. Beginning with the zero-inflation portion of the hurdle models in Table 3, which model the likelihood of *no* protest occurring after the election, the standalone coefficient for election fraud is not significant in Model 5. Figure \@ref(fig:margin-plot-gdp) shows that there are no significant effect across values of *lagged GDP growth*.

More fraudulent elections do appear associated with an increasing risk of protest initiation as margins of victory shrink, as shown in the top panel of Figure  \@ref(fig:margin-prob-fig). Here, typical values for low and high levels of fraud are used (one standard deviation below and above the mean, respectively). The top panel shows that more fraudulent elections are associated with a significantly higher risk of protest initiation, but only when winning vote-shares are extremely small.

\newpage

\singlespace

```{r include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary)

model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary_noc)

## Plots

####Margin

#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

```

```{r logit-table, echo=FALSE}
models2 <- list()
models2[['Model 5']] <- model.margin.binary_noc
models2[['Model 6']] <- model.margin.binary

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

\newpage
\doublespace

Table \@ref(tab:logit-table) shows the results for Hypothesis 2, which is expected to be supported across all three schools of thought. Model 5 is a base model, where incumbent margin is interacted with *intentional irregularities*; Model 6 includes the same interaction along with additional control variables that could be associated with both fraud and post-election protest. The expected, significant interaction effect is detected in both models, and it is in the expected direction: the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.

Figure \@ref(fig:margin-prob-fig) illustrates just how tenuous this effect is, however. The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables). The figure highlights that the predicted probability of protest is only statistically significant for the two values of fraud at extremely low levels of incumbent vote-share. This represents an exceedingly unlikely outcome along both variables. As the figure shows, there are only a small number of observations in the range where the predicted probabilities are statistically significant. At the same time, the values for *intentional irregularities* are also highly unusual. They represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.

While all three schools of thought are consistent with H2, the results here may be taken as circumstantial evidence in favor of the fraud-as-strength model. Fraud only appears significantly associated with increased protest risk when the incumbent's result in the election is very low and fraud is very high. This is suggestive of the model where opposition actors interpret the result as showing extremely low genuine support for the incumbent, perhaps enough that it outweighs the signal of strength sent by fraud for some oppositions. Middle values for vote-share combined with high values of fraud suggest durable regime strength, bringing expected protest close to zero. In the alternative view where fraud conveys weakness, we might expect the probability of protest to rise more sharply as incumbent vote-share declines, and to reach higher predicted probabilities. This pattern is only suggestive, of course, and for more definitive evidence I look to protest duration.

```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
plot_binary_margin_typ_n
```

## Protest wave duration

[CLEAN UP]

Substantively, even if one does not fully discount the claim that the null results shown here represent equilibrium behavior, it is difficult to dispute that the resulting outcome is weighted heavily in favor of governments. In the dataset of 1110 elections used here, only 19 elections in five countries are coded as having an *intentional irregularities* score of less than -2 (indicating extremely fraud-free elections). By contrast, 121 elections in 27 countries score higher than 2 (indicating highly fraudulent elections). The frequency of fraud, the rarity of protest, and the finding that fraud severity is not correlated with the scope of protest events all point to the same conclusion: protest risk is not a substantial constraint on election fraud in electoral authoritarian and electoral democratic regimes.

```{r, include=FALSE}
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 62, 0, 1)))

nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)

km <-  with(nelda2.sub.km, Surv(d.events.num, surv_status))

survfit2(Surv(d.events.num, surv_status) ~ 1, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
  ggsurvfit() +
#   add_confidence_interval() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )

cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
summary(cox1)
cz1 <- cox.zph(cox1, transform = "identity")
print(cz1)
plot(cz1[16])
abline(h=0)
abline(v = 21)
abline(v = 61)

cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
summary(cox1.cen)

cox1t <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(v2elirreg.inv), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(15, 60, 90),
Boundary.knots = FALSE))
summary(cox1t)

cz <- cox.zph(cox1, transform = "identity") #function(time) 1/(log(time + 1)))
print(cz)
plot(cz) # The points on the plot are residuals of v2elirreg.inv (and other vars) based on hte predicted value of X and time t. The line shows the estimated value of the coefficient on variable X at time t compared to its baseline. You want it to be zero for consistent effects over time. V2elirreg.inv is positive for about two weeks early on. The substantive effect of this can be shown by taking two values of v2elirreg.inv (i.e. first and third quartile), taking the difference, and finding exp(coefficient_residual * difference). The result shows here that protests in states in the third quartile of fraud are exp(.75*1.4) = 2.86 times more likely to die than those in the first quartile during those couple weeks. So this suggests that if protests survive the first week or two, the state's capacity starts to kick in?

cox1t2 <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(n.events.post), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(10, 70, 90),
Boundary.knots = FALSE))
summary(cox1t2)

sandbox_data <- survSplit(Surv(d.events.num, surv_status) ~ 
                      jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                      data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), cut = c(2), episode = "tgroup")

cox1.2 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + fraud_level*(as.factor(tgroup)) + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = sandbox_data)
summary(cox1.2)

cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ration", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox2.cen)

cox3 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + gdpgro.1lag*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag , data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox3)

cox4 <- coxph(Surv(d.events.num, surv_status) ~ jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox4)

```

To further test the mechanism outlined in the theory above, I examine the relationship between fraud and the survival of protest movements in cases where protests began but the incumbent eventually survived. Restricting the sample to these cases allows for a consistent interpretation of results: longer protest waves indicate a more even match between the resources and resiliency of the protesters and the capacity and cohesiveness of the regime, while shorter-lived protest movements are those which were more easily defeated. If significant election fraud is indicative of regime resilience, the expectation is that more fraudulent elections will be associated with more quickly dispatched protest movements.^[Protest waves are defeated in the large majority of cases. There are 211 cases where protest occurs, out of which only 13 see new elections or the incumbent step down.]

Table \@ref(tab:cox-models-tab) reports the results of Cox proportional hazard models of protest duration. In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived; those with a negative sign indicate factors associated with a slower demise. The coefficient for *election fraud* is positive and significant in Model 7. Exponentiation of the coefficient for election fraud gives a hazard ratio of 3.3, indicating that for each one-unit increase in fraud, a protest movement becomes three times more likely to end at any given time. This supports H3b, and fails to support H3a.

Further investigation of the relationship between fraud and protest duration yields additional insights. In Model 8, incumbent vote-share is interacted with *election fraud*. Figure \@ref(fig:margin-dur-fig) shows the marginal effect. As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end. This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.

```{r margin-dur-fig, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
p.duration.margin
```

The relationship between the level of fraud and the survival of protest movements can also be demonstrated non-parametrically using Kaplan-Meier curves for varying levels of fraud. In Figure \@ref(km-fig), observations are divided into three groups. High fraud observations are those where *election fraud* is scored at one standard deviation above the mean or more, low fraud observations are those scored at one standard deviation below the mean or less, with moderate levels in between. The figure shows a clear difference between the low-fraud group and the other two. Low-fraud cases are much more likely to see their protest movements survive past the first day than their counterparts. For much of the range of the data, they remain roughly twice as likely to survive at any given date than protest movements in medium- or high-fraud cases. This is evidence in favor of the idea that high levels of fraud is an indication of regime resilience, and contradicts the claim that fraud either reveals weakness or galvanizes political opposition, as it is the low-fraud cases where governments struggle to put an end to protest movements on average.

```{r km-fig, fig.cap = "Kaplan-Meier curves for protest duration and high, low, and medium levels of fraud", fig.height= = 7, echo=F, message=F, fig.pos='H'}
nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))

survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
  ggsurvfit() +
#   add_confidence_interval() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )
```

In Model 7, all covariates' effects are modeled as time-invariant. However, diagnostics of the model suggest that this assumption may not be upheld for *election fraud*.^[Schoenfeld residuals for *election fraud* are not independent of time, as calculated using the `survival` package in R [@survival-package].] To adjust for this, Model 12 allows the coefficient for *election fraud* to take on different values at different points in time. Based on the diagnostics, I allow the value of the coefficient to vary for three periods: 1 - 11 days, 12 - 25 days, and 26 days to the maximum value. Model 12 shows the result of this analysis; in it, *election fraud* has a positive and significant coefficient for the earliest time period. This indicates that, for the first 11 days of a protest movement, states that engaged in more election fraud are more likely to be able to quash the protests (whether by repression or accommodation). This is contrary to the expectations of deterrence models. After this period, there is no significant relationship between the level of fraud and the survival of the protest wave. That is, if a protest movement has survived longer than (roughly) a week and a half after the election, the severity of fraud does not influence its continued duration. It is also noteworthy that this formulation is a much better fit to the data than the linear effect in Model 11, with a much lower root mean square error.

These results indicate that the influence of a fraudulent election on the duration of mass protest diminishes over time. It also suggests that most protests are failed bets. Upon taking to the streets, protesters are more likely to find themselves facing a capable incumbent regime the more fraudulent the election was. Finally, if protests survive beyond the short term, the severity of fraud no longer has any effect. Instead, structural indicators\textemdash the strength of the opposition in the legislature, the degree to which parties are organized across the country, and (to a lesser extent) the openness of civil society are all associated with longer-lived protest movements. Similarly, a more fractured incumbent regime, as indicated by a term-limited incumbent, also make protest survival more likely. All of this is supportive of the theoretical model laid out above.

```{r cox-models-tab, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 7']] <- cox1
models3[['Model 8']] <- cox2
models3[['Model 9']] <- cox1t

modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 5: Cox proportional hazard models of protest duration', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", 
                 "winner.share.upd" = "Incumbent vote-share",
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
                 "tt(v2elirreg.inv)1" = "Election fraud (day 2-20)", "tt(v2elirreg.inv)2" = "Election fraud (day 21-90)"
                 ))

```

\doublespace

# Discussion

Altogether, the findings challenge the idea that the fear of protest helps deter election manipulation. To the extent that fraud is a signal of incumbent strength, it is an unreliable one. Since even strong incumbents commit fraud, the decision to take to the streets is often a losing gamble\textemdash anticipating a weak incumbent, they instead find one that is capable of repressing, coopting, or conciliating an emerging protest movement. Indeed, while there are 194 cases of protest out of the 661 elections in this dataset where incumbents claimed victory, only 8 resulted in either the incumbent being replaced or the election being re-run due to mass protests.\footnote{This information is determined using the NELDA variables NELDA37 and NELDA41.} This suggests a `success rate' for post-election protest of about four percent *conditional on protest initiation*, and about one percent for all elections in electoral authoritarian and electoral democratic regimes.

The results indicate that we should be cautious when building and interpreting theories of politics outside liberal democratic regimes that emphasize the risk of mass protest as a trip-wire to be avoided, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models. The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes. Both protest and election fraud are complex phenomena with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a significant factor under most circumstances.

This has several implications for understanding election manipulation and regime resilience. By overlooking the poor quality of election fraud as a signal, or by treating protest primarily as a binary event, the existing literature on electoral protest has overemphasized the risk to incumbents of engaging in electoral manipulation. The results presented here show that fraud has no bearing on how numerous or large protest events are. Moreover, heightened protest risk appears to be associated with *more* election fraud, rather than less. Instead, the conditions that are correlated with more damaging protest are generally institutional factors, especially executive term limits, which are either more amenable to control or can be anticipated by incumbents.

These are pessimistic findings for proponents of democracy. That electoral protest is less risky than is often assumed, in turn, implies that governments and ruling parties have a freer hand to engage in electoral manipulation than many models posit. Since the severity of fraud has no apparent bearing on whether or not a major protest wave will emerge as a serious threat to the regime, government have an incentive to manipulate as much as possible to secure a large winning vote-share. This perverse incentive undermines the case that free and fair elections in an electoral democracy represent a self-enforcing equilibrium backed up by the risk of citizen protest [@fearon_self-enforcing_2011], or that risk of popular mobilization can substantially moderate election manipulation in non-democracies. In effect, it is the existence of an election itself [@lucardiStrengthExpectationElections2019], and the opportunity it offers for opposition mobilization [@howard2006], that creates protest risk for incumbents, giving them a relatively free hand to manipulate to the extent they can recruit agents to do the work.

This study also suggests implications for the study of election manipulation. First, the study implies that supply-side theories of electoral manipulation\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important than protest risk for understanding election integrity, which may drive future research as well as policy interventions. Third, and relatedly, it undermines some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity. As noted earlier, prior studies argue that these actors restrain manipulation by revealing information about fraud and intensifying protest risk. But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work. Scholarly attention might turn more profitably toward supply-side explanations for manipulation, such as the organizational capacity of the state [@simpser2013], the popularity of the government [@rundlett2016deliver], and the risks faced by front-line agents [@Harvey2022].

Second, it suggests avenues for future research. It is clear that electoral protests charged by allegations of manipulation do occur at times. While the main argument of this paper is that election fraud itself does not predict these protests, there may still be connections between manipulation and protest to uncover. For example, it could be that other forms of manipulation\textemdash perhaps intimidation or violence\textemdash are associated with higher risk of escalating protest. Perhaps some forms of electoral manipulation are particularly offensive to certain groups\textemdash young people, highly educated people, people with ideological commitments to democracy, etc. To the extent that countries vary according to the size of such populations, they may vary in their underlying protest risk.

As a final point, the results also speak to incidents of mobilization around unfounded allegations of election-rigging, such as those of January 6, 2020 in the United States or 2022-2023 in Brazil. Indeed, out of the 293 instances of post-election protest in this dataset, 18 occurred after very clean elections (those one standard deviation below the mean for election fraud)\textemdash a small but not minuscule percentage. These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation. This possibility tends to reduce the 'opportunity cost' to incumbents of engaging in real fraud, while also underlining the fact that the initiation and course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.

These results are consistent with different specifications presented in an online appendix. In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data (appendix page 1), and when a measure of overall election integrity is used in place of the measure of election fraud (appendix page 15).

# Conclusion

Protest risk is often considered to be a major risk to incumbent governments aiming to rig elections. This risk has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors and other restraints on the ruling party. However, I argue that such models overlook the ambiguity inherent in election manipulation efforts, as well as the quick decay of even such limited information once protests begin. Empirical research on the subject has largely treated protest as a binary variable, an approach which obscures the distinction between small, one-off protests and much more costly protest waves.

This paper addresses these concerns by testing the relationship between election fraud and the number and maximum size of protest events that occur after an election. Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 1110 elections to show that election fraud has no significant relationship with the number or size of protests that occur in a protest wave. Additional tests suggest that this null finding is not the result of equilibrium behavior predicted in deterrence models. Instead, most fraudulent elections go un-protested, and most post-election protests appear to be strategic gambles gone wrong. The initiation of protest more often reveals the incumbent to be strong enough to manage protest through repression, cooptation, or other concessions. The results suggest that theories which treat protest risk as an unqualified deterrent to election manipulation should be reconsidered.

The absence of protest risk as a guardrail is felt in both electoral authoritarian regimes and in electoral democracies\textemdash those democracies that are most at risk of democratic erosion. The evidence thus suggests that mass demonstrations in favor of advancing (or restoring) democratic elections are rare and unlikely to be driven by the severity of manipulation---a gloomy picture for supporters of democratization---and that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.

```{r, include=FALSE}
ggplot(data=nelda2.sub, aes(x=v2elirreg.inv, y = jitter(any.ecav.protest), color = as.factor(v2x_regime))) +
  geom_point()

```

Predicting post-election protest counts and getting residuals:

```{r, include=FALSE}
test3 <- glm.nb(n.events.post ~  jucon.1lag + legcon.1lag + altinf.1lag + libdem.1lag + gdpgro.1lag +  + osorg.1lag  + natlpartyorg.lag +   inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election  + regional_success_lag, data = nelda2.sub)
summary(test3)

ggplot(mapping = aes(x=predict(test3, type = "response"), y= test3$model$n.events.post)) + geom_point() + coord_cartesian(xlim=c(0, 500))

newdata <- modelr::add_predictions(nelda2.sub, test3)
newdata <- newdata %>% mutate(pred_error = n.events.post - pred)
  #because this is set up so that the predicted value is subtracted from the actual, underpredicting
  # protest is a positrive value (i.e. actual 5 - 0 predicted = 5)

test4 <- lm(pred_error ~ winner.share.upd + v2elirreg.inv + presidential, data = newdata)
summary(test4)

test5 <- lm(pred_error ~ winner.share.upd*v2elirreg.inv + presidential, data = newdata)
summary(test5)

sjPlot::plot_model(test5, type = "pred", terms = c("v2elirreg.inv", "winner.share.upd [.4, .7]"))

plot(newdata$v2elirreg.inv, newdata$pred_error)
```

Predicting the level of election manipulation, and then seeing if actual values above prediction are associated with protest

```{r, include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(inc.confident.num = ifelse(inc.confident == "N/A", NA, ifelse(inc.confident == "yes", 1, 0)))

###For testing missingness
subsample <- nelda2.sub %>% dplyr::select(v2elirreg.inv, jucon.1lag, legcon.1lag, gdpgro.1lag,  natlpartyorg.lag,  presidential, urban.pct, rentierstate,  taxstate,  unscheduled_election, regional_success_lag, physinteg.1lag, antisys.1lag, csenv.1lag, preelection_concerns, inc.termlimit.num, liberal_no_jucon, v2elirreg.inv.lag, patronage.appts.1lag,
                                                  state.owned.1lag)
summary(subsample)
############################ Note: Drop legcon from analysis because of high missingness and high cor with lib_no_jucon

test6 <- lmer(v2elirreg.inv ~ jucon.1lag +  gdpgro.1lag +  natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + physinteg.1lag + antisys.1lag + csenv.1lag + preelection_concerns + inc.termlimit.num   + liberal_no_jucon + v2elirreg.inv.lag +
                patronage.appts.1lag +  state.owned.1lag +  v2elembaut + (1 | stateid), data = nelda2.sub) 
summary(test6)

testdata <- broom.mixed::augment(test6)
  #This adds fitted values and residuals.
  #Residuals are actual values minus predicted value, so
  #over-predicting fraud will result in a negative number (actual fraud < predicted )
  #Under-predicting fraud--what we are interested in--will result in a positive number (actual > predicted)

## Need to add rownames to nelda2.sub, so that can merge variables from nelda2.sub into testdata
nelda2.sub <- nelda2.sub %>% mutate(.rownames = rownames(nelda2.sub))
neldavars <- nelda2.sub %>% dplyr::select(.rownames, n.events.post, inc.lose, winner.share.upd, country, year, diem_explicheat_max, diem_fraud_max, country_id_year, surv_status,  inc.replaced.prot, d.events.num,
                                          legcon.1lag, osorg.1lag, altinf.1lag, libdem.1lag,  winner.share.upd, n.events.pre)
testdata <- testdata %>% left_join(neldavars, by = ".rownames")

model_protest <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag + .resid*winner.share.upd,
                               ziformula=~.,
                              data = testdata %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest)

model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + .resid  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data = testdata %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest2)

plot(testdata$winner.share.upd, testdata$.resid)

testdata <- testdata %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
testdata <- as.data.frame(testdata)
model_protest_binary <- glm(no_protest ~ .resid*winner.share.upd, family = binomial(link = "logit"),
                             data = testdata %>% filter(inc.lose == 0))
plot_binary_margin_typ_n <- sjPlot::plot_model(model_protest_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"))
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Excess election fraud", y = "Predicted probability", title = "Predicted probability of no protest", color = "Incumbent vote-share") + theme_bw()

model_protest2_binary <- glm(no_protest ~ .resid + winner.share.upd, family = binomial(link = "logit"),
                             data = testdata %>% filter(inc.lose == 0))

sjPlot::plot_model(model_protest2_binary, type = "pred", terms = c(".resid", "winner.share.upd [.2, .8]"), show.data = T)
sjPlot::plot_model(model_protest2_binary, type = "est", vline.color = "black") + theme_bw()

### Seeing how residuals compare to DIEM data

summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = testdata))
plot(testdata$diem_explicheat_max, testdata$.resid)
plot(testdata$diem_fraud_max, testdata$.resid)
  ##This shows that the resids are basically noise with regard to election observer reports, which should be good; I think this helps push back against the idea that V-Dem coders are giving higher scores to protest elections. But it would be good to push deeper and look at the correlation only among the most up-coded resids.

high_resids <- testdata %>% filter(.resid >= .222)
summary(lm(.resid ~ diem_explicheat_max + diem_fraud_max, data = high_resids))

ggplot(data = testdata, aes(x = jitter(diem_fraud_max), y = .resid)) + geom_point(aes(size = n.events.post))

plot(high_resids$diem_explicheat_max, high_resids$.resid)
plot(high_resids$diem_fraud_max, high_resids$.resid)
  #We do see more outliers here, good to think deeper--there are two major outliers that are also big protest events

### Duration

cox1_b <- coxph(Surv(d.events.num, surv_status) ~ winner.share.upd  + .resid, data = testdata %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
summary(cox1_b)

## DIEM and protest

model_protest_diem <- glmmTMB(n.events.post ~   winner.share.upd +  diem_explicheat_max ,
                               ziformula=~.,
                              data = nelda2.sub %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest_diem)

```

```{r, include=FALSE}
nelda2.sub <- nelda2.sub %>% group_by(country) %>% mutate(elirreg_mean = mean(v2elirreg.inv, na.rm=T))
nelda2.sub <- nelda2.sub %>% mutate(elirreg_mean_diff = v2elirreg.inv - elirreg_mean)

model_protest1 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data =nelda2.sub %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest1)
sjPlot::plot_model(model_protest1, type ="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]")) + coord_cartesian(ylim = c(0, 50))

nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))
model_protest_binary <- glm(no_protest ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd*elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                             data =nelda2.sub %>% filter(inc.lose == 0),
                                family=binomial(link = "logit"))
summary(model_protest_binary)
sjPlot::plot_model(model_protest_binary, type="pred", terms = c("elirreg_mean_diff" , "winner.share.upd [.2, .8]"))

model_protest2 <- glmmTMB(n.events.post ~  jucon.1lag + legcon.1lag + osorg.1lag + altinf.1lag + libdem.1lag  + natlpartyorg.lag +  gdpgro.1lag  +   winner.share.upd + elirreg_mean_diff  + presidential + inc.termlimit.num  +  physinteg.1lag  + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag,
                               ziformula=~.,
                              data =nelda2.sub %>% filter(inc.lose == 0),
                                family=truncated_nbinom2)
summary(model_protest2)

```

\newpage
# References