### 2024-02-09 Output/appendix_v4.Rmd
Changes to appendices and fixes in main text

---
title: "Appendix: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: \usepackage{booktabs}
---

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
library(survival)
library(modelsummary)
library(glmmTMB)
library(huxtable)
library(kableExtra)
library(simPH)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))

#Coding as factors

nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Replacing NA with 0 for protest size

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))

## Fixing missing values of winner.share.upd for liberal democracies

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))

## Set max_size zero for elections with no protest

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Excluding liberal democracies

nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.

## Remove outlier
nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)

## Coding median size as 0 when n.events = 0

nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))

## Any protest

nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))

## Surv status
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))

```

# Alternative measures of election manipulation

To ensure that the results presented in the main text are not biased by the chosen measure of election fraud, this section replaces that measure with a broader indicator of election integrity. This variable is `v2xel_frefair,' from V-Dem, which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election. This variable ranges from 0 to 1, with higher values indicating freer and fairer elections. The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables, using the ECAV data on protest. Since the dependent variable of these models is on an inverse scale, compared to the variable *intentional irregularities*, we should expect to see coefficients for key variables take on opposite-signed values from the models in the main text.

```{r manipdv-full, echo=FALSE}
m1 <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  +   (1 | stateid), data = nelda2.sub)

#summary(m1)

m1a <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag  +  (1 | stateid), data = nelda2.sub) 
#summary(m1a)

m1b <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election  +  (1 | stateid), data = nelda2.sub)

#summary(m1b)

m1c <- lmer(v2xel_frefair ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election   +  (1 | stateid), data = nelda2.sub) 
#summary(m1c)

models <- list()
models[['Model 1']] <- m1
models[['Model 2']] <- m1a
models[['Model 3']] <- m1b
models[['Model 4']] <- m1c

modelsummary(models, stars = T,  title = 'Random-intercept multilevel model of overall election integrity', gof_omit = "AIC|BIC",
             coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", "presidential" = "Presidential", "n.events.pre" = "N. of pre-election protest",
                 "physinteg.1lag" = "Physical integrity (lag)",
                 "physintlag.sq" = "Phys. integ. (lag), squared", "unscheduled_election1" = "Off-schedule election",
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate1" = "Rentier state", "taxstate1" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests",
                 "state.owned.1lag" = "Private ownership (lag)",
                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
                 "regional_success_lag" = "Regional protest diffusion"))
```

As expected, the first model shows that the number of pre-election protests is negatively associated with overall election integrity\textemdash contrary to deterrence expectations and in line with the argument of the main text. We also see that overall election integrity is lowest at middle levels of repression; this is, again, in line with the expectations of the incumbent strength model and contrary to deterrence expectations. No significant effect on overall election integrity is observed for successful electoral protests in the region.

```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F, echo=FALSE}

model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd +  presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary)

model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary.int)

model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary_noc)

```

Moving to the logit models of post-election protest, we observe the same interaction relationship between overall election integrity, incumbent vote-share, and protest as is shown for fraud in the main text. That is, less free-and-fair elections are only associated with a higher protest risk and low levels of incumbent vote-share.

```{r logit-table, echo=FALSE}
models2 <- list()
models2[['Model 5']] <- model.margin.binary
models2[['Model 6']] <- model.margin.binary.int
models2[['Model 7']] <- model.margin.binary_noc

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                 "v2xel_frefair" = "Overall election integrity",
                 "winner.share.upd" = "Incumbent vote-share",
                 "v2xel_frefair:winner.share.upd" = "Election integ. : incumbent vote-share"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

\newpage

```{r overall-binary, fig.cap="Overall election integrity, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))  +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election integrity") + theme_bw()

```

```{r, echo=FALSE}

cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2xel_frefair  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
#summary(cox1)

cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2xel_frefair*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
#summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.14, .43]"))  #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election free and fair") + theme_bw()

```

```{r cox-models-tab-overall, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 8']] <- cox1
models3[['Model 9']] <- cox2

modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", 
                 "winner.share.upd" = "Incumbent vote-share",
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2xel_frefair" = "Overall election integrity",
                 "n.events.post" = "N. post-election protest"
                 ))

```

```{r margin-dur-fig-overall, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H'}
p.duration.margin
```

Likewise, the models of protest movement survival are also in line with those in the main text. Table \@ref(tab:cox-models-tab-overall) shows that cleaner elections are associated with a slower end to protest movements (here, a negative coefficient). When elections are more unfair overall, incumbents are able to more quickly bring protest to an end. Figure \@ref(fig:margin-dur-fig-overall) shows that this effect is significant in interaction with low-to-moderate incumbent vote-shares.

These three sets of models all replace a measure of *election fraud* with a broader measure of overall election integrity. In all three cases, the results broadly confirm the incumbent-strength model and are unsupportive of the deterrence model.

\newpage

# Repression and pre-election protest

The main analysis already controls for possible confounding by the state's repressive capacity; i.e. the possibility that in highly repressive states, protest is deterred and fraud can proceed unopposed. However, it may still be useful to check an alternative explanation\textemdash perhaps manipulation is deterred in highly repressive states with considerable pre-election unrest (that is, when protest appears to be exceeding the ability of the state to contain it). To check this possibility, a variation of Model 1 from the main text is run, but with an interaction between the number of pre-election protests and the lagged *physical integrity* variable. If protest risk deters manipulation in this way, there should be a negative and significant effect for the number of pre-election protests in highly repressive states.

Figure \@ref(fig:prot-rep-fraud) shows the absence of such an effect. In fact, there is no significant change in the predicted level of fraud at lower levels of physical integrity, across all levels of pre-election protest. We do observe a significant effect for low repression regimes; however, it is in the opposite direction predicted by this counterargument. That is, more numerous pre-election protest is associated with higher predicted levels of fraud in less repressive states, not harsher ones. The figure also shows that less repressive regimes appear to respond to greater civic unrest with more fraud. Altogether, this robustness check shows that pre-election protest in a repressive environment does not appear to deter fraud, and supports the view that manipulation may be used to indicate regime durability in a challenging period.

```{r prot-rep-fraud, message=FALSE, warning=F, echo=FALSE, fig.cap="Marginal effect of pre-election protest on fraud, conditional on physical integrity", out.width = '75%', fig.align='center'}
m1_rep <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut + physinteg.1lag*n.events.pre + (1 | stateid), data = nelda2.sub) 
#summary(m1_rep)
p.m1rep <- sjPlot::plot_model(m1_rep, type = "pred", terms = c("n.events.pre", "physinteg.1lag [.2, .7]")) #.25 and .75 quantile
p.m1rep + labs(title = "Predicted values of election fraud", x = "Number of pre-election protests", y = "Election fraud", color = "Physical integrity (lag)") + theme_bw()
```

\newpage

# Chosen successors

The models in the main text control for the presence of term-limited executive incumbents, based on the logic in Hale [@hale_patronal_2014]. However, it could be the case that the nature of a term-limited election affects the behavior of protesters and election manipulators alike; that is, perhaps when there is a designated successor, the heir apparent is better able to mobilize the resources of the state to manipulate elections and suppress protest. Using a variable from NELDA that codes for designated successors, the following models add a categorical variable for the incumbent's status, which can take on the following values: *executive eligible*, *chosen successor*, *contested executive election*, and *legislative election*. Including this variables requires the exclusion of the original variables *presidential election* and *incumbent term limited*.

```{r successor-table, echo=FALSE, message=FALSE, warning=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Executive eligible"))
nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))

## Protest binary

model.margin.int.suc <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + incumb.status  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.int.suc)

## Duration

cox2.suc <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + incumb.status +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv*winner.share.upd  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
#summary(cox2.suc)

### Table

models4 <- list()
models4[['Model 10']] <- model.margin.int.suc
#models4[['Model 6']] <- model.d.base2.hurd
models4[['Model 11']] <- cox2.suc

t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, coefs = c("Legislative election" = "incumb.statusLegislative election", "Chosen successor" = "incumb.statusChosen successor", "Contested exec. election" = "incumb.statusContested exec. election", "Intentional irregularities" = "v2elirreg.inv", "Incumbent vote-share" = "winner.share.upd", "Intentional irregularities : Incumbent vote-share" = "v2elirreg.inv:winner.share.upd"),  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Protest initiation", "Protest end", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 14)
theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)

```

```{r n-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
dat <- ggpredict(model.margin.int.suc, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), condition = c(incumb.status = "Executive eligible")) 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Probability", title = "Predicted probability of protests", color = "Election fraud") + theme_bw()
```

```{r duration-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
p.s.margin <- sjPlot::plot_model(cox2.suc, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]")) +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio of protest end", color = "Election fraud") + theme_bw()

```

As Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text. For these figures, *incumbent status* was set to an executive election in which the incumbent is eligible. In this model, higher levels of fraud increase the likelihood of protest following the election, at low levels of incumbent vote-share; this is in line with expectations of both the deterrence and incumbent-strenght models. However, the results for protest survival models support the incumbent-strength model.

Table \@ref(tab:successor-table), abridged here to show the key variables, shows that there is no effect of successor type on the duration of electoral protest. There is, however, the same relationship between election fraud, incumbent vote-share, and protest survival that is found in the main text. That is, incumbents who report low vote-shares but make use of large-scale fraud are better able to end protest movements than those who produce less fraud.

```{r successor-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
plot.combined.prob <- cowplot::plot_grid(p.n.margin, p.s.margin, ncol=1)
plot.combined.prob
```

\newpage

# Removing controls to address multicollinearity

The models in the main text include a suite of control variables aimed at controlling for common causes of election manipulation and protest, in order to minimize the risks of confounding and omitted variable bias. However, this approach runs the risk that individual coefficients may be highly sensitive to model specification due to multicollinearity. In particular, many of the control variables\textemdash such as *legislative constraints*, *judicial independence*, and *physical integrity* are likely to be correlated with pre-election protest. To address this risk, the models in Tables \@ref(tab:logit-table-small) and \@ref(tab:cox-models-tab-small) exclude non-electoral control variables. That is, they include variables related to the election itself, the measure of *pre-election protest*, and the explanatory variables from the main models.

The results echo those from the main text. Fraudulent elections are associated with a higher risk of protest initiation at low levels of ruling-party vote-share (Figure \@ref(fig:margin-prob-fig-small)), though these same conditions are associated with a higher likelihood that protest movements will fail quickly \@ref(fig:margin-dur-fig-small).

```{r include=FALSE, echo=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.margin.binary <- glm(any_protest ~  v2elirreg.inv +  winner.share.upd  + presidential + inc.termlimit.num  +  n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary)

model.margin.binary.int <- glm(any_protest ~ v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary.int)

## Plots

####Margin

#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

```

\newpage

```{r logit-table-small, echo=FALSE}
models2 <- list()
models2[['Model 12']] <- model.margin.binary
models2[['Model 13']] <- model.margin.binary.int

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                 "v2elirreg.inv" = "Election fraud",
                 "winner.share.upd" = "Incumbent vote-share"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

\newpage

```{r margin-prob-fig-small, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
plot_binary_margin_typ_n
```

```{r, include=FALSE, echo=FALSE}
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))

cox1 <- coxph(Surv(d.events.num, surv_status) ~   winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
#summary(cox1)

cox2 <- coxph(Surv(d.events.num, surv_status) ~  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag +  n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
#summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]"))  #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

```

```{r cox-models-tab-small, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 14']] <- cox1
models3[['Model 15']] <- cox2

modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
               c("winner.share.upd" = "Incumbent vote-share",
                 "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "unscheduled_election" = "Off-schedule election",
                 "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
                 "n.events.post" = "N. post-election protests"
                 ))

```

```{r margin-dur-fig-small, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
p.duration.margin
```

\newpage