### 2024-07-15 Scripts/authoritarian_only_latex.Rmd
Beginning post-BJPS updates

---
output:
    pdf_document:
      keep_tex: yes
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
indent: true
abstract: "The risk of mass protest figures prominently in many models of election manipulation, as a deterrent to tampering with elections. Yet election manipulation is common outside liberal democracies, while protest is rare (and successful protest rarer still). Using data on election fraud and duration data for protest events from 647 elections from 1990 to 2012, this paper argues that the risk to incumbents has been overstated in prior work. It argues that incumbents who solve the principal-agent dilemmas associated with election fraud are likely to be highly resilient to protest. There are two main findings. First, a riskier pre-election environment for incumbents is associated with more fraud, not less. Second, when protests do occur, greater election-day fraud is associated with a faster demise for protest movements. This pessimistic finding implies that election fraud is less risky than much prior work assumes, and helps explain the long-term stability of electoral authoritarianism as a regime type."
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{setspace}
  - \usepackage{float}
  - \usepackage{caption}
  - \captionsetup[table]{labelformat=empty}
---

\doublespace

The December 2011 elections to the Russian State Duma delivered a significant setback for the ruling party. It claimed only forty-nine percent of the vote, and while it retained the majority in the legislature, it lost more than 75 seats from the previous convocation. This narrow 'victory' occurred despite widespread election-day manipulation, detected in many cases by domestic election observers. The election triggered one of the most substantial protest movements in post-Soviet Russian history. More than 100,000 people turned out in protest on two occasions in Moscow, with tens of thousands more taking to the streets in regional cities [@Dollbaum2017]. Protests continued regularly in advance of the March 2012 presidential election, declining afterward.

Despite the significant challenge, the ruling party was able to defuse the protest movement over a period of months. Through a combination of repression and concessions [@golosov2012], the ruling party was able to reassert itself and continue a long process of sidelining the political opposition in Russia [@gelman_2015_troubled]. The post-2011 protest wave is one much-studied case of the phenomenon of electoral protest aimed at punishing electoral manipulation. Yet the Russian example raises questions for that broader phenomenon: does election fraud itself make protest more likely, or only in the context of a weak electoral showing for the ruling party? Given the ruling party's poor result and substantial electoral fraud, why was it able to dispatch the mobilized opposition with relative ease?

These questions get to the heart of a debate in the study of electoral manipulation and protest: does election fraud signify regime weakness or strength, and in what ways? The various theoretical perspectives brought to bear on the topic have been built on different assumptions, and yield different implications for authoritarian governance and democratization. For some models, election fraud is a tool of weak incumbents. As a result, incumbents avoid fraud when it might be exposed and reveal their weakness to potential protesters\textemdash an auspicious prediction for democratization and ultimately the under-pinning of 'self-enforcing' democracy [@fearon_self-enforcing_2011]. Another school of thought holds that election fraud can be an indicator of incumbent strength. This approach, which is expanded upon below, is a pessimistic account for democracy. If this model is correct, it is authoritarian politics rather than democracy that is self-enforcing\textemdash the ability to commit large-scale election fraud helps win elections, signal strength, deter challengers under most conditions, and put down protest movements if they do begin. This paper argues that this pessimistic model best fits the overall pattern of the data, and is supported by specific, novel hypothesis tests.

In this paper, I argue \textemdash in keeping with a growing literature on the informational aspects of elections in authoritarian regimes [@guriev2020;@smyth2020elections;@Geddes1989;@Miller2015a;@Pop-Eleches2015]\textemdash that an election outcome produces a signal of the relative strength of the incumbent. As Magaloni [-@magaloni2006] and others have noted, incumbents typically work to produce a higher margin of victory in the election, in order to demonstrate their strength and induce cooperation from other actors in society. In some theories, this works best if the votes for the incumbent are largely 'genuine' (i.e. not fabricated), or produced using the incumbent's organizational resources (i.e. vote-buying and intimidation) [@Harvey2020;@magaloni2006]. These two kinds of votes correspond to two of the three 'pillars of stability' identified by Gerschewski [-@Gerschewski2013]: legitimation and cooptation. Fraud is considered a form of cheap talk in these models, which opposition actors can discount if they become aware of it [@little_elections_2012].

However, these models overlook the fact that engineering widespread fraud requires solving principal-agent problems through consolidated control over patronage resources and the criminal and judicial apparatus [@Harvey2019;@Harvey2022]. As a result, high levels of fraud are a sign of incumbent strength in both cooptation and repression (Gerschewski's third pillar). Both of these tools can be useful for defusing or dismantling an emergent protest movement. As a result, effective fraudsters should also be more effective at defeating protest movements. Consequently, in this view, there is little penalty for severe manipulation in terms of protest risk, and little deterrent effect.

Testing this proposition in more depth is difficult. If, for example, incumbents manipulate just the right amount to avoid mass protest in equilibrium\textemdash and are deterred from manipulating more by that risk\textemdash we would observe no statistical relationship between the severity of fraud and the likelihood of protest initiation. High and low levels of fraud would generally be associated with no protest. Beginning to untangle the relationship between election fraud and protest requires looking beyond standard binary measures of protest. Instead, this paper tests outcomes "off the equilibrium path," to use the language of formal modeling. Does greater risk to the incumbent in the pre-election period result in reduced election fraud? When protest movements do arise after an election, is fraud associated with weaker governments? Using data from 614 non-democratic elections across 97 countries, from 1989 to 2012, this paper finds that the answer to both questions is no. In fact, greater pre-election risk is associated with worse fraud, and more severe election fraud is associated with a *faster* demise for protest waves. This indicates that, far from being a sign of weakness, in most cases the ability to generate significant electoral fraud is an indicator of durable regime strength. This finding holds even when incumbents claim small vote-shares, suggesting that electoral fraud capacity is an independent component of regime stability.

Given these results, it is likely that the downside risk for incumbents that engage in election fraud in non-democracies has been overemphasized. Fraud itself is a sign of the robustness of some of the same tools that can be used to co-opt the opposition, constrain elites, or intensify repression during a crisis; far from being a sign of weakness, it can be a sign of resilience. This can deter protest, especially in conjunction with high claimed vote-shares for the incumbent. Even in the unlikely event that protests do break out in response to large-scale fraud, a government that has engaged in widespread fraud has many tools it may deploy to curtail them.

These results help explain several empirical patterns at once: the relative paucity of electoral protests, the rarity of their success in overturning incumbent regimes, and governments' attempts to claim implausibly high margins of victory in conjunction with high levels of fraud. At a theoretical level, they also suggest that protest risk should be de-emphasized in research on the causes of election manipulation, in favor of alternative explanations which emphasize the role of front-line actors in delivering manipulation. Finally, they suggest a pessimistic outlook for democratization by election in authoritarian regimes.

# Literature review: election manipulation and mass protest

Prior research has often considered that electoral manipulation has both costs and benefits for ruling parties. The benefits include better odds of winning the official tally [@lehoucq2002], of course, as well as any signals of strength and stability that a larger vote-share [@magaloni2006] or display of power might provide [@simpser2013;@gehlbach2015electoral]. On the other side, election manipulation is generally characterized as having both operational costs (which the incumbent may be unwilling or unable to pay [@Greene2010]) and legitimacy costs which reduce other actors' level of cooperation with the incumbent [@birch2011]. The most visible, and arguably most studied, form of such legitimacy costs is mass post-election protest against the incumbent.

The cost of protest is conceptualized as the risk that the incumbent may be unseated and/or the resource cost needed to repress or buy off the incumbent's opponents. Various theories model the decision to engage in post-election protest either as a strategic one by opposition groups [@magaloni2010;@Luo2020;@przeworski2015], or more spontaneous coordination by citizens expressing a shared grievance [@tucker2007;@Brancati2016]. In either case, the risk of protest is a classic component of models of electoral manipulation, as a restraint on incumbents who would prefer to use fraud to increase their vote-shares, all else held equal. This argument can be found in theories about the conduct of non-democratic elections, as well as in the literature on self-enforcing democracy [@fearon_self-enforcing_2011;@przeworskiSelfenforcingDemocracy2006]. In the latter literature, generally free and fair elections are thought to be a self-enforcing equilibrium through the threat of mass protest by wronged groups if the incumbent government rigs the results. While the focus here is on elections in authoritarian cases, some implications for self-enforcing democracy are discussed at the close of the paper.

<!-- Broadly, prior work has considered two major explanations for the severity of manipulated elections: the costs and benefits of manipulation to the ruling party and government (a demand-side framework) and the costs and benefits for the front-line actors who do the work of manipulating elections (a supply-side framework). The risk of costly protest is the central, though not sole^[In some studies, like [@birch2011], the costs of manipulation are attributed to a more diffuse loss of legitimacy; such costs *may* manifest as protest, but may also manifest as a reduction in general cooperation with the regime. However, it is worth noting the argument of [@gehlbach2015electoral] that election manipulation can *induce* cooperative behavior from some agents.], factor which affects the demand-side cost of manipulation in most models. -->

<!-- In the major formulations of this argument, such as Fearon  [-@fearon_self-enforcing_2011] and Przeworski [-@przeworskiSelfenforcingDemocracy2006], 'democracy' is understood in procedural terms. For Przeworski, democracy requires that "authorization to rule is derived from elections" which designate winners and losers according to a set of rules [@przeworskiSelfenforcingDemocracy2006, p. 312]. Fearon adopts a similar definition for electoral democracy [@fearon_self-enforcing_2011, p. 1662]. Such arrangements are not limited to liberal democracies, and in fact constitute the large majority of regimes today. Many previous works argue that the risk of protest can mitigate election manipulation in regimes where elections are held\textemdash including established democracies [@millerRepublicIfYou2021], electoral authoritarian regimes [@Egorov2021], and any regime where elections are held but may be manipulated [@Little2015;@Luo2020]. In most such models, protests are successful in overthrowing the government with some probability, but impose a cost on both the protesters and the government regardless of success (e.g. [@Luo2020]); the logic thus extends even to highly repressive and entrenched regimes. Nevertheless, this body of work has yet to grapple with the empirical finding that election integrity is relatively poor across these regimes outside of a narrow club of 'consolidated' democracies.  -->

Much of the work on this subject has been done in the formal modeling tradition. In game-theoretic research where the risk of protest is shown to deter fraud, the election is usually conceived as a signaling game. In such models, incumbents are characterized as 'weak' or 'strong' based on their support in the electorate; weak incumbents may need to resort to fraud in order to secure re-election. As a result, the opposition perceives election fraud as a reliable signal of incumbent weakness, and will stage protests in order to push out the weak incumbent. For example, the model developed by Little, Tucker, and LaGatta [-@Little2015] predicts that "Just as stronger election results for the incumbent lead to less protest because they indicate that she is popular, reports of more fraud lead to more protest because they indicate that the incumbent is, for a given election result, less popular" (p. 1153). In equilibrium, this can lead weaker incumbents to step down rather than face protest due to massive fraud\textemdash a pro-democracy outcome. Similar results are developed by Chernykh and Svolik [-@chernykh2015third], Luo and Rozenas [-@Luo2020] and [-@Luo2017], and Magaloni [-@magaloni2010]. These are normatively optimistic models upon which a sizable literature has been founded: fraud can be kept in check, and ultimately reduced, through the development institutions that reveal information about election integrity: election monitoring [@Hyde2014;@daxecker2012], more independent courts [@chernykh2015third], media activity [@naiFourthEstate], and civil society coordination [@birch2017getting].

Another school of thought holds that electoral protest is driven by citizens' grievances rather than by a strategic, organized opposition [@Brancati2016]. In this view, election manipulation can act as a shared grievance among individuals who prefer to see the ruling party lose\textemdash especially when it appears that fraud may have been decisive [@kuntz_more_2009]. Belief in a stolen election is a particularly motivating kind of grievance, because it acts to focus all the diverse personal grievances individuals may have against the regime into one moment of opportunity for action [@tucker2007], possibly initiating a protest cascade [@Kuran1991]. Just as above, these models tend to predict greater protest risk when margins of victory are low, suggesting a stolen election, and/or grievances are high in the presence of fraud.

Empirical work has found some support for the microfoundations of this theory. For example, ballot-stuffing and voter coercion have been shown to generate stronger feelings of anger than pre-election manipulation of ballot access [@Szakonyi2021], and access to information about manipulation appears to reduce support for the ruling party [@Reuter2021], increase support for protest and reduce the perceived legitimacy of the regime [@Williamson2021]. This argument has found support in studies of the relationship between economic grievances, manipulated elections, and protest [@Rod2019;@Brancati2016].

The grievance-based approach suggests that protests hinge more on citizens' subjective beliefs about fraud and the broader social context than on the actual level of malfeasance. A study of individuals in Nigeria found that individuals' subjective perceptions of fraud makes them more likely to participate in protest, not more objective measures [@Daxecker2019b]. Subjective belief in fraud was also shown to be correlated with approval of mass protests after Russia's 2011 election [@Chaisty2013]. Furthermore, simply being on the losing side in an election is associated with reduced belief that the election was fair [@Cantu2015], and with increased support for electoral protest [@Anderson2006]. These studies imply only a limited deterrent effect of protest on fraud in non-democracies. If the incumbent may pay the cost of protest whether or not it engages in fraud (due to subjective, and perhaps mistaken, beliefs about fraud and economic conditions), it may as well engage in manipulation in order to gain the benefits.

This implication connects to a second class of formal models, which hold that fraud operates as a distortion of the signal of the incumbent's true strength. In this type of model, such as those developed by Little [-@littleFraudMonitoringNoncompetitive2015] or Egorov and Sonin [-@Egorov2021], fraud may be used by incumbents to bluff their opponents into being more supportive than they would be if they knew the incumbent's true popularity. However, greater transparency makes such bluffs more costly. The typical result in these models is that protest risk can *partially* deter fraud. In equilibrium, the incumbent will generally choose the level of fraud anticipated by the opposition and no more\textemdash but also no less.

Finally, there is a third school of thought, which holds that election manipulation may indicate regime strength, especially if manipulation is blatant or produces large margins of victory [@simpser2013]. In the next section, I expand on this school of thought in more detail.

There is a general, though not complete, consensus in empirical work that election fraud is associated with increased risk of protest. Several cross-national studies have found links between election manipulation and binary measures of protest [@Brancati2016;@Hyde2014;@Harvey2020;@Rod2019]. A study of African elections from 1997 to 2009 shows that election fraud is associated with an increased number of post-election conflicts [@daxecker2012]. However, Hafner-Burton et al [-@Hafner-Burton2016] find that the positive association between election fraud and post-election protest falls short of conventional statistical significance. Studies of one archetypal example, the 2011 election and protest cycle in Russia, have found a positive relationship between fraud and protest by some measures [@Lankina2017], and no relationship by others [@ananyevDictatorsSignalStrength2022]. Taking an alternative approach, Rozenas [-@Rozenas2016b] finds that governments facing pre-election protests manipulate less severely; however, the dependent variables for that study are largely defined by the structural conditions of the election (i.e. whether opponents are allowed to run), rather than by the election-day manipulation that usually figures in theoretical work. This result is consistent with the finding that parties are more likely to protest elections after ex ante changes to the rules [@chernykh2014].

While these works have identified a connection between fraud and protest, in general, they cannot distinguish between the competing assumptions of the two theories discussed here, or the third possible approach discussed in below. For instance, all three approaches predict that protest should be most likely when evidence of fraud is high and the incumbent's vote-share is low (or, relatedly, grievances are high). Such a scenario could occur due to 1) strategic protest against an incumbent revealed to be weak through reliance on fraud, 2) mass-led protest against a 'stolen' election, or 3) risky strategic protest against a low-popularity incumbent despite institutional reservoirs of strength. The implications of all three scenarios are substantively distinct, however. To untangle them, it is necessary to test the assumptions underlying the different models. In particular, do post-election protest movements encounter weak incumbents as signaling models predict?  If not, the deterrence model is undermined.

<!--  One critique of this view, although still within the information/strategic framework, comes from Simpser's [-@Simpser2011] pathbreaking book. Simpser argues that when wanipulation efforts are excessive---well beyond what is needed simply to win---they demonstrate the organizational capacity, resources, and manpower of the ruling party [@Simpser2013]. Large-scale manipulation is, thus, a signal of strength rather than weakness, and can deter protest. By contrast, partial manipulation efforts that fail to secure a large majority vote-share signals the government lacks both popular support and mobilizational capacity, making opposition protest more likely. -->

# Theory

As discussed above, one view of election fraud and protest views fraud as a tool of weak incumbents. Another view, associated with Simpser [-@simpser2013], holds that election fraud demonstrates incumbent strength. In Simpser's original account, we should expect electoral manipulation to signal strength when it is readily observable or excessive, producing large margins of victory. A strong government may manipulate heavily in order to produce a very large majority [@magaloni2006], to induce regime insiders to remain cooperative [@gehlbach2015electoral], or because of a bandwagon effect among front-line agents [@rundlett2016deliver]. This argument is in keeping with Seeberg's [-@Seeberg2019] finding that higher-capacity states are weakly associated with more election fraud.

I build on Simpser's logic, and on subsequent research, by arguing that that extensive election fraud itself requires the use of resources that are sources of incumbent strength. Simpser's model is concerned with whether or or not other actors can perceive the regime's signal of strength; hence the focus on excessive or blatant manipulation. I argue here that even when large-scale election fraud is marginal (i.e. produces narrow margins of victory), it can indicate regime strength independent of popularity. This argument  draws on the recent strand of research on election manipulation that emphasizes the principal-agent problems inherent in such campaigns. Political principals require the cooperation of large numbers of front-line agents in order to generate extensive election fraud, agents who may shirk if their perceived interests are not aligned with the incumbent [@rundlett2016deliver].

To the extent that election results in non-democracies send a signal, it is composed of a mix of three factors [@Pop-Eleches2015]: the incumbent's genuine popularity, its ability to mobilize otherwise reluctant voters (through patronage, clientelism, and coercion), and its ability to control the bureaucratic apparatus (including the criminal and judicial systems) to generate fraud. Large-scale election fraud requires that the incumbent have consolidated control over patronage resources in society; this makes cooperation with the incumbent attractive for front-line agents, who are less willing to risk losing access to the incumbent's network [@Harvey2019]. However, such efforts also require that incumbents be able to shield front-line agents from negative consequences if their illegal efforts are publicly exposed [@Harvey2022]. In other words, the incumbent must have sufficient control over the police and the courts, such that ordinary individuals expect little punishment for committing criminal election fraud on their behalf. Each of these capacities is required to commit large-scale fraud, and both are useful for disarming protest movements. To summarize, the more extensive the election-fraud effort, the more successfully the incumbent has solved the associated principal-agent problems, and the more likely it will be able to use those tools to quash protest.

The initial emergence of protest triggers a response by the government and ruling party that can take the form of repression, concessions, or toleration [@tillyMobilizationRevolution1978;@lipskyProtest1968;@Hummel2019;@raslerConcessionsRepression1996;@Bishara2015;@Frantz2014]. Concessions may be aimed at protesting parties themselves [@reuter2015;@mooreRepressionDissentSubstitution2000;@Hafner-Burton2016;@Piven1979] or the more moderate public, in order to keep them on the sidelines of the conflict [@Hummel2019]. In either case, large-scale election fraud indicates control over considerable patronage resources, which in turn indicates the capacity to engage in such payments (or withhold them from clients in exchange for passive support [@Rosenfeld2017]).

Ruling parties may also lean on repression, calling upon the security services to disperse protests, arrest leaders, and perhaps employ harsher methods including disappearance, torture, and execution [@Escriba_Folch2013;@Davenport2007]. Though such crackdowns risk provoking a backlash, they are aimed at demobilizing the opposition by increasing the costs of participation in overt dissent [@DeVogel2022]. The ability to protect regime agents from any legal consequences for carrying out orders, a requirement for successful election fraud, is also an asset when asking law enforcement agents to carry out possibly violent repression [@MAGALONI_RODRIGUEZ_2020;@stavro_police_2023;@abouharbFactoJudicialIndependence2013].

In addition, as Gehlbach and Simpser [-@gehlbach2015electoral] argue, effective electoral manipulation can induce regime agents to exert effort on behalf of the incumbent. Elite defections, which can presage real challenges for the incumbent during a crisis, are also rarer in less democratic regions, where election manipulation is likely more severe [@Reuter2019]. Just as election manipulation has been held to solve a coordination problem for regime opponents, it may also solve a coordination problem for regime agents: when successful, a fraud campaign signals to agents that other agents are likely to stick with the incumbent due to its revealed capacities.

This argument\textemdash that successful, large-scale fraud draws on the same resources a state may use to suppress protest\textemdash turns the deterrent logic on its head. Instead of signalling weakness, large-scale fraud indicates a well of resilience. Consequently, we should expect incumbent regimes to be aggressive in their use of election fraud, especially in periods of uncertainty, in order to reveal their capacity to resolve principal-agent problems and signal resilience despite low popularity or an assertive opposition.

Since the incumbents' claimed vote-share is part of the signal generated by the election, the effect of election fraud should only be considered in conjunction with the margin of victory. A large margin of victory combined with election fraud sends an unambiguous signal of ruling-party strength, with uncertainty around its components. That is, large margins of victory can be composed of a number of values of 'natural' popularity, mobilized voters, and election fraud. This makes large margins of victory difficult signals for opposition actors to interpret. As margins of victory decline, the relative value of the signal becomes clearer. A narrow margin of victory combined with extensive fraud is a clear signal of low 'natural' popularity. While this may indicate an opportune moment for opposition actors, the large extent of fraud is nevertheless itself an independent indicator of regime strength. The combination of patronage resources and control over the criminal and judicial systems required for mass fraud can be turned to coopting or repressing protest movements. This argument is the mirror-image of the one present in Trinh [-@trinhTeaLeafElections2022], who argues that when incumbents manipulate heavily, they limit the quality of information that they can extract from the election. By the same token, a heavily manipulated election also confounds the signal received by opposition actors; a strong, fraudulent result is a signal of regime stability whose exact composition is uncertain.

The argument can be summarized as follows. Incumbents will generally seek to manipulate elections as much as they can, to capitalize on both the direct and indirect benefits of electoral manipulation [@simpser2013]. To do so, they must solve principal-agent problems by consolidating control over patronage resources and shielding agents from negative consequences of their actions. The election outcome sends an ambiguous signal of the incumbents' resources\textemdash including natural popularity, resource control, and control over the legal apparatus\textemdash to possible regime opponents. Strategic opposition actors, which may include both organized groups and individuals weighing costs and benefits of participation, interpret this signal with some prior expectation of the incumbent's capacity to manipulate. However, even when the incumbent's popularity is low, its ability to engage in election fraud is indicative of resources that can be used to terminate protest movements.

This theory makes predictions for protest initiation, protest duration, and the level of fraud employed. First, if large-scale election fraud indicates strength, we should expect that incumbents will respond to pre-election signs of opposition strength by increasing the level of fraud in order to send a more forceful signal. This runs directly counter to the prediction of the classic models, which hold that a greater risk of post-election protest should deter fraud.

> H1: Pre-election environments that are more uncertain for the incumbent will be associated with increased election fraud.

Likewise, in an election where the incumbent claims victory, protest should be most likely when levels of fraud are high and claimed vote-shares are low. A low vote-share for the incumbent combined with large-scale fraud indicates that the incumbent's 'natural' popularity is very low. As claimed vote-share increases for a given level of fraud, the incumbent appears better supported and harder to challenge.^[Of course, the strongest signal of weakness would be one where vote-share and fraud are low, but in this case it is likely that the incumbent has accepted an election loss, rather than falsify a victory. This scenario is explored more in the appendix.] This hypothesis is consistent across all three schools of thought, but should nevertheless be evaluated.

> H2: The likelihood of protest initiation will be highest when incumbent vote-share and election fraud are low.

Finally, if large-scale protest is indicative of regime resiliency, we should expect that election fraud will be associated with reduced protest duration among those incumbents who successfully fend off protest movements. That is, among strong incumbents who were capable of defeating post-election protest movements, the resources associated with large-scale fraud should enable them to more swiftly quell the movement. Those who lack such resources are expected to take longer to dispatch the protest movement. Again, these expectations are counter to the standard protest-deterrent logic.

> H3: Among incumbents that survive post-election protest movements, more extensive election fraud should be associated with shorter protest waves.

# Data and methods

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
library(glmmTMB)
library(modelsummary)
library(huxtable)
library(simPH)
library(broom)
library(broom.helpers)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))

#Coding as factors

nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Replacing NA with 0 for protest size

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))

## Fixing missing values of winner.share.upd for liberal democracies

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))

## Set max_size zero for elections with no protest

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Excluding liberal democracies

nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.

## Remove outlier
nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)

## Coding median size as 0 when n.events = 0

nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))

```

These hypotheses are tested using data from three datasets: NELDA [@hyde_2012_marinov], V-Dem version 11.1 [@vdem11.1], and the Electoral Contention and Violence dataset [@Daxecker2019]. The unit of observation for this study is an election period as recorded in NELDA. This framework prevents selecting on the dependent variable, by capturing both protested and non-protested elections. I limit the sample to authoritarian regimes which hold multiparty national elections. This excludes fully closed autocracies (where no national elections are held de jure), as well as electoral and liberal democracies.^[The specific variable used is 'v2x_regime'; I limit the sample to those regime-years coded 1 for either the current or lagged year. This coding rule will include cases that are coded as electoral democratic after an election but electoral authoritarian before.]

While it is possible that the theoretical mechanisms considered here may be at work in democracies as well, democracies are excluded from this analysis for two reasons. First, the theory presented above is derived from prior work on authoritarian elections; the factors that drive the theory are expected to be weaker in democratic regimes. As regimes trend toward liberal democracy, patronage consolidation should decline and rule of law should increase by definition. Consequently, the theory expects that fraud should be rare in such cases, even in the absence of a protest deterrent: incumbents simply lack the tools. At the same time, the more liberal-democratic a state becomes, the more alternatives their are to protest as a means to check the executive. More independent courts and more active executive constraints have been shown to be associated with improved election integrity [@Harvey2022;@Bishop2016a]; it is where these factors are weakest\textemdash the electoral authoritarian regimes\textemdash where we should expect to see a more pronounced role for protest risk in curtailing election fraud under the deterrence logic. The case selection strategy here is thus one that both fits the theory and makes for a strong test of the classic fraud-protest model.

Of course, treating regime type as a sharply defined binary indicator, rather than as a continuum, will introduce some bias around edge cases. As a robustness check for this case selection process, I also run the main analyses on a larger sample that adds electoral democratic regimes. While the strength of the findings is somewhat attenuated, as is expected, the results are generally upheld in the broader sample as shown in the appendix.

Models where the dependent variable is the level of election fraud include data from all cases. For the models that focus on the initiation of post-election protest, I exclude cases where the incumbent party lost, based on coding in NELDA\textemdash this limits the observations to those that fit the population of interest: protest following an election in which the incumbent won, possibly through reliance on election fraud.^[Excluding cases where incumbents stepped down after losing an election is necessary to avoid conflating pro- and anti-incumbent protest. However, this design does influence the interpretation of the results from these models. Threats to inference from this section are evaluated in more detail in the discussion and appendix.] For models of protest-wave duration, only cases with a positive number of post-election protests are included. The date range is limited by the availability of data on protest from ECAV, as discussed below, which ranges from 1989 to 2012. The resulting dataset includes 614 total election-period observations across 97 countries.

## Dependent variables and modeling strategies

To capture the severity of *election fraud*, I use the V-Dem measure of intentional election irregularities, a category which includes "double IDs, intentional lack of voting materials, ballot-stuffing, misreporting of votes, and false collation of votes." I operationalize electoral manipulation using this variable, rather than alternative measures of election integrity, because fraud of this kind is thought to be most likely to trigger protest [@Linebarger2020]---it is often incumbents' tool of last resort [@Sjoberg2016], may not mobilize real voters [@Harvey2020], and can be more damaging to legitimacy than pre-election manipulation [@birch2011;@Szakonyi2021]. This measure of election integrity should thus serve as a strong test of the protest-oriented model. It also matches the conception of election manipulation used in much of the relevant formal literature.^[For example, election manipulation is treated as post-hoc adjustment of the results in Little et al [-@Little2015], Little [-@Little2012], Magaloni [-@magaloni2010], and Fearon [-@fearon_self-enforcing_2011], among others.] The variable is used as a dependent variable in the first set of models that test H1, and as a potential explanatory variable in the subsequent models of protest initiation and duration. Two alternative measures are used as robustness checks in the appendix.

ECAV records protests and other events related to an election, ranging from six months before to three months after the vote (or three months after the final round of a multi-round election). To qualify, an action must be contentious; rallies in support of the incumbent are excluded. I only include events labeled as protests, occupations, or blockades, excluding acts like shootings, bombings, and coups. ECAV data is in event-day format; to capture the *number of protests* before the election, I take the sum number of events recorded in ECAV before election day. To evaluate hypotheses regarding *protest duration* I take the number of days between the last protest recorded in ECAV and the first, for each election.^[The duration of a protest wave with only one protest is taken to be one day. While the main analysis assumes that the last protest in the dataset marks the end of the protest movement, an alternative specification in the appendix assumes right-censoring for a portion of the observations due to the 90-day observation window.]

The large majority of elections experience no post-election protest: in `r nelda2.sub %>% filter(n.events.post == 0) %>% nrow()` of the `r nrow(nelda2.sub)` elections, no protests are recorded. Figure \@ref(fig:elex-hist) reports the distribution of election fraud across the dataset, as well as protest duration. As the figure shows, while election fraud is common among this set of regimes, protests are rare, and, when they do occur, are usually short in duration. This distribution\textemdash with large numbers of zeroes\textemdash is in line with the theoretical proposition that most fraudulent elections do not produce signals of weakness that prompt opposition protest. Still, approximately one-third of elections experienced some protest, and so protest initiation is modeled using standard logistic regression. *Election fraud*, which follows a more normal distribution, is modeled using multilevel models with random intercepts by country. Lastly, protest wave duration is modeled using Cox proportional hazards models.

In the survival models, the sample is restricted to the 156 cases where post-election protests occurred. If significant election fraud is indicative of regime resilience, the expectation is that more fraudulent elections will be associated with more quickly dispatched protest movements. As an additional robustness check, I subsequently exclude the 7 cases where incumbents were removed following protests. This allows me to exclude the possibility that incumbents who make greater use of fraud are *weaker*, and more quickly removed from office, than their more electorally honest peers. A brief further analysis of the seven cases in which incumbents were ousted is also included in the discussion.

```{r, echo=FALSE, warning=FALSE, message=FALSE, include=FALSE}

p.hist.events <- ggplot(nelda2.sub, aes(x=n.events.post)) + geom_histogram(bins = 90) + xlab("Number of protest events") + theme_bw()

p.hist.duration <- ggplot(nelda2.sub, aes(x=d.events.num)) + geom_histogram(bins=100) + xlab("Protest wave duration") + theme_bw()

p.hist.fraud <- ggplot(nelda2.sub, aes(x = v2elirreg.inv)) + geom_histogram(bins=100) + xlab("Election fraud") + theme_bw()

 ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= jitter(n.events.post))) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("n.events.post (jittered)") + theme_bw()

#  ggplot(nelda2.sub, aes(y = v2elirreg.inv, x= max_size)) + geom_point(aes(color = factor(inc.fav.cancel.protests))) + xlab("max size") + theme_bw()
```

```{r elex-hist, fig.cap="Distributions of election fraud, post-election protest, and protest duration", fig.height=4, echo=FALSE, message=FALSE, warning = FALSE}
plot.combined.hist <- cowplot::plot_grid(p.hist.fraud, p.hist.duration, ncol=1)
plot.combined.hist
```

```{r, include=F}
incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
                                            'Legislative constraints' = legcon.1lag, 
                                            'Alternative information' = altinf.1lag,
                                          'Liberal democracy index' = libdem.1lag,
                                            'GDP growth rate' = gdpgro.1lag, 
                                            'Civil society openness' = osorg.1lag, 
                                            #'Public sector corruption' = pubseccorrup.lag,
                                            'National party organization' = natlpartyorg.lag, 
                                             'Election fraud' = v2elirreg.inv, 
                                            'Presidential election' = presidential,
                                            'Incumbent term-limited' = inc.termlimit.num,
                                            'Physical integrity index' = physinteg.1lag,
                                            'Urbanization' = urban.pct, 
                                            'Rentier state' = rentierstate,
                                            'Tax state' =  taxstate,
                                            'Winning vote-share' = winner.share.upd, 
                                            'Number of pre-election protests' = n.events.pre,
                                            'Off-schedule election' = unscheduled_election,
                                          'Regional protest diffusion' = regional_success_lag,
                                         "Privatization" =  state.owned.1lag,
                                         "Professional bureaucracy" = patronage.appts.1lag,
                                        "EMB autonomy"= v2elembaut)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))

incl.vars <- incl.vars %>% mutate(`Presidential election` = as.numeric(`Presidential election`))
incl.vars <- incl.vars %>% mutate(`Rentier state` = as.numeric(`Rentier state`))
incl.vars <- incl.vars %>% mutate(`Tax state` = as.numeric(`Tax state`))
incl.vars <- incl.vars %>% mutate(`Off-schedule election` = as.numeric(`Off-schedule election`))

```

\singlespace

```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)
datasummary_skim(incl.vars, output = "kableExtra", title = "Table 1: Summary statistics and variable distributions", histogram = F)  %>%
    kable_styling(font_size = 9, full_width = FALSE)
```
\doublespace

```{r, include=FALSE}
library(ggcorrplot)
corr <- round(cor(incl.vars, use = "pair"), 2)

ggcorrplot(corr, hc.order = TRUE, type = "lower")
```

## Explanatory variables

The first set of models test Hypothesis 1, on the relationship between protest risk and fraud. Protest risk is operationalized in three ways. First, the *number of pre-election protests* in ECAV is taken as *prima facie* evidence of a more turbulent and risky election environment. Given that pre-election protest is likely to be a strong predictor of post-election protest\textemdash by indicating strong social networks [@mateo2022], pre-existing opposition groups and civil society capacity [@Bunce2006], and/or popular grievances [@Brancati2016]\textemdash the deterrent logic suggests that incumbents will pull back on election-day fraud when the pre-election environment is characterized by protest [@Rozenas2016b]. On the other hand, H1 predicts that incumbents will generate more manipulation in this context.

Second, since a state with greater capacity for repression may be better able to deter protest and thus 'get away with' more manipulation [@Egorov2021,@Ritter2016], I measure repressive capacity using the *physical integrity* index from V-Dem, lagged one year. Higher values for this variable indicate greater freedom from torture and political killing. In keeping with the 'murder in the middle' framework [@feinLifeIntegrityViolationsDemocracy1995], the highest protest risks (and thus most reduced fraud) might be found at middle values of repression; regimes with the most demonstrated repressive capacity deter potential challengers from taking to the streets in the first place, and the least repressive states defuse protest through accommodation [@pierskallaProtestDeterrenceEscalation2010;@mooreRepressionDissentSubstitution2000]. I thus include the variable's square term to account for the 'murder in the middle' framework.

Finally, the literature on the diffusion of electoral protest holds that successful electoral protest in one country increases the likelihood of electoral protest in similar countries nearby [@Bunce2006a] (see Brancati and Lucardi [-@Brancati2019] for an alternative view, however). This diffusion mechanism is modeled using each country's geopolitical region as coded in Teorell et al [-@qog_teorell] in conjunction with binary variables indicating a successful protest movement from NELDA. A 'successful protest' is recorded if a protest movement resulted in either a new election being run or the incumbent stepping down. The *regional diffusion* dummy variable takes on a value of one for all countries in the same region as a successful protest, for elections within one calendar year of the successfully overturned election.^[For example, the variable takes on a positive value for elections held in the post-communist region, within one year of the September 24, 2000 election in Serbia that sparked the 'Bulldozer Revolution'. All told there are 15 successful electoral protests in the dataset: Bulgaria 1990, Lesotho 1998, Cote d'Ivoire 2000 and 2010, Peru 2000, Serbia 2000, Georgia 2003, Ukraine 2003-2004, Azerbaijan 2005, Kyrgyzstan 2005, Haiti 2006, and Cameroon 2007.]

The models incorporate several potential explanatory variables, in addition to the *election fraud* variable described above. The *lagged GDP growth rate*, taken from V-Dem, is used to capture economic grievances: lower GDP growth rates, especially negative ones, should be expected to increase citizens' sense of grievance. The winning party's *vote-share* is also taken from the V-Dem dataset.\footnote{This variable has considerable missingness, which I reduced in part by updating missing values in V-Dem using data from Wikipedia and Psephos.} Finally, whether or not the chief executive is *term limited* is taken from NELDA.

## Control variables

As identified in prior research, other factors that may influence both fraud and protest include the previous repressive history of the government [@Davenport2019], the institutionalization of opposition parties [@trejo_ballot_2014], indicators of government *control over civil society* (the 'v2cseeorgs' variable from V-Dem), the proportion of national political parties with *permanent party organizations* (V-Dem's 'v2psorgs'), *judicial constraints* on the executive ('v2x_jucon') and the available of *alternative sources of information*. Since urban areas can make electoral manipulation more difficult [@ziblatt2009;@larreguy2016parties] while also making protest more likely [@Wallace2013a], I also include data on *urbanization* taken from the United Nations [@DepartmentofEconomicandSocialAffairsPopulationDivision2019]. All models include *legislative constraints* on the executive, a V-Dem variable which captures the extent to which the legislative body in the country acts independently in practice to limit executive power. Additional V-Dem variables capture *privatization* of the economy ('v2clstown') and *bureaucratic professionalization* ('v2stcritrecadm'), both of which might give incumbents influence over election fraud and subsequent protest. Lastly, a V-Dem measure the *autonomy of election managment bodies* ('v2elembaut') is included.

Several binary control variables are also included. First, a variable from NELDA is used to indicate *presidential* elections, since these may include both higher levels of manipulation [@simpser2013] and a greater risk of protest. Another NELDA variable indicates if the incumbent executive is not running due to *term limits*. I include dummy variables that provide a categorical evaluation of state capacity. These are taken from the ordinal version of the V-Dem variable "state source of fiscal revenue" (v2stfisccap). If the V-Dem data codes a state as not capable of raising revenue or primarily relying on loans and foreign aid, I code it as a *weak state*. If the state is described as primarily relying on resource rents and expropriation, I code it is a *rentier state*. States that rely primarily on taxes on property or economic transactions are coded as *tax states*. This distinction is relevant since access to natural resource rents may both make protest less likely and repression more effective [@Girod2018]. Lastly, since strong non-democratic governments may be able to adjust the electoral calendar strategically (to affecting both election manipulation and protest risk), I include a variable from NELDA that indicates *off-schedule elections*.

Table \@ref(tab:summary-table) shows summary statistics for all of these variables. To address concerns about multicollinearity, I present models in the appendix that exclude control variables associated with pre-election protest.

\newpage

\singlespace

```{r manipdv, echo=FALSE}

m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag +  v2elembaut + (1 | stateid), data = nelda2.sub) 
#summary(m1a)

m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag  + unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)

#summary(m1b)

m1c <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate  + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1c)

m1 <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate + n.events.pre  + state.owned.1lag + patronage.appts.1lag + unscheduled_election  + v2elembaut +  (1 | stateid), data = nelda2.sub)

#summary(m1)

m1a <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   +  presidential +  urban.pct + rentierstate + taxstate +  unscheduled_election + state.owned.1lag + patronage.appts.1lag  + regional_success_lag + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1a)

m1b <- lmer(v2elirreg.inv ~ jucon.1lag + legcon.1lag  + gdpgro.1lag +  osorg.1lag  + natlpartyorg.lag   + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + state.owned.1lag + patronage.appts.1lag +  unscheduled_election + v2elembaut +  (1 | stateid), data = nelda2.sub)

#summary(m1b)

m1d <- lmer(v2elirreg.inv ~  libdem.1lag + gdpgro.1lag +    presidential +  urban.pct + rentierstate + taxstate  +  unscheduled_election + n.events.pre  + state.owned.1lag + patronage.appts.1lag + v2elembaut +  (1 | stateid), data = nelda2.sub) 
#summary(m1d)

modelsummary(c(m1, m1a, m1c), output = "kableExtra",  stars = T, gof_omit = c("R2|RMSE"), title = 'Table 2: Random-intercept multilevel model of election-day fraud', coef_omit = "SD", coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "state.owned.1lag" = "Privatization",
                 "patronage.appts.1lag" = "Professional bureaucracy",
                 "v2elembaut" = "EMB autonomy"
                 )) %>% kable_styling(latex_options = "HOLD_position")

##Note: state owned variable is inverse coded, so should be labeled as private ownership in economy
##Note: Patronage appointments ialso inverse coded. Higher values = more professionalism
```

\newpage

\doublespace

# Results

## Election-day fraud and protest risk

The results of the first set of multilevel models are not promising for argument that protest risk deters election manipulation. Hypothesis 1 is supported in the models shown in Table  \@ref(tab:manipdv). Models 1 through 3 apply different measures of pre-election protest risk; all of them behave contrary to the standard expectation in the literature. In Model 1, the number of pre-election protests is associated with an increase in election-day fraud. While the effect is small\textemdash ten pre-election protests only tend to increase election fraud by about six percent of a standard deviation\textemdash it runs counter to the expectations of the deterrence logic. A similar effect is found for *diffusion* in Model 2; recent successful protest in a nearby country is also associated with a modest increase in expected fraud. This suggests that the level of election manipulation, rather than being deterred by a highly contentious environment, may be deployed to make up for anticipated poor election results and/or to signal organizational strength.

Model 3 shows that fraud first increases with physical integrity, but then declines. In other words, fraud is lowest on the extremes of repressive capacity\textemdash where repression is severe and where it is most limited. This is contrary to the expectation that a higher risk of protest is associated with reduced fraud.  Under the expectation that protest risk is highest for these middle values [@pierskallaProtestDeterrenceEscalation2010], H1 is supported by each measure; fraud is higher when protest risk is higher. Models with fixed effects by country, in the appendix, show that these findings holds even within cases.

## Fraud and the likelihood of protest

\singlespace

```{r include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.margin.binary <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary)

model.margin.binary.int <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary.int)

model.margin.binary_noc <- glm(any_protest ~ jucon.1lag + legcon.1lag +  altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
summary(model.margin.binary_noc)

## Plots

####Margin

#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

```

```{r logit-table, echo=FALSE}
models2 <- list()
models2[['Model 4']] <- model.margin.binary
models2[['Model 5']] <- model.margin.binary.int
#models2[['Model 6']] <- model.margin.binary_noc

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state", "v2elirreg.inv" = "Election fraud",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

\newpage
\doublespace

Table \@ref(tab:logit-table) shows the results for Hypothesis 2, which is expected to be supported across all three schools of thought. Model 4 is a base model, where incumbent vote-share and *intentional irregularities* are standalone variables; Model 5 includes their interaction. The expected, significant interaction effect is detected in Model 5, and it is in the expected direction: the occurrence of any protest is most likely when incumbent vote-share is low and fraud is high.

Figure \@ref(fig:margin-prob-fig) illustrates just how tenuous this effect is, however. The figure shows the predicted probability of post-election protest for two specific values of election fraud, across all values of incumbent vote-share, holding all other variables constant at their means (continuous variables) or modes (categorical variables). The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively. The figure highlights that the predicted probability of protest is only statistically significant for these unusual values of fraud at extremely low levels of incumbent vote-share. This represents an exceedingly unlikely outcome along both variables. As the figure shows, there are only a small number of observations in the range where the predicted probabilities are statistically significant.

While all three schools of thought are consistent with H2, the results here may be taken as circumstantial evidence in favor of the fraud-as-strength model. Model 4 shows that fraud alone has no bearing on protest occurrence; while the level of fraud only appears significant when the incumbent's result in the election is very low. This is suggestive of the model where opposition actors interpret the result as showing extremely low genuine support for the incumbent, perhaps enough that it outweighs the signal of strength sent by fraud. Middle values for vote-share combined with high values of fraud suggest durable regime strength, bringing expected protest close to zero. In the alternative view where fraud conveys weakness, we might expect the probability of protest to rise more sharply as incumbent vote-share declines, and to reach higher predicted probabilities. This pattern is only suggestive, of course, and for more definitive evidence I look to protest duration.

```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
plot_binary_margin_typ_n
```

## Protest wave duration

```{r, include=FALSE}
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))

cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
summary(cox1)

cox1.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0))
summary(cox1.cen)

#cox1t2 <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election +  regional_success_lag + n.events.post + v2elirreg.inv + tt(n.events.post), data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0), tt = function(x, t, ...)  x* nsk(t, knots=c(10, 70, 90),
#Boundary.knots = FALSE))
#summary(cox1t2)

cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]")) + coord_cartesian(ylim = c(0, 10)) #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

cox2.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
summary(cox2.cen)

cox3.cen <- coxph(Surv(d.events.num, surv_status_censor_assume) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post + median.participants, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
summary(cox3.cen)

```

Table \@ref(tab:cox-models-tab) reports the results of Cox proportional hazard models of protest duration. In such models, coefficients with a positive sign indicate factors that are associated with shorter-lived protests; those with a negative sign indicate factors associated with a slower demise. The coefficient for *election fraud* is positive and significant in Model 7, indicating that more fraudulent elections are associated with less resilient protest movements. Exponentiation of the coefficient for election fraud gives a hazard ratio of 3.3, indicating that for each one-unit increase in fraud, a protest movement becomes three times more likely to end at any given time. This supports H3.

One threat to inference in this design derives from the 90-day post-election window considered by ECAV. It is possible that some protest movements considered 'dead' in Models 7 and 8 are, in fact, 'alive,' if subsequent protests occurred after 90 days and are unobserved. As a test of the robustness of the conclusions, in Models 9 and 10, observations are assumed to be right-censored if a protest occurred within three-weeks of the 90-day cutoff. This amounts to 15 of the 156 observations in the sample. The results remain consistent with those from the uncensored models. While the coefficient on *election fraud* falls slightly short of conventional statistical significance in Model 9, it retains the same size and sign compared to Model 7. The interaction effect in Model 10 increases in size, while retaining its direction and statistical significance. This suggests that the main models are not drawing overly incorrect inferences through censoring of longer-lasting protest movements. Model 11 adds an additional control, the median number of participants at events in the protest wave. Larger protests might indicate relative opposition strength, however, since there is considerable missingness in the variable, I use it as a robustness check. Including this control variable reduces the size of the interaction coefficient, but it retains its direction and statistical significance.

Further investigation of the relationship between fraud and protest duration yields additional insights. In Model 8, incumbent vote-share is interacted with *election fraud*. Figure \@ref(fig:margin-dur-fig) shows the marginal effect. As shown in the figure, for all incumbent vote-shares near and below fifty percent, large-scale fraud significantly increases the likelihood that a protest movement will end. This result also supports the fraud-as-strength model, and shows that only as vote-share increases above fifty percent\textemdash creating a stronger signal of popularity\textemdash does a government in a low-fraud election have the same likelihood of defeating a protest movement at a given time.

```{r margin-dur-fig, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
p.duration.margin
```

The relationship between the level of fraud and the survival of protest movements can also be demonstrated non-parametrically using Kaplan-Meier curves for varying levels of fraud. In Figure \@ref(fig:km-fig), observations are divided into three groups. High fraud observations are those where *election fraud* is one standard deviation above the mean or more, low fraud observations are those one standard deviation below the mean or less, with moderate levels in between. The figure shows a clear difference between the low-fraud group and the other two. Low-fraud cases are much more likely to see their protest movements survive past the first day than their counterparts. For much of the range of the data, they remain roughly twice as likely to survive at any given date than protest movements in medium- or high-fraud cases. This is evidence in favor of the idea that high levels of fraud is an indication of regime resilience, and contradicts the claim that fraud indicates weakness or galvanizes political opposition, as it is the low-fraud cases where governments struggle to put an end to protest movements on average.

```{r km-fig, fig.cap="Kaplan-Meier curves for protest duration across levels of fraud", fig.height= 5, echo=F, message=F, fig.pos='H'}
nelda2.sub <- nelda2.sub %>% mutate(fraud_level = ifelse(v2elirreg.inv >= (.98 + .97), "High",
                                                            ifelse(v2elirreg.inv <= (.98 - .97), "Low", "Moderate")))
nelda2.sub.km <- nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0)

p.km <- survfit2(Surv(d.events.num, surv_status) ~ fraud_level, data = nelda2.sub.km %>% filter(d.events.num < 300)) %>% 
  ggsurvfit(linetype_aes = T) +
#   add_confidence_interval() +
  labs(
    x = "Days",
    y = "Overall survival probability"
  )

p.km
```

```{r cox-models-tab, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 7']] <- cox1
models3[['Model 8']] <- cox2
models3[['Model 9']] <- cox1.cen
models3[['Model 10']] <- cox2.cen
models3[['Model 11']] <- cox3.cen

modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Table 4: Cox proportional hazard models of protest duration', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", 
                 "winner.share.upd" = "Incumbent vote-share",
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.post" = "N. of post-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud"
                 ))

```

\doublespace

# Discussion

The results can be summarized as follows. Pre-existing risk to incumbents is positively, rather than negatively, associated with election fraud. The severity of election fraud itself is not associated with protest initiation, except when combined with very low vote-shares for the incumbent. Finally, incumbents who have employed large-scale fraud appear better able to put an end to those protest movements that do emerge, compared to those who have presided over cleaner elections. This is true even when incumbent vote-shares are low. Altogether, the findings are a substantial challenge to several prior assumptions and findings in the literature on election fraud, non-democratic elections, and authoritarian stability.

The results strongly contradict the assumptions of models which hold that election fraud indicates regime weakness. Instead, across the three tests here, fraud appears to be an indicator of regime resilience. This is a substantively important distinction. If fraud indicates weakness, incumbents that would lose fair elections can be deterred from stealing the election by the risk of post-election protest. Such models predict that a democratic exchange of power can occur in such scenarios, backed up by the risk of protest. The results here would appear to cast doubt on those predictions.

These results are most consistent with the view that even election fraud can be an indicator of regime strength. Where this paper goes beyond the traditional formulation of this view\textemdash apart from testing the underlying assumptions\textemdash is the observation that large-scale fraud can be indicative of incumbent strength even when it is not excessive. That is, even when ruling party vote-shares are low (perhaps especially when this is so), incumbents which showed themselves capable of large-scale fraud are more capable of ending protest movements than low-fraud incumbents. This holds true despite controlling for indicators of other possible explanations of the relationship between low fraud and long-lived protest: civil society strength, media openness, and median protest size.

A very brief consideration of the seven cases of successful electoral protest\textemdash Ukraine (2004), Georgia (2003), Kyrgyzstan (2005), Serbia (2000), Haiti (2006) and Cote d'Ivoire (2000 and 2010)\textemdash in comparison with the other cases can also help aid interpretation of the results. Table ... [CONSIDER ADDING TABLE] shows the average values for several key variables for three groups: incumbents defeated by protest, incumbents who survived protest, and cases where there was no electoral protest. The table shows that defeated incumbents did not hold elections that were noticeably more fraudulent than those in where incumbents survived protest: 1.13 for the survivors to 1.15 for the defeated. However, those that were defeated claimed substantially smaller vote-shares: 42 percent compared to 55 percent for the survivors. Meanwhile, the unprotested group average 55 percent vote-share and a fraud score of 0.87. This supports the view that (relatively) low fraud and high vote-shares are a reliable signal of strength (preventing protest), while high levels of fraud and low incumbent vote-shares indicate that the opposition's popularity might overcome the incumbent's structural advantages in a struggle in the street. This course of action is uncertain, however, as the unsuccessful protest group shows. Lastly,

What does all this tell us? First, these results may inform how researchers think about the causes of election manipulation. They suggest that factors like the opposition's capacity to organize street protest in limiting election fraud have been overestimated. In fact, these results suggest that incumbents who have the resources will respond to such a scenario by *increasing* their production of fraud. Consequently, theories of electoral manipulation that focus on factors other than protest risk\textemdash such as resource availability [@greene_why_2007], principal-agent problems [@rundlett2016deliver], and legal risks to manipulators [@Harvey2022]\textemdash may be more important for understanding election integrity, which may drive future research as well as policy interventions. Additionally, it may change some current understandings of the role of election monitors, courts, and other third parties in upholding election integrity. As noted earlier, prior studies argue that these actors restrain manipulation by revealing information about fraud and intensifying protest risk. But if protest risk is largely untethered from the severity of fraud, it is likely that other mechanisms are at work. Scholarly attention might turn more profitably toward other explanations for manipulation, such as the organizational capacity of the state [@simpser2013], the popularity of the government [@rundlett2016deliver], and the risks faced by front-line agents [@Harvey2022].

In turn, this revision to our understanding of election manipulation has implications for broader topics like authoritarian persistence and democratization. We observe that protest initiation is most likely when fraud is high and incumbent vote-share is low. This could be indicative of strategic behavior by the rational opposition actors found in most formal models, or by less strategic behavior by citizens en masse. For such citizens, election manipulation may act as a shared grievance, especially when it appears that fraud may have been decisive [@kuntz_more_2009;@tucker2007]. A close election marked by fraud intensifies the connection between emotional and strategic protest incentives. As Schedler writes, "Grievances provide moral justifications for collective protest: we need to do something. Opportunities provide strategic justifications: we can achieve something" [@Schedler2013,p.304]. Yet this highlights a dilemma for opposition supporters, since it is exactly these scenarios wherein post-election protest is most likely to fail quickly.

There is a long-standing debate over the role of elections in democratization, of which the manipulation-protest nexus is a component [@Schedler2013;@Morgenbesser2019;@edgellWhenWhereElections2018]. At the core, much of this debate revolves around the extent to which non-democratic elections affect the ruling party's cost of suppressing the opposition versus tolerating it [@dahlPolyarchyParticipationOpposition1971]. Election manipulation, including fraud, directly reduces the cost of toleration, by making it less likely that opposition groups win elections. To the extent that it makes electoral protest more likely and/or more costly, however, it increases the cost of suppression. One view holds that election fraud has the potential to lay the groundwork for democratization by incentivizing the ruling party to tolerate organized opposition which, one day, it may be unable to repress after an election. This logic can be found in studies of the color revolutions [@bunce2010], and in the observation that competitive authoritarian regimes are more likely to democratize after an election than hegemonic ones [@donno2013elections]

By showing that election fraud does not make protest more likely (most of the time) and that the infrastructure of fraud can actually reduce the cost of post-election protest to the regime, this result flips that logic on its head. Instead, election fraud can both reduce the cost of toleration *and* the cost of suppression. This implies that electoral authoritarianism is something of a cul-de-sac: the capacity to generate severe election fraud makes it appealing for closed authoritarian regimes to introduce elections, while the manageable risk of protest makes it unappealing to democratize. This logic also creates incentives for ruling parties in weakly consolidated democracies to take measures that will enable them to engage in electoral manipulation; namely, the consolidation of patronage resources and erosion of rule of law. This provides additional theoretical support for understanding what some have called a 'wave of autocratization' or a 'democratic recession' in recent years [@luhrmannThirdWaveAutocratization2019;@diamondDemocraticRegressionComparative2021]. If the same tools that produce broad election fraud also reduce the cost of protest, then electoral authoritarianism exerts a sort of gravitational pull as a regime type.

This paper is primarily concerned with whether or not election manipulation is indicative of regime weakness or resilience; its primary evidence is finding that fraud appears to increase with pre-election uncertainty and that fraud is associated with more resilient incumbents in the face of protest. Due to space constraints, two other predictions are not tested in depth here. First, one of the key predictions of the protest-deterrent approach is that weaker incumbents will generate little fraud and voluntarily step down after losing elections. There is indeed less election fraud observed in those election-years: the mean estimate of *election fraud* in the 129 observations where incumbents lost is 0.322, much lower than the 1.20 average in the group where incumbents claimed victory. Were these incumbents deterred from committing worse fraud? Or are they simply systematically more limited in their ability to generate election fraud?

```{r, include=F, warning=FALSE, message=FALSE, echo=FALSE}
## Anova
anova_dat <- nelda2.sub %>% dplyr::select(inc.lose, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut)

# boxplots and t-tests for the 4 variables at once
for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
  boxplot(anova_dat[, i] ~ anova_dat$inc.lose, # draw boxplots by group
    ylab = names(anova_dat[i]), # rename y-axis with variable's name
    xlab = "Incumbent loss"
  )
  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
  print(t.test(anova_dat[, i] ~ anova_dat$inc.lose)) # print results of t-test
}

aov_output <- aov(state.owned.1lag ~ inc.lose, data = nelda2.sub)
summary(aov_output)
```

```{r anova-table, echo=FALSE, warning=FALSE, message=FALSE}
p.values <- c(3.009e-15, 6.014e-11, 6.149e-10, 0.01146,  3.758e-07,  0.6281, 0.003519,
              0.1852, 0.007405, 0.003086, 2.2e-16)
win_lose_diff <- c(.87, -0.15, -0.16, .06, -.47, .04, 5.3, -1.46, -.265, -.257, -.893)
variables <- c("Election fraud", "Judicial constraints", "Legislative constraints", "GDP growth (lag)", "Civil society openness (lag)", "Natl. party institutionalization (lag)", "Urban",
               "N. pre-election protests", "Privatization", "Professional bureaucracy", "EMB autonomy")
anova_dat <- data.frame(cbind(variables,win_lose_diff, p.values))
anova_dat <- anova_dat %>% mutate(win_lose_diff = as.numeric(win_lose_diff))
anova_dat <- anova_dat %>% mutate(p.values = as.numeric(p.values))
anova_dat <- anova_dat %>% rename(Variables = variables)
anova_dat <- anova_dat %>% rename("Win-Loss Difference" = win_lose_diff)
anova_dat <- anova_dat %>% rename("P-value" = p.values)

kable(anova_dat, booktabs = T, digits = 3, caption = "Table 5: Mean variable values for winning incumbents minus losing incumbents, with ANOVA p-values")  %>%
    kable_styling(font_size = 9, full_width = FALSE)

```

The overall pattern in the data leans toward the latter interpretation. Table \@ref(tab:anova-table) shows the mean differences between the two groups, with ANOVA p-values for those differences. Negative values indicate that the losing incumbents had higher values for the variable, on average. For instance, among the cases where the incumbent lost, judicial independence, professionalization of the bureaucracy, and the autonomy of election management bodies are all markedly higher. Additionally, the losing group shows lower average economic growth. All of these should tend to reduce the availability of fraud. Of the factors considered here, only *number of pre-election protests* clearly behaves as the deterrence logic would predict, with the losing group subject to one more protest, on average, than the winning group. However, unlike the factors which point toward the alternative hypothesis, this difference is not statistically significant. Thus, while future research might test this aspect of the deterrence model more rigorously, the general pattern is supportive of the view put forth in this paper. Further checks are performed in the appendix, including models which show that predicted levels of fraud based on structural factors alone are strongly predictive of incumbent defeat, while pre-election measures of risk are not.

The second question this paper does not closely examine is the role of election fraud in sustaining (or not) the small number of incumbents who are defeated by mass protest. These cases are: Peru 2000, Cote d'Ivoire 2000, Georgia 2003, Malaysia 2004, Ukraine 2004, Azerbaijan 2005, and Kyrgyzstan 2005. The causes of the post-communist cases, which account for most of the seven cases, has been much debated [@LucanWay2008;@Bunce2009]. Yet for most of these cases, election fraud either held steady after the 'successful' revolution, or increased before partially declining again some years later. These cycles, identified as a feature of 'patronal politics' [@hale_patronal_2014] could indicate that in these cases, fraud patterns and protest were caused by more fragmented elites who failed to resolve the principal-agent problems needed to engage in mass protest and control the streets. A thorough qualitative comparative analysis of the cases would be informative, but a sketch can be drawn that supports the theory here.

Ukraine's Orange Revolution, for example, combined very high fraud with a very low incumbent vote-share. The analysis of residuals conducted in the appendix finds that fraud in the 2004 election far exceeded expectations based on structural factors. Combined with a term-limited incumbent, a strategic opposition may have interpreted these signals as indicating a deeply unpopular ruling party which would struggle to put down a determined protest movement due to a fragmented elite. On the other hand, the 2005 Tulip Revolution in Kyrgyzstan came on the heels of moderate election fraud that fell well below structural expectations (i.e. large negative residuals). This struggle to generate election fraud may have been interpreted as a sign of a weak ruling party, leading to successful protest.

Nevertheless, electoral protest is rare in non-democracies, and successful protest extremely so. The overall results indicate that we should be cautious when building and interpreting theories that emphasize the risk of mass protest as a trip-wire to be avoided in non-democracies, and add empirical evidence to Przeworski's [-@przeworskiFormalModelsAuthoritarian2022] critique of such models. The rarity of post-election protest, especially successful protest, suggests that such events are highly contingent; they hinge on a multitude of factors that go beyond the regime's efforts to bias election outcomes. Both protest and election fraud are complex phenomena with multiple drivers, including resource costs, signaling, information-gathering, principal-agent dynamics, and more; these findings indicate that protest risk is not a practical constraint on fraud under most circumstances.

Finally, this paper does not examine the role of protest risk in insulating established democracies from democratic backsliding [@fearon_self-enforcing_2011]. This question deserves further investigation, but it is possible to draw a few implications from this study. First, even if protest risk fails to serve as a deterrent in non-democracies, it may still do so in democracies. In non-democracies, protest initiation is most likely when fraud is very high and incumbent vote-shares very low. The likelihood of protest initiation in a democracy may well be higher, triggering at lower levels of fraud and higher incumbent vote-shares, since opposition actors are likely to be better resources in a liberal democracy (and the government more constrained). At the same time, if the theory described here is accurate, liberal democracies would be constrained against using large-scale election fraud even in the absence of protest risk: they lack the patronage resources and arbitrary enforcement of laws needed for such a scenario.

It is also worth noting that mass electoral protest can serve an illiberal purpose, as illustrated by the events of January 6, 2020 in the United States or 2022-2023 in Brazil. Indeed, out of the 156 instances of post-election protest in this dataset, 10 occurred after relatively clean elections (*intentional irregularities* < 0)\textemdash a small but not minuscule percentage. These incidents demonstrate that, while crying fraud may sometimes be a useful tool for protest entrepreneurs, losing an election alone can be a catalyst for protest, irrespective of the actual level of election manipulation. This possibility tends to reduce the 'opportunity cost' to incumbents of engaging in real fraud, while also underlining the fact that the initiation and course of protest is often driven more by the relative resources and opportunities available to the disaffected and the incumbent government.

<!-- This project suggests additional avenues for future research. It is clear that electoral protests charged by allegations of manipulation do occur at times. While the main argument of this paper is that election fraud itself does not predict these protests, there may still be connections between manipulation and protest to uncover. For example, it could be that other forms of manipulation\textemdash perhaps intimidation or violence\textemdash are associated with higher risk of escalating protest. Perhaps some forms of electoral manipulation are particularly offensive to certain groups\textemdash young people, highly educated people, people with ideological commitments to democracy, etc. To the extent that countries vary according to the size of such populations, they may vary in their underlying protest risk. -->

These results are consistent with different specifications presented in an online appendix. In particular, the main results are supported when data from the Mass Mobilization dataset are used in place of the ECAV data, and when alternative measures of election integrity are used in place of the measure of election fraud. They are also supported when electoral democracies are included in the dataset. The relationship between fraud levels and incumbent defeat is tested. Additional robustness checks include models with country fixed-effects and alternative specifications for selected control variables.

# Conclusion

There are competing views over the nature of election fraud, incumbent strength, and protest. The risk of post-election protest has a central role in several prominent models of electoral manipulation, serving to deter manipulation and to incentivize the toleration of election monitors and other restraints on the ruling party. Others contend that election fraud can indicate incumbent strength. Building on the principal-agent literature of election fraud, I argue that large-scale fraud is indicative of incumbent resilience even when regime popularity (and thus vote-shares) are low. Conducting election fraud requires consolidated control over patronage resources and the ability to protect front-line agents from the consequences of illegal behavior; this capacity also helps insulate incumbents from the effects of mass protest. Resolving this tension in the literature has important implications for the study of election fraud and authoritarian stability.

This paper attempts to test the assumptions of these two models by investigating the relationship between pre-election risk and election fraud, as well as between election fraud and protest occurrence and duration. Drawing on data from the V-Dem, NELDA, and ECAV datasets, it uses cross-national data on 661 elections in electoral authoritarian regimes to show that 1) when pre-election risks to incumbents are higher, subsequent fraud increases rather than decreases, 2) protest occurrence is most likely when incumbent vote-shares are very low and election fraud is high, and 3) incumbents that engaged in larger-scale election fraud are able to more quickly defeat protest movements.

All together, these results suggest that protest risk largely fails to serve as a constraint on election manipulation in authoritarian regimes. The absence of protest risk as a guardrail suggests that mass demonstrations in favor of advancing (or restoring) democratic elections is unlikely to be driven by the severity of manipulation, and that electoral authoritarianism as a regime-type is likely to be a stubborn equilibrium. This is a gloomy picture for supporters of democratization, and suggests that researchers should consider other mechanisms beyond mass protest in their models of electoral manipulation.

\newpage
# References