### 2025-11-14 Output/appendix_v4.Rmd
Restructuring sections; rewriting fitted vs. residuals section to clarify and focus on addressing endogeneity concern.

---
title: "Appendix: A fragile guardrail---The minimal effect of election fraud on protest intensity in non-democracies"
author: "Cole J. Harvey (Oklahoma State University)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    bookdown::pdf_document2: default
linkcolor: blue
bibliography: C:/Users/colej/Documents/Research projects/bib_ch.bib
#biblio-style: "apalike"
fontsize: 12pt
link-citations: yes
toc: false
csl: C:/Users/colej/Documents/Research projects/ji_manipulation_protest/american-journal-of-political-science.csl
header-includes: 
  - \usepackage{booktabs}
  - \usepackage{longtable}
---

```{r setup, include = FALSE}
library(tidyverse)
library(countrycode)
library(readxl)
library(lme4)
library(interplot)
library(lubridate)
library(summarytools)
library(stargazer)
library(ggeffects)
library(survival)
library(modelsummary)
library(glmmTMB)
library(huxtable)
library(kableExtra)
library(simPH)
library(broom)
library(broom.helpers)

nelda2.sub <- read.csv(here::here("Data", "nelda2_sub_2024-02-06.csv"))

nelda2.sub <- nelda2.sub %>% mutate(legit_no_proc = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol))
nelda2.sub <- nelda2.sub %>% mutate(legit_total = (v2exl_legitperf + v2exl_legitlead + v2exl_legitideol + v2exl_legitratio))

##Naming nelda variables

nelda2.sub <- nelda2.sub %>% rename(boycott = nelda14)
nelda2.sub <- nelda2.sub %>% mutate(boycott = ifelse(boycott == "yes", 1, ifelse(boycott == "no", 0,
                                                                                 NA)))

##Creating economic crisis variable

nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = ifelse(is.na(e_miinflat) == T | is.na(gdpgro.1lag) == T, NA, ifelse(e_miinflat > 15 | gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis = as.factor(econ.crisis))

nelda2.sub <- nelda2.sub %>% mutate(gdp.crisis = ifelse(is.na(gdpgro.1lag) == T, NA, ifelse(gdpgro.1lag <= -0.012, 1, 0)))
nelda2.sub <- nelda2.sub %>% mutate(inflat.crisis = ifelse(is.na(e_miinflat) == T, NA, ifelse(e_miinflat > 15, 1, 0)))

nelda2.sub <- nelda2.sub %>% mutate(grievance.index = ifelse(is.na(e_migdppcln) == T, NA, e_migdppcln*v2xeg_eqdr))

## Removing cases where inc. loses
#nelda2.sub <- nelda2.sub %>% filter(inc.lose == 0)  #See how this affects things

##Intmon as a factor
nelda2.sub <- nelda2.sub %>% mutate(v2elintmon = as.factor(v2elintmon))

nelda2.sub <- nelda2.sub %>% mutate(electoral.dem = ifelse(v2x_regime == 2, 1, 0))

#Square of physint

nelda2.sub <- nelda2.sub %>% mutate(physintlag.sq = physinteg.1lag^2)

## Categorical variable for incumbent status

nelda2.sub <- nelda2.sub %>% mutate(incumb.status = as.factor(ifelse(inc.termlimit == "yes" & chosen.successor == 1, "Chosen successor",
                                                           ifelse(inc.termlimit == "yes" & chosen.successor == 0, "Contested exec. election", 
                                                                  ifelse(inc.termlimit == "no", "Executive eligible", "Legislative election"))))) 
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Legislative election"))

#Coding as factors

nelda2.sub <- nelda2.sub %>% mutate(presidential = factor(presidential))
#nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = factor(inc.termlimit.num))
nelda2.sub <- nelda2.sub %>% mutate(rentierstate = factor(rentierstate))
nelda2.sub <- nelda2.sub %>% mutate(taxstate = factor(taxstate))
nelda2.sub <- nelda2.sub %>% mutate(unscheduled_election = factor(unscheduled_election))
nelda2.sub <- nelda2.sub %>% mutate(inc.lose = factor(inc.lose))
nelda2.sub <- nelda2.sub %>% mutate(econ.crisis.nelda.1lag = factor(econ.crisis.nelda.1lag))

#Successful protest = new elections or new incumbent

nelda2.sub <- nelda2.sub %>% mutate(protest.success = ifelse(inc.replaced.prot == 1 | new.elections.held == 1, 1, 0))

# Replacing NA with 0 for term limits variable in known parliamentary countries in legislative elections

nelda2.sub <- nelda2.sub %>% mutate(inc.termlimit.num = ifelse(is.na(inc.termlimit.num) == T & v2ex_elechog == 0, 0, inc.termlimit.num))

## Replacing NA with 0 for protest size

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(is.na(max_size)==T & n.events.post == 0, 0, max_size))

## Fixing missing values of winner.share.upd for liberal democracies

nelda2.sub <- nelda2.sub %>% mutate(winner.share.upd = ifelse(is.na(winner.share.upd)==T & v2x_regime == 3, winner.margin, winner.share.upd))

## Set max_size zero for elections with no protest

#nelda2.sub <- nelda2.sub %>% mutate(max_size = ifelse(n.events.post == 0, 0, max_size))

## Excluding liberal democracies

nelda2.sub <- nelda2.sub %>% filter(v2x_regime == 1 | regime.1lag == 1) #Includes only authoritarian, add e-dems with | v2x_regime == 2  #Current version includes lagged 1 in order to capture pre-election auth.

## Remove outlier
nelda2.sub <- nelda2.sub %>% filter(d.events.num <= 100)

# Clean inc lose
nelda2.sub <- nelda2.sub %>% mutate(inc.lose_clean = as.factor(ifelse(inc.lose == "<NA>", NA, inc.lose)))

#Protest cases only

protest.cases <- nelda2.sub %>% filter(any.ecav.protest == 1 & inc.lose == 0)

## Coding median size as 0 when n.events = 0

nelda2.sub <- nelda2.sub %>% mutate(median.participants = ifelse(n.events.post == 0, 0, median.participants))

## Any protest

nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(n.events.post > 0, 1, 0))

## Surv status
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))

```

# Summary statistics

Table \@ref(tab:summary-table) provides summary statistics for the variables used in the main text.

```{r, include=F}
incl.vars <- nelda2.sub %>% dplyr::select('Judicial independence' = jucon.1lag, 
                                            'Legislative constraints' = legcon.1lag, 
                                            'Alternative information' = altinf.1lag,
                                          'Liberal democracy index' = libdem.1lag,
                                            'GDP growth rate' = gdpgro.1lag, 
                                            'Civil society openness' = osorg.1lag, 
                                            #'Public sector corruption' = pubseccorrup.lag,
                                            'National party organization' = natlpartyorg.lag, 
                                             'Election fraud' = v2elirreg.inv, 
                                            'Presidential election' = presidential,
                                            'Incumbent term-limited' = inc.termlimit.num,
                                            'Physical integrity index' = physinteg.1lag,
                                            'Urbanization' = urban.pct, 
                                            'Rentier state' = rentierstate,
                                            'Tax state' =  taxstate,
                                            'Winning vote-share' = winner.share.upd, 
                                            'Number of pre-election protests' = n.events.pre,
                                            'Off-schedule election' = unscheduled_election,
                                          'Regional protest diffusion' = regional_success_lag,
                                         "Privatization" =  state.owned.1lag,
                                         "Professional bureaucracy" = patronage.appts.1lag,
                                        "EMB autonomy"= v2elembaut)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))

incl.vars <- incl.vars %>% mutate(`Presidential election` = as.numeric(`Presidential election`))
incl.vars <- incl.vars %>% mutate(`Rentier state` = as.numeric(`Rentier state`))
incl.vars <- incl.vars %>% mutate(`Tax state` = as.numeric(`Tax state`))
incl.vars <- incl.vars %>% mutate(`Off-schedule election` = as.numeric(`Off-schedule election`))

```

```{r summary-table, echo=FALSE, message=FALSE, warning=FALSE}
library(kableExtra)
datasummary_skim(incl.vars, output = "kableExtra", title = "Summary statistics and variable distributions", histogram = F)  %>%
    kable_styling(font_size = 9, full_width = FALSE)
```

\newpage

# Robustness check: Two alternative measures of election manipulation

To ensure that the results presented in the main text are not idiosyncratic to the chosen measure of election fraud, this section replaces that measure with two alternative indicators of election integrity. The first variable is `v2xel_frefair,' from V-Dem, which is a comprehensive indicator of quality of the electoral process, including measures of electoral management body autonomy and capacity, vote-buying, intimidation, fraud, and an overall assessment of the fairness of the election. This is a retroactively coded, expert assessment (as is the variable in the main text), but captures a wider array of possible abuses of electoral integrity.

The second measure reflects contemporaneous judgments by election observers, as reported in the Dataset on International Election Monitoring, or DIEM [@kelley2011]. I use the variable "Explicit cheating," as a parallel to the V-Dem "intentional irregularities" variable from the main text. This variable is categorical, and so captures less variation than the original measure. However, it helps address the concern that the findings could be systematically biased by the human coders utilized in the V-Dem project [@Little_Meng_2024]. While there is some evidence that such bias is small [@weidmann2023], it is still useful to check the V-Dem coders' assessments against election monitors' reports.

Both variables produce the same substantive results in the Cox proportional hazards models of protest duration; more election manipulation is associated with shorter protests.

## Overall election integrity

The 'election integrity' variable from V-Dem (v2xel_frefair) ranges from 0 to 1, with higher values indicating freer and fairer elections. Note that the sign of this variable is the inverse of the variable used in the main text. The models presented below utilize this variable in place of the main measure of election fraud, while retaining all other control and explanatory variables.

The same interaction relationship between overall election integrity, incumbent vote-share, and protest as is shown for fraud in the main text. Table \@ref(tab:cox-models-tab-overall) shows that cleaner elections are associated with a slower end to protest movements. When elections are more unfair overall, incumbents are able to more quickly bring protest to an end. Figure \@ref(fig:margin-dur-fig-overall) shows that this effect is significant in interaction with low-to-moderate incumbent vote-shares.

```{r models-hurdle-fulldv, warning = FALSE, message = FALSE, include = F, echo=FALSE}

model.margin.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary.full)

model.margin.binary.int.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag  + natlpartyorg.lag + v2xel_frefair +  +    winner.share.upd*v2xel_frefair   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary.int.full)

#summary(model.margin.binary_noc)

```

```{r logit-table2, echo=FALSE, include = FALSE}
models2 <- list()
models2[['Model 1']] <- model.margin.binary.full
models2[['Model 2']] <- model.margin.binary.int.full

modelsummary(models2,  stars = T, title = 'Logit models of protest occurrence', gof_omit = "Log|RMSE", coef_rename = 
               c("lji.lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                 "v2xel_frefair" = "Overall election integrity",
                 "winner.share.upd" = "Incumbent vote-share",
                 "v2xel_frefair:winner.share.upd" = "Election integ. : Incumbent vote-share"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

\newpage

```{r overall-binary, fig.cap="Overall election integrity, winning vote-share, and probability of protest initiation", fig.height=6, echo=FALSE, message=FALSE, fig.align='center', include=FALSE, warning=FALSE}
sjPlot::plot_model(model.margin.binary.int.full, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.12, .46]"))  +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election integrity") + theme_bw()

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}

nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(inc.replaced.prot == 1, 0, surv_status))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))

cox1 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2xel_frefair   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
#summary(cox1)

cox2 <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag  +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + v2xel_frefair*winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + unscheduled_election + regional_success_lag + n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0))
#summary(cox2)
#summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2xel_frefair [.14, .43]"))  #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election free and fair") + theme_bw()

```

```{r cox-models-tab-overall, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 3']] <- cox1
models3[['Model 4']] <- cox2

modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", 
                 "winner.share.upd" = "Incumbent vote-share",
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "v2xel_frefair" = "Overall election integrity",
                 "n.events.post" = "N. post-election protest"
                 ))

```

```{r margin-dur-fig-overall, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H', warning=FALSE}
p.duration.margin
```

\newpage

## Election observers' reports of fraud

This dataset codes the judgments of international election monitoring missions as categorical variables. The explicit cheating variable is categorical, ranging from the 0 for the cleanest elections to 3 for the most fraudulent. While this data is closely tied to the facts on the ground, as reported by election monitors, it covers a much smaller range of cases than V-Dem: of the 661 observations in the full dataset, 471 do not have DIEM values. This is, in part, due to differences in the time periods covered (DIEM ranges from 1980 to 2004, while the main dataset covers 1989 - 2012). This problem is compounded by the fact that most elections in the dataset go un-protested. Of the 156 cases where protest occurs, 110 lack data in DIEM.  While this limits the usefulness of the DIEM variable, it can still provide some confirmatory information for the main analysis.

Table \@ref(tab:diem-table-1) reports two model where the DIEM measure as the explanatory variable in a survival model of protest movements, both as standalone variable and in interaction with the incumbent's vote-share. Here the difficulty with missingness becomes apparent, with only 44 observations in the models. Figure \@ref(fig:margin-dur-fig-diem) shows that higher 'explicit cheating' has a significant effect on protest duration around the fifty-percent vote-share mark, though the effect is small. Given the small number of observations, it is even more reassuring that the expected relationships between fraud and a quick end to protest movements are detected.

```{r, echo=FALSE}

cox1_diem <- coxph(Surv(d.events.num, surv_status) ~ legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  unscheduled_election + regional_success_lag + diem_explicheat_max   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
#summary(cox1_diem)

cox2_diem <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + winner.share.upd   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  unscheduled_election + regional_success_lag + winner.share.upd*diem_explicheat_max   + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" &  n.events.post > 0))
summary(cox2_diem)

```

```{r diem-table-1, echo=FALSE, message=FALSE, warning=FALSE}
models <- list()
models[['Model 5']] <- cox1_diem
models[['Model 6']] <- cox2_diem

modelsummary(models, output = "kableExtra",  stars = T, title = 'Models using DIEM measure of election fraud',  coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity", "diem_explicheat_max" = "Explicit cheating (DIEM)", "n.events.post" = "N. post-election protests", "winner.share.upd" = "Incumbent vote-share"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

```{r margin-dur-fig-diem, fig.cap="Fraud, winning vote-share, and duration of protest (DIEM)", fig.height=7, echo=FALSE, message=FALSE, fig.pos='H', warning=FALSE}
p.duration.margin_diem <- sjPlot::plot_model(cox2_diem, type = "pred", terms = c("winner.share.upd", "diem_explicheat_max [0, 3]"))  #This is the 10th and 90th quantiles
p.duration.margin_diem <- p.duration.margin_diem +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Explicit cheating (DIEM)") + theme_bw() + coord_cartesian(ylim = c(0, 20))
```

# Addressing concerns about selection effects

## Predicted and residual levels of fraud

The main text focuses on elections where the incumbent claims victory. However, classic regime-weakness models also predict that some incumbents who will voluntarily avoid electoral manipulation and accept a loss at the ballot box. If this is the case, it could introduce selection bias into the sample for the duration analyses in the main paper by sorting only the strongest incumbents into the sample that faces protest.

Fully investigating this possibility should be the focus of future work, but some initial analyses are provided here to show that this is not the case. The analysis in this subsection proceeds in three steps. First, using all the observations in the dataset\textemdash winning and losing incumbents\textemdash I estimate the predicted level of election fraud for each case, using only structural variables (i.e. professionalization of the bureaucracy, electoral management body autonomy, urbanization, etc) in an OLS model. Factors associated with deterrence, like the number of pre-election protests are excluded. This should produce an estimate of the level of manipulation the incumbent could produce in the absence of strategic deterrence.

Second, the residuals from this first model are then used as a dependent variable in second model. The residuals reflect the degree to which an incumbent over- or under-produced manipulation, relative to their expected baseline. If the deterrence interpretation is correct (which may lead to selection effects), the indicators of protest risk should be negatively associated with these residuals. In other words, higher protest risk should lead incumbents to under-produce manipulation. Under the regime-strength logic, we might expect the opposite, as incumbents work hard to build and indicate strength in riskier environments. This model is also an OLS model.

Lastly, two logit models of incumbent loss are presented, using the fitted values from the first model and indicators of protest risk as predictors. The fitted values indicated the expected, structural level of electoral manipulation. Here, the deterrence logic would expect that indicators of pre-election protest risk will make incumbents more likely to lose, while the regime-strength model would expect that incumbent loss will be largely driven by structural factors.

Table \@ref(tab:fraud-loss) shows the results. Model 6 shows how structural factors are predictors of the level of observed manipulation. We observe some expected results; for instance, more autonomous EMBs and more professionalized bureaucracies are associated with lower fraud. Model 7 shows that indicators of protest risk\textemdash the number of pre-election protests, and the existence of a recent successful protest movement in the region\textemdash are not associated with lower fraud. In fact, just the opposite: incumbents appear to manipulate more heavily under riskier conditions. This is a key finding, and is evidence against the classic deterrent approach; it also suggests that weaker governments are not voluntarily dialing back manipulation to avoid protest, mitigating concerns about selection effects in the main models.

Models 8 and 9 take the fitted values from the first model and employ them as an explanatory variable for incumbent loss, while controlling for pre-election protest. As both models show, the more incumbents invest in the structural prerequisites of manipulation, the less likely they are to lose elections. By contrast, we observe little or no effect for pre-election protest risk. Model 8 shows no effect for regional protest diffusion on incumbent loss, while Model 9 shows a significant (p = 0.05) but substantively small effect for pre-election protest.

All told, this analysis supports the argument in the main text, which holds that the capacity to manipulate heavily is associated with incumbent durability. The results in this section are consistent with a world in which losing incumbents are not voluntarily surrendering power in the face of a committed opposition (on average), but are failing to produce sufficient election manipulation to win due to structural constraints. This should reduce the concern that the main findings are driven by an endogenous process, in which incumbents who face more organized and capable opposition actors are strategically avoiding protest by producing less fraud.

```{r, include=F, echo=FALSE, message=FALSE, warning=F}
## Anova
anova_dat <- nelda2.sub %>% dplyr::select(inc.replaced.prot, v2elirreg.inv,  jucon.1lag, legcon.1lag, gdpgro.1lag,  osorg.1lag, natlpartyorg.lag,  urban.pct, n.events.pre, state.owned.1lag, patronage.appts.1lag, v2elembaut)

# boxplots and t-tests for the 4 variables at once
for (i in 2:ncol(anova_dat)) { # variables to compare are variables 2 to end
  boxplot(anova_dat[, i] ~ anova_dat$inc.replaced.prot, # draw boxplots by group
    ylab = names(anova_dat[i]), # rename y-axis with variable's name
    xlab = "Incumbent replaced after protest"
  )
  print(paste(c("ANOVA for variable ", (names(anova_dat[i])))))
  print(t.test(anova_dat[, i] ~ anova_dat$inc.replaced.prot)) # print results of t-test
}

aov_output <- aov(state.owned.1lag ~ inc.replaced.prot, data = nelda2.sub)
summary(aov_output)
```

```{r, echo=FALSE, message=FALSE, include=FALSE, warning=FALSE}
m.fraud.pred <- lm(v2elirreg.inv ~ jucon.1lag +  gdpgro.1lag + presidential   +   urban.pct + rentierstate + taxstate + physinteg.1lag + physintlag.sq + state.owned.1lag + patronage.appts.1lag  + v2elembaut, data = nelda2.sub)

summary(m.fraud.pred)

nelda2.sub.fraudpred <- nelda2.sub %>% mutate(.rownames = rownames(nelda2.sub))

new_frame <- augment(m.fraud.pred)
new_frame <- new_frame %>% dplyr::select(.fitted, .rownames, .resid)

nelda2.sub.fraudpred <- nelda2.sub.fraudpred %>% left_join(new_frame, by = ".rownames")

model.lose_reg <- glm(inc.lose ~ .fitted + regional_success_lag , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
summary(model.lose_reg)

model.lose_events <- glm(inc.lose ~ .fitted + n.events.pre , data = nelda2.sub.fraudpred, family = binomial(link="logit"))
summary(model.lose_events)

model.fraud.resid <- lm(.resid ~ n.events.pre +  regional_success_lag , data = nelda2.sub.fraudpred)
summary(model.fraud.resid)
```

```{r fraud-loss, echo=FALSE}
models2 <- list()
models2[['Model 6']] <- m.fraud.pred
models2[['Model 7']] <- model.fraud.resid
models2[['Model 8']] <- model.lose_reg
models2[['Model 9']] <- model.lose_events

row <- data.frame(cbind("", "Fraud", "Fraud residual", "Inc. loss", "Inc. loss"))
attr(row, 'position') <- c(0)

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Models of manipulation and residuals (OLS) and incumbent loss (logit)', coef_omit = "as.factor", gof_omit = "RMSE|Log.", coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", "v2elirreg.inv" = "Election fraud",
                 "winner.share.upd" = "Incumbent vote-share",
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                 "state.owned.1lag" = "Private ownership (lag)",
                 "patronage.appts.1lag" = "Professional bureaucracy (lag)",
                 "v2elembaut" = "EMB autonomy",
                 ".fitted" = "Expected fraud level"
                 ), add_rows = row) %>% kable_styling(latex_options = "HOLD_position")

## Figure out how to Add a row beneath the model numbers, with the DV
```

\newpage

# Incumbent defeat by protest

The classic model of election manipulation as an indicator of weakness would suggest that incumbents who rely more on manipulation will be more likely to fall when protest movements do emerge. The main text does not examine successful protests, which are a small minority of cases, in order to avoid measurement error in the duration models. Successful protest is, instead, examined here; there is little to suggest that incumbents who manipulate more are more vulnerable to defeat by protesters.

First, Figure \@ref(fig:successplot) shows cases of 'successful' and unsuccessful protest at varying levels of election fraud. Protests are coded as successful if they lead to incumbent ouster or to new elections as recorded in NELDA. As the figure shows, the large majority of protests are unsuccessful by this measure. Moreover, the small number of successful cases appear spread over approximately the same range of election fraud as unsuccessful cases; we do not observe a clustering of successful protest at higher levels of fraud.

Next, Table \@ref(tab:oust-mods) reports three logit regression models of successful protest; the dataset for these models includes all cases with a least one protest event. Model 9 is a minimal model that includes *intentional election irregularities*, *incumbent vote-share* and a control for *presidential elections*. Model 10 includes a larger number of control variables, while Model 11 adds the interaction term. As the results in the table show, greater election manipulation is not associated with a higher likelihood of incumbent defeat in any model. Figure \@ref(fig:oust-plot) shows no significant effect of election manipulation on protest success across all values of incumbent vote-share.  Taken together, this is further disconfirmation of the fraud-as-weakness model.

```{r successplot, fig.cap="Election fraud and successful protest", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}

ggplot(protest.cases, aes(v2elirreg.inv, protest.success)) +
  geom_jitter(width = 0, height = 0.1)+
  xlab("Electoral fraud") +
  ylab("Successful protest (jittered)") +
  theme_bw()
```

```{r oust-mods, echo=FALSE, message=FALSE, warning=FALSE}

model.oust3 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag  + winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag + v2elirreg.inv , family = binomial(link = "logit"), data = protest.cases %>% filter(inc.lose_clean == "1"))
summary(model.oust3)

model.oust4 <- glm(protest.success ~  osorg.1lag + gdpgro.1lag +  winner.share.upd  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +   unscheduled_election + taxstate + rentierstate + regional_success_lag +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
summary(model.oust4)

model.oust5 <- glm(protest.success ~  presidential +  v2elirreg.inv*winner.share.upd , family = binomial(link = "logit"),  data = protest.cases %>% filter(inc.lose_clean == "1"))
summary(model.oust5)

models2 <- list()
models2[['Model 9']] <- model.oust5
models2[['Model 10']] <- model.oust3
models2[['Model 11']] <- model.oust4

modelsummary(models2, output = "default",  stars = T, title = 'Logit models of incumbent ouster by protest', coef_omit = "SD", gof_omit = "AIC|BIC|Log|F|RMSE", coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "state.owned.1lag" = "Privatization",
                 "patronage.appts.1lag" = "Professional bureaucracy",
                 "v2elirreg.inv" = "Election fraud",
                 "winner.share.upd" = "Winner vote-share"
                 ))
```

```{r oust-plot, fig.cap="Fraud, winning vote-share, and incumbent ouster", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
sjPlot::plot_model(model.oust5, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.4, 1.88]")) + labs(x = "Winner vote-share", y = "Pred. probability of successful protest", color = "Election fraud") + theme_bw() 
```

# Chosen successors

The models in the main text control for the presence of term-limited executive incumbents, based on the logic in Hale [@hale_patronal_2014]. However, it could be the case that the nature of a term-limited election affects the behavior of protesters and election manipulators alike; that is, perhaps when there is a designated successor, the heir apparent is better able to mobilize the resources of the state to manipulate elections and suppress protest. Using a variable from NELDA that codes for designated successors, the following models add a categorical variable for the incumbent's status, which can take on the following values: *executive eligible*, *chosen successor*, *contested executive election*, and *legislative election*. Including this variables requires the exclusion of the original variables *presidential election* and *incumbent term limited*.

```{r successor-table, echo=FALSE, message=FALSE, warning=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(incumb.status = relevel(incumb.status, ref = "Executive eligible"))
nelda2.sub <- nelda2.sub %>% mutate(no_protest = ifelse(n.events.post == 0, 1, 0))

## Protest binary

model.margin.int.suc <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + incumb.status  +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate + n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.int.suc)

## Duration

cox2.suc <- coxph(Surv(d.events.num, surv_status) ~  legcon.1lag + altinf.1lag +  osorg.1lag + gdpgro.1lag + natlpartyorg.lag + incumb.status +  physinteg.1lag + physintlag.sq + urban.pct + rentierstate + taxstate +  unscheduled_election + regional_success_lag + v2elirreg.inv*winner.share.upd  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
#summary(cox2.suc)

### Table

models4 <- list()
models4[['Model 13']] <- model.margin.int.suc
#models4[['Model 6']] <- model.d.base2.hurd
models4[['Model 14']] <- cox2.suc

t5a <- huxreg(models4, tidy_args = c("component" = "cond"), borders = 0, number_format = 2, coefs = c("Legislative election" = "incumb.statusLegislative election", "Chosen successor" = "incumb.statusChosen successor", "Contested exec. election" = "incumb.statusContested exec. election", "Intentional irregularities" = "v2elirreg.inv", "Incumbent vote-share" = "winner.share.upd", "Intentional irregularities : Incumbent vote-share" = "v2elirreg.inv:winner.share.upd"),  omit_coefs = c("sd__(Intercept)", "sd__Observation"), statistics = character(0)) %>% set_caption("Models of number and size of protests") %>% insert_row("", "Protest initiation", "Protest end", after = 0) %>% insert_row("Controls", "Yes", "Yes", after = 14)
theme_compact(t5a, header_rows = TRUE, header_cols = FALSE)

```

```{r n-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
dat <- ggpredict(model.margin.int.suc, terms = c("winner.share.upd [0, .1, .2, .3, .4, .5, .6, .7, .8, .9, 1]", "v2elirreg.inv [-.45, 1.64]"), condition = c(incumb.status = "Executive eligible")) 
p.n.margin <- plot(dat)
p.n.margin <- p.n.margin +  labs(x = "Incumbent vote-share", y = "Probability", title = "Predicted probability of protests", color = "Election fraud") + theme_bw()
```

```{r duration-protests-margin, echo=FALSE, message=FALSE, warning=FALSE}
p.s.margin <- sjPlot::plot_model(cox2.suc, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.45, 1.64]")) +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio of protest end", color = "Election fraud") + theme_bw()

```

As Figure \@ref(fig:successor-fig) show, the results are largely similar to those in the main text. For these figures, *incumbent status* was set to an executive election in which the incumbent is eligible. In this model, higher levels of fraud increase the likelihood of protest following the election, at low levels of incumbent vote-share; this is in line with expectations of both the deterrence and incumbent-strenght models. However, the results for protest survival models support the incumbent-strength model.

Table \@ref(tab:successor-table), abridged here to show the key variables, shows that there is no effect of successor type on the duration of electoral protest. There is, however, the same relationship between election fraud, incumbent vote-share, and protest survival that is found in the main text. That is, incumbents who report low vote-shares but make use of large-scale fraud are better able to end protest movements than those who produce less fraud.

```{r successor-fig, fig.cap="Fraud, winning vote-share, and protest initiation / defeat", fig.height=6, echo=FALSE, message=FALSE, fig.align='center'}
plot.combined.prob <- cowplot::plot_grid(p.n.margin, p.s.margin, ncol=1)
plot.combined.prob
```

\newpage

# Removing controls to address multicollinearity

The models in the main text include a suite of control variables aimed at controlling for common causes of election manipulation and protest, in order to minimize the risks of confounding and omitted variable bias. However, this approach runs the risk that individual coefficients may be highly sensitive to model specification due to multicollinearity. In particular, many of the control variables\textemdash such as *legislative constraints*, *judicial independence*, and *physical integrity* are likely to be correlated with pre-election protest. To address this risk, the models in Tables \@ref(tab:logit-table-small) and \@ref(tab:cox-models-tab-small) exclude non-electoral control variables. That is, they include variables related to the election itself, the measure of *pre-election protest*, and the explanatory variables from the main models.

The results echo those from the main text. Fraudulent elections are associated with a higher risk of protest initiation at low levels of ruling-party vote-share (Figure \@ref(fig:margin-prob-fig-small)), though these same conditions are associated with a higher likelihood that protest movements will fail quickly \@ref(fig:margin-dur-fig-small).

```{r include=FALSE, echo=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.margin.binary <- glm(any_protest ~  v2elirreg.inv +  winner.share.upd  + presidential + inc.termlimit.num  +  n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary)

model.margin.binary.int <- glm(any_protest ~ v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +  n.events.pre + unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary.int)

## Plots

####Margin

#plot_binary_margin_max <- sjPlot::plot_model(model.margin.binary, type = "pred", terms = c("winner.share.upd [all]", "v2elirreg.inv [-2.3, 3.2]"))
#plot_binary_margin_max <- plot_binary_margin_max +  labs(x = "", y = "Predicted probability", title = "Predicted probability of any protest", color = "Election fraud") + theme_bw()

plot_binary_margin_typ_n <- sjPlot::plot_model(model.margin.binary.int, type = "pred", terms = c( "winner.share.upd", "v2elirreg.inv [-.05, 2]"), show.data = T)
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Incumbent vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()

```

\newpage

```{r logit-table-small, echo=FALSE}
models2 <- list()
models2[['Model 15']] <- model.margin.binary
models2[['Model 16']] <- model.margin.binary.int

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Logit models of protest occurrence', coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity",
                 "v2elirreg.inv" = "Election fraud",
                 "winner.share.upd" = "Incumbent vote-share"
                 )) %>% kable_styling(latex_options = "HOLD_position")
```

\newpage

```{r margin-prob-fig-small, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
plot_binary_margin_typ_n
```

```{r, include=FALSE, echo=FALSE}
library(survival)
library(ggsurvfit)
nelda2.sub <- nelda2.sub %>% mutate(surv_status = ifelse(d.events.num > 0, 1, NA))
nelda2.sub <- nelda2.sub %>% mutate(surv_status_censor_assume = ifelse(d.events.num == 0, NA, 
                                                                       ifelse(d.events.num >= 69, 0, 1)))

cox1 <- coxph(Surv(d.events.num, surv_status) ~   winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag + v2elirreg.inv  + n.events.post , data = nelda2.sub %>% filter(inc.lose == "0" & n.events.post > 0 & inc.replaced.prot == 0))
#summary(cox1)

cox2 <- coxph(Surv(d.events.num, surv_status) ~  v2elirreg.inv*winner.share.upd  + presidential + inc.termlimit.num  +   unscheduled_election + regional_success_lag +  n.events.post, data = nelda2.sub %>% filter(inc.lose == "0") %>% filter(d.events.num > 0 & inc.replaced.prot == 0))
#summary(cox2)
p.duration.margin <- sjPlot::plot_model(cox2, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [-.05, 2]"))  #This is the 10th and 90th quantiles
p.duration.margin <- p.duration.margin +  labs(x = "Incumbent vote-share", y = "Hazard ratio", title = "Hazard ratio for protest movement end", color = "Election fraud") + theme_bw()

```

```{r cox-models-tab-small, echo=FALSE, warning=FALSE}
models3 <- list()
models3[['Model 17']] <- cox1
models3[['Model 18']] <- cox2

modelsummary::modelsummary(models3, gof_omit = "AIC|BIC|Log", longtable=F, stars = T, title = 'Cox proportional hazard models of protest duration', coef_rename = 
               c("winner.share.upd" = "Incumbent vote-share",
                 "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "unscheduled_election" = "Off-schedule election",
                 "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "regional_success_lag" = "Regional protest diffusion", "v2elirreg.inv" = "Election fraud",
                 "n.events.post" = "N. post-election protests"
                 ))

```

\newpage

```{r margin-dur-fig-small, fig.cap="Fraud, winning vote-share, and duration of protest", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
p.duration.margin
```

\newpage

# Using country fixed effects

Models that include fixed effects by country can be used to estimate the effects of explanatory variables within countries, rather than being driven by a subset of cases where fraud and protest both tend to be high (for example). Table \@ref(tab:logit-table-fe), where logit models of protest initiation are presented, shows results consistent with those in the main text. Fraud alone has no significant relationship with protest initiation, but it does in the context of low incumbent vote-share. Fixed effects models are not employed for the proportional hazards models of protest survival; due to the small number of cases for which multiple protest waves occur, there is insufficient within-country variation to make use of such models.

```{r include=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.margin.binary.fe <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd + v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  n.events.pre + unscheduled_election + regional_success_lag + as.factor(country), family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary)

model.margin.binary.int.fe <- glm(any_protest ~ jucon.1lag + legcon.1lag + gdpgro.1lag + altinf.1lag + libdem.1lag +  + osorg.1lag  + natlpartyorg.lag + v2elirreg.inv +  +    winner.share.upd*v2elirreg.inv  + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct +  n.events.pre + unscheduled_election + regional_success_lag +  as.factor(country), family = binomial(link = "logit"), data = nelda2.sub %>% filter(inc.lose == "0"))
#summary(model.margin.binary.int)

```

```{r logit-table-fe, echo=FALSE}
models2 <- list()
models2[['Model 19']] <- model.margin.binary.fe
models2[['Model 20']] <- model.margin.binary.int.fe

modelsummary(models2, output = "kableExtra",  stars = T, title = 'Table 3: Logit models of protest occurrence', coef_omit = "as.factor", gof_omit = "RMSE|Log.", coef_rename = 
               c("jucon.1lag" = "Judical independence (lag)", "legcon.1lag" = "Leg. constraints (lag)", 
                 "gdpgro.1lag" = "GDP growth rate (lag)", "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)", "econ.crisis.nelda.1lag1" = "Economic crisis", 
                 "natlpartyorg.lag" = "National party organization (lag)", "presidential1" = "Presidential election",
                 "inc.termlimit.num" = "Incumbent term-limited", "libdem.1lag" = "Liberal democracy index (lag)", "v2elirreg.inv" = "Election fraud",
                 "winner.share.upd" = "Incumbent vote-share",
                 "urban.pct" = "Urbanization", "rentierstate" = "Rentier state", "taxstate" = "Tax state",
                 "n.events.massmob.pre" = "Number of pre-election protests", "median.participants.mm.pre" = "Median size of pre-election protests", "unscheduled_election" = "Off-schedule election",
                 "rentierstate1" = "Rentier state", "taxstate1" = "Tax state", "n.events.pre" = "Number of pre-election protests", "unscheduled_election1" = "Off-schedule election", "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", "regional_success_lag" = "Regional protest diffusion", "antisys.1lag" = "Anti-system CS activity"
                 ), add_rows = data.frame(cbind("Country fixed effects", "Y", "Y"))) %>% kable_styling(latex_options = "HOLD_position")
```

# Fraud and the likelihood of protest

Table \@ref(tab:logit-table) shows the results for Hypothesis 1. Model 1 is a base model, where incumbent loss, winner vote-share, and *election fraud* are standalone variables. In keeping with the general (though inconsistent) findings in the prior literature, there is a significant positive relationship between fraud severity and the likelihood of protest initiation. The winner's claimed margin of victory is negatively associated with protest initiation. Model 2 includes the three-way interaction between *election fraud*, incumbent loss, and winning vote-share, with no controls other than a dummy variable for *presidential elections*. This allows for the largest possible sample size, and shows that the coefficient signs are consistent with those of the full Model 3.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
nelda2.sub <- nelda2.sub %>% mutate(any_protest = ifelse(d.events.num > 0, 1, 0))
#nelda2.sub <- nelda2.sub %>% mutate(no_size = ifelse(max_size == 0, 1, 0))

model.binary_noint <- glm(any_protest ~ inc.lose_clean + v2elirreg.inv + winner.share.upd, data = nelda2.sub, family = binomial(link = "logit"))

model.binary.base <-  glm(any_protest ~ winner.share.upd*v2elirreg.inv*inc.lose_clean + presidential, data = nelda2.sub, family = binomial(link = "logit"))

model.binary.full <- glm(any_protest ~ lji.lag + legcon.1lag + gdpgro.1lag + altinf.1lag  +  + osorg.1lag   +    winner.share.upd*v2elirreg.inv*inc.lose_clean   + presidential + inc.termlimit.num  +  physinteg.1lag + physintlag.sq + urban.pct + n.events.pre +  unscheduled_election + regional_success_lag, family = binomial(link = "logit"), data = nelda2.sub)

plot_binary_margin_typ_n <- sjPlot::plot_model(model.binary.full, type = "pred", terms = c("winner.share.upd", "v2elirreg.inv [.08, 2.07]", "inc.lose_clean"))
plot_binary_margin_typ_n <- plot_binary_margin_typ_n +  labs(x = "Winner's vote-share", y = "Predicted probability", title = "Predicted probability of protest", color = "Election fraud") + theme_bw()
```

```{r logit-table, echo=FALSE}
models2 <- list()
models2[['Model 1']] <- model.binary_noint
models2[['Model 2']] <- model.binary.base
models2[['Model 3']] <- model.binary.full
#models2[['Model 6']] <- model.margin.binary_noc

modelsummary(models2, output = "default",  stars = T, title = 'Table 3: Logit models of protest occurrence', gof_omit = c("Log|RMSE"), coef_map = c("(Intercept)" = "(Intercept)",
             "inc.lose_clean2" = "Incumbent loss",
             "v2elirreg.inv" = "Election fraud",
             "winner.share.upd" = "Winner vote-share",
"winner.share.upd:v2elirreg.inv" = "Winner vote-share:Election fraud",
"winner.share.upd:inc.lose_clean2" = "Winner vote-share:Inc. loss",
"v2elirreg.inv:inc.lose_clean2" = "Election fraud:Inc. loss",
"winner.share.upd:v2elirreg.inv:inc.lose1" = "Election fraud:Winner vote-share:Inc. loss",
  "lji.lag" = "Judical independence (lag)",                                                                 "legcon.1lag" = "Leg. constraints (lag)",                                                                 "gdpgro.1lag" = "GDP growth rate (lag)",                                                                   "altinf.1lag" = "Alternative info. (lag)", 
                 "osorg.1lag" = "Civil soc. openness (lag)",
                 "inc.termlimit.num" = "Incumbent term-limited", 
                 "libdem.1lag" = "Liberal democracy index (lag)", 
                 "urban.pct" = "Urbanization", 
                 "rentierstate1" = "Rentier state", 
                 "taxstate1" = "Tax state", 
                "presidential1" = "Presidential",
                 "unscheduled_election1" = "Off-schedule election",
                 "n.events.pre" = "Number of pre-election protests", 
                 "physinteg.1lag" = "Phys. integrity index (lag)",
                 "physintlag.sq" = "Phys. int. lag squared", 
                 "regional_success_lag" = "Regional protest diffusion",

                 "median.participants" = "Median protest participants",

                "v2elirreg.inv:winner.share.upd"= "Election fraud:Winner vote-share")
                 )

#reg_table |> tab_options(
#  latex.use_longtable = TRUE, 
#  data_row.padding = px(5)
#  ) 
```

Model 3 is the main model, which includes the full set of controls. Given the complexity of the model, Figure \@ref(fig:margin-prob-fig) illustrates the effects of election fraud on protest risk across various levels of winning vote-share and incumbent loss^[All other variables constant at their means (continuous variables) or modes (categorical variables).]. The values for election fraud represent one standard deviation above and below the mean value, or about the 15th and 85th percentiles, respectively.

The direction of the effect supports Hypothesis 1. When an incumbent claims victory (left panel) and their vote-share is low, electoral protest is significantly likelier when fraud is elevated than when the election is relatively clean. Furthermore, protest against winning incumbents is significantly likelier when fraud is high and vote-shares are low, compared to an election with an equivalent level of fraud but high claimed vote-shares. As noted above, such a finding could be consistent with either the regime-weakness or regime-strength models of manipulation, under different conceptualizations of protest. To distinguish between the two, we turn to the models of protest duration.

```{r margin-prob-fig, fig.cap="Fraud, winning vote-share, and probability of protest initiation", fig.height=5, echo=FALSE, message=FALSE, fig.pos='H'}
plot_binary_margin_typ_n$data$facet <-  ifelse(plot_binary_margin_typ_n$data$facet == "1", "Incumbent won", "Incumbent lost")

plot_binary_margin_typ_n
```